---
title: "Fig4_rRNA motifscan visualization"
author: "Jie Hua"
date: 2026-03-29
lang: en
date-modified: 2026-03-29
keywords: ["rDNA Quantification","Hybrid Cells"]
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: true
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load the required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(tidyverse)
library(stringr)
library(data.table)
library(ggplot2)
library(dplyr)
library(tibble)
library(grid)
```

# Define the data folder

```{r}
#| echo: true
#| warning: false

folder <- "F:\\Big Paper\\Processed data\\Fig4_rDNA_motifscan"
#folder <- "H:\\HybridCells_Data\\rDNA_motifscan"
```

# Read the *.csv* files

```{r}
#| echo: true
#| warning: false

files <- list.files(folder, pattern = "\\.csv$", full.names = TRUE)

# 读取并处理
df_all <- map_dfr(files, function(f) {
  
  # 提取文件名
  fname <- basename(f)
  
  # 提取 sample（.rDNA 前面的部分）
  sample_name <- sub("\\.rDNA.*$", "", fname)
  
  # 读取数据
  df <- read_csv(f, show_col_types = FALSE)
  
  # 添加 sample 列
  df %>%
    mutate(sample = sample_name)
})

```

# Extract data for rDNA quantification

```{r}
#| echo: true
#| warning: false

sample_map <- tibble(
  sample = c("Ras3","ML82","5A", "5B_barcode19","6B","7B_barcode23"),
  new_sample = c("Ras3","ML82","HYC1", "HYC2","HYC3","HYC4")
)

df_rDNA <- df_all %>%
  filter(sample %in% sample_map$sample) %>%
  left_join(sample_map, by = "sample") %>%
  mutate(sample = new_sample) %>%
  select(-new_sample)


```

# Extract data for total rDNA

```{r}
#| echo: true
#| warning: false

bam_summary_file <- "F:\\Big Paper\\Processed data\\Fig3_Dosage and Transcriptome\\bam_summary_genome.csv"

bam_summary <- fread(bam_summary_file)



# =============================
# 1. sample mapping
# =============================
sample_map <- tibble(
  sample = c(
    "Ras3",
    "ML82",
    "5A",
    "5B_barcode19",
    "6B",
    "7B_barcode23"
  ),
  bam_sample = c(
    "Ras3_q30",
    "ML82_q30",
    "05a_q30",
    "05b_barcode19_q30",
    "06b_q30",
    "07b_barcode23_q30"
  ),
  new_sample = c(
    "Ras3",
    "ML82",
    "HYC1",
    "HYC2",
    "HYC3",
    "HYC4"
  )
)



# =============================
# 2. normalize by Total_Bases
# =============================
df_norm <- df_all %>%
  filter(sample %in% sample_map$sample) %>%
  left_join(sample_map, by = "sample") %>%
  left_join(
    bam_summary %>%
      as_tibble() %>%
      select(
        bam_sample = Sample,
        Total_Bases
      ),
    by = "bam_sample"
  ) %>%
  mutate(
    Total_Bases = as.numeric(Total_Bases),
    
    # raw normalized value
    normalized_raw = total_hits / Total_Bases,
    
    # easier for plotting: hits per Gb
    normalized = total_hits / Total_Bases * 1e9,
    
    # species from motif name
    species = case_when(
      grepl("_dmel$", motif) ~ "dmel",
      grepl("_dsim$", motif) ~ "dsim",
      TRUE ~ NA_character_
    ),
    
    # final sample name for plotting
    sample = new_sample
  )

# =============================
# 3. plot data
# =============================
sample_colors <- c(
  "HYC1" = "#61649F",
  "HYC2" = "#61649F",
  "HYC3" = "#61649F",
  "HYC4" = "#61649F",
  "ML82" = "#ed556a",
  "Ras3" = "#fb8b05"
)

motif_order <- c(
  "18S_dmel", "28S_dmel",
  "18S_dsim", "28S_dsim"
)

df_plot <- df_norm %>%
  filter(
    motif %in% motif_order,
    !is.na(normalized),
    is.finite(normalized)
  ) %>%
  filter(
    !(sample == "ML82" & species != "dsim"),
    !(sample == "Ras3" & species != "dmel")
  ) %>%
  mutate(
    motif = factor(motif, levels = motif_order),
    sample = factor(
      sample,
      levels = c("Ras3", "ML82", "HYC1", "HYC2", "HYC3", "HYC4")
    )
  )

# Y-axis maximum
y_max <- max(df_plot$normalized, na.rm = TRUE) * 1.1

my_family <- "sans"

# =============================
# 4. plot
# =============================
p1 <- ggplot(df_plot, aes(x = sample, y = normalized, fill = sample)) +
  geom_col(
    width = 0.7,
    colour = NA,
    linewidth = 0.1
  ) +
  
  facet_wrap(~motif, ncol = 4) +
  
  scale_fill_manual(values = sample_colors) +
  
  scale_y_continuous(
    expand = c(0.05, 0),
    limits = c(0, y_max),
    breaks = pretty(c(0, y_max), n = 3),
    minor_breaks = waiver(),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  
  labs(
    x = NULL,
    y = "Total hits per Gb"
  ) +
  
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Facet labels
    strip.text.x = element_text(
      angle = 0,
      hjust = 0.5,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 1, r = 0, b = 3, l = 0)
    ),
    strip.text.y.left = element_text(
      angle = 0,
      hjust = 1,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 0, r = 3, b = 0, l = 1)
    ),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      color = "black",
      family = my_family,
      angle = 0,
      hjust = 0.5
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_blank(),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Minor ticks
    axis.minor.ticks.x = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.y = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),
    
    # No legend
    legend.position = "none",
    
    # Facet spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    strip.placement = "outside",
    strip.switch.pad.grid = unit(0, "pt"),
    strip.clip = "off",
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  )

```

Save the plot

```{r}
#| echo: true
#| warning: false

p1_canvas <- p1 + ggview::canvas(180, 50, units="mm",dpi=900,bg="transparent")
p1_canvas

#ggsave( "Fig4A.pdf",p1,dpi=1200,units="mm",width=180, height=50)


```

# Extract data for rDNA quantification per X chromosome

```{r}
#| echo: true
#| warning: false


sample_map <- tibble(
  sample = c("Ras3","ML82","5A", "5B_barcode19","6B","7B_barcode23"),
  new_sample = c("Ras3","ML82","HYC1", "HYC2","HYC3","HYC4")
)

df_rDNA <- df_all %>%
  filter(sample %in% sample_map$sample) %>%
  left_join(sample_map, by = "sample") %>%
  mutate(sample = new_sample) %>%
  select(-new_sample)


df_norm <- df_rDNA %>%
  group_by(sample) %>%
  mutate(
    # baseline
    baseline_dmel = mean(total_hits[str_detect(motif, "^FBgn")], na.rm = TRUE),
    baseline_dsim = mean(total_hits[str_detect(motif, "^LOC")], na.rm = TRUE)
  ) %>%
  ungroup() %>%
  filter(str_detect(motif, "^(18S|28S)")) %>%
  
  mutate(
    species = case_when(
      str_detect(motif, "dmel") ~ "dmel",
      str_detect(motif, "dsim") ~ "dsim"
    ),
    rDNA_type = case_when(
      str_detect(motif, "^18S") ~ "18S",
      str_detect(motif, "^28S") ~ "28S"
    ),
    
    normalized = case_when(
      species == "dmel" ~ total_hits / baseline_dmel,
      species == "dsim" ~ total_hits / baseline_dsim
    )
  )
```

## Plot the copy number per X chromosome

```{r}
#| echo: true
#| warning: false

# 🎨 自定义颜色（稍微优化一下更统一）
sample_colors <- c(
  "HYC1" = "#61649F",
  "HYC2" = "#61649F",
  "HYC3" = "#61649F",
  "HYC4" = "#61649F",
  "ML82" = "#ed556a",
  "Ras3" = "#fb8b05"
)



# 🔥 motif顺序（关键）
motif_order <- c(
  "18S_dmel", "28S_dmel",
  "18S_dsim", "28S_dsim"
)

df_plot <- df_norm %>%
  # ❗去掉 NA / Inf
  filter(!is.na(normalized), is.finite(normalized)) %>%
  
  # ❗控制 ML82 / Ras3
  filter(
    !(sample == "ML82" & species != "dsim"),
    !(sample == "Ras3" & species != "dmel")
  ) %>%
  
  # ❗设置顺序
  mutate(
    motif = factor(motif, levels = motif_order),
    sample = factor(sample, levels = c("Ras3","ML82", "HYC1", "HYC2", "HYC3", "HYC4"))
  )

# Y-axis maximum number
y_max <- max(df_plot$normalized, na.rm = TRUE) * 1.1

my_family <- "sans"

p2 <- ggplot(df_plot, aes(x = sample, y = normalized, fill = sample)) +
  geom_col(
    width = 0.7,
    colour = NA,
    linewidth = 0.1
  ) +
  
  facet_wrap(~motif, ncol = 4) +
  
  scale_fill_manual(values = sample_colors) +
  
  scale_y_continuous(
    expand = c(0.05, 0),
    limits = c(0, y_max),
    breaks = pretty(c(0, y_max), n = 3),
    minor_breaks = waiver(),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  
  labs(
  x = NULL,
  y = expression(Hits^rDNA / Hits^"mean of BUSCO genes")
) +
  
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Facet labels
    strip.text.x = element_text(
      angle = 0,
      hjust = 0.5,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 1, r = 0, b = 3, l = 0)
    ),
    strip.text.y.left = element_text(
      angle = 0,
      hjust = 1,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 0, r = 3, b = 0, l = 1)
    ),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      color = "black",
      family = my_family,
      angle = 0,
      hjust = 0.5
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_blank(),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Minor ticks
    axis.minor.ticks.x = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.y = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),
    
    # No legend
    legend.position = "none",
    
    # Facet spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    strip.placement = "outside",
    strip.switch.pad.grid = unit(0, "pt"),
    strip.clip = "off",
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  )

```

Save the plot

```{r}
#| echo: true
#| warning: false
p2_canvas <- p2 + ggview::canvas(180, 50, units="mm",dpi=900,bg="transparent")
p2_canvas

#ggsave( "Fig4B.pdf",p2,dpi=1200,units="mm",width=180, height=50)
```

# Calculate the rRNA/rDNA ratio

## Extract rRNA and rDNA counts

```{r}
#| echo: true
#| warning: false

sample_map_DNA <- tibble(
  sample = c("Ras3","ML82","5A", "5B_barcode19","6B","7B_barcode23"),
  new_sample = c("Ras3","ML82","HYC1", "HYC2","HYC3","HYC4")
)

df_rDNA <- df_all %>%
  # ❗只保留你指定的样本
  filter(sample %in% sample_map_DNA$sample) %>%
  
  # ❗精确映射
  left_join(sample_map_DNA, by = "sample") %>%
  
  mutate(sample = new_sample) %>%
  
  select(-new_sample)



sample_map_RNA <- tibble(
  sample_core = c("Ras3", "ML82", "5A", "5B", "6B", "7B"),
  new_sample  = c("Ras3", "ML82", "HYC1", "HYC2", "HYC3", "HYC4")
)

df_rRNA <- df_all %>%
  # ❗只选 transcriptome
  filter(str_detect(sample, "transcriptome")) %>%
  
  mutate(
    # 🔥 提取 core
    sample_core = str_extract(sample, "^[^_]+"),
    
    # 🔥 提取 replicate
    rep = str_extract(sample, "rep\\d+")
  ) %>%
  
  # 🔥 映射
  left_join(sample_map_RNA, by = "sample_core") %>%
  
  mutate(
    sample = paste0(new_sample, "_", rep)
  ) %>%
  
  select(-sample_core, -new_sample, -rep)


# Extract to a summary table
df_merge <- df_rRNA %>%
  # 🧩 提取对应 DNA 的 sample（去掉 _repX）
  mutate(sample_dna = str_remove(sample, "_rep\\d+")) %>%
  
  # 🧬 提取 RNA
  select(sample, sample_dna, motif, RNA = total_hits) %>%
  
  # 🔗 和 DNA 对齐
  left_join(
    df_rDNA %>%
      select(sample, motif, DNA = total_hits),
    by = c("sample_dna" = "sample", "motif" = "motif")
  ) %>%
  
  # 🧹 整理输出
  select(sample, motif, RNA, DNA)
```

## Calculate rRNA ratio

```{r}
#| echo: true
#| warning: false

df_result <- df_merge %>%
  filter(str_detect(motif, "^(18S|28S)")) %>%
  
  mutate(
    rDNA_type = str_extract(motif, "^(18S|28S)"),
    species   = str_extract(motif, "dmel|dsim"),
    
    RNA_over_DNA = if_else(DNA == 0, 0, RNA / DNA)
  ) %>%
  
  group_by(sample, rDNA_type) %>%
  
  mutate(
    RNA_percentage = RNA / sum(RNA, na.rm = TRUE),
    
    # 🔥 正确处理（重点）
    relative_RNA_percentage = {
      total <- sum(RNA_over_DNA, na.rm = TRUE)
      if (total == 0) {
        rep(0, n())
      } else {
        RNA_over_DNA / total
      }
    }
  ) %>%
  
  ungroup()
```

## Plot the result of RNA percentage

## RNA percentage

```{r}
#| echo: true
#| warning: false

df_plot <- df_result %>%
  mutate(sample_base = stringr::str_remove(sample, "_rep\\d+")) %>%
  
  group_by(sample_base, rDNA_type, species) %>%
  summarise(
    RNA_percentage = mean(RNA_percentage, na.rm = TRUE),
    .groups = "drop"
  )

sample_levels <- c("Ras3", "ML82", "HYC1", "HYC2", "HYC3", "HYC4")

df_plot <- df_plot %>%
  mutate(sample_base = factor(sample_base, levels = sample_levels))

my_family <- "sans"

p3 <- ggplot(df_plot, aes(
  x = sample_base,
  y = RNA_percentage,
  fill = species
)) +
  geom_col(
    width = 0.7,
    colour = NA,
    linewidth = 0.1
  ) +
  
  facet_wrap(~rDNA_type, ncol = 2) +
  
  scale_fill_manual(
  values = c(
    "dmel" = "#0eb0c9",
    "dsim" = "#f2481b"
    ),
    labels = c(
      "dmel" = "D. mel",
      "dsim" = "D. sim"
    ),
    name = "Source"
  ) +
  
  
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = c(0.05, 0),
    breaks = seq(0, 1, by = 0.5),
    minor_breaks = seq(0, 1, by = 0.1),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  
  labs(
    x = NULL,
    y = "RNA Percentage"
  ) +
  
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Facet labels
    strip.text.x = element_text(
      angle = 0,
      hjust = 0.5,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 1, r = 0, b = 3, l = 0)
    ),
    strip.text.y.left = element_text(
      angle = 0,
      hjust = 1,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 0, r = 3, b = 0, l = 1)
    ),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      color = "black",
      family = my_family,
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_blank(),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Minor ticks
    axis.minor.ticks.x = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.y = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),
    
    # Compact legend
    legend.position = "right",
    legend.title = element_text(
      size = 6,
      color = "black",
      family = my_family,
      face ="italic"
    ),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 0, l = 2)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    # Facet spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    strip.placement = "outside",
    strip.switch.pad.grid = unit(0, "pt"),
    strip.clip = "off",
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  ) +
  guides(
    fill = guide_legend(
      keywidth = unit(6, "pt"),
      keyheight = unit(6, "pt"),
      ncol = 1,
      bycol = TRUE,
      label.position = "right",
      label.hjust = 0,
      override.aes = list(
        colour = NA,
        alpha = 0.95
      )
    )
  )

```

Save the plot

```{r}
#| echo: true
#| warning: false
p3_canvas <- p3 + ggview::canvas(75, 50, units="mm",dpi=900,bg="transparent")
p3_canvas

#ggsave( "Fig4C.pdf",p3,dpi=1200,units="mm",width=75, height=50)

```

## Plot the result of relative RNA percentage

```{r}
#| echo: true
#| warning: false

df_plot_rel <- df_result %>%
  mutate(sample_base = str_remove(sample, "_rep\\d+")) %>%
  
  group_by(sample_base, rDNA_type, species) %>%
  summarise(
    relative_RNA_percentage = mean(relative_RNA_percentage, na.rm = TRUE),
    .groups = "drop"
  )

sample_levels <- c("Ras3", "ML82", "HYC1", "HYC2", "HYC3", "HYC4")

df_plot_rel <- df_plot_rel %>%
  mutate(sample_base = factor(sample_base, levels = sample_levels))

p4 <- ggplot(df_plot_rel, aes(
  x = sample_base,
  y = relative_RNA_percentage,
  fill = species
)) +
  geom_col(
    width = 0.7,
    colour = "grey20",
    linewidth = 0.2
  ) +
  
  facet_wrap(~rDNA_type, ncol = 2) +
  
   scale_fill_manual(values = c("dmel" = "#0D6D98","dsim" = "#FF3636"),name="Source") +
  
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = c(0.05, 0),
    breaks = seq(0, 1, by = 0.5),
    minor_breaks = seq(0, 1, by = 0.1),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  
  labs(
    x = NULL,
    y = "Relative RNA Percentage"
  ) +
  
  theme_bw(base_size = 6) +
  theme(
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    panel.grid = element_blank(),
    
    panel.spacing.x = unit(0.6, "lines"),
    panel.spacing.y = unit(0.5, "lines"),
    
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.7),
    
    strip.text = element_text(size = 7, face = "bold"),
    axis.text.x = element_text(size = 5.5, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 5.5),
    axis.title.y = element_text(size = 7),
    
    axis.ticks = element_line(linewidth = 0.2),
    axis.ticks.length = unit(1.5, "pt"),
    
    axis.minor.ticks.length = unit(0.8, "pt"),
    
    # 新增图例位置（右上角）
    legend.justification = c("right", "top"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = -2),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2),
    legend.title = element_text(size = 6, color = "black"),
    legend.text  = element_text(size = 6, color = "black"),
    legend.key.width  = unit(5, "pt"),
    legend.key.height = unit(5, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.spacing.x = unit(0.5, "pt"),
    strip.clip="off",
    
    plot.margin = margin(4, 2, 0, 2)
  )
```

Save the plot

```{r}
#| echo: true
#| warning: false


p4_canvas <- p4 + ggview::canvas(75, 50, units="mm",dpi=900,bg="transparent")
p4_canvas

#ggsave( "relative rDNA percentage.png",p3,dpi=1200,units="mm",width=75, height=50)

```

# Check the ratio between *D. mel* and *D. sim*

```{r}
#| echo: true
#| warning: false


library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(grid)

# =============================
# 1. 计算 RNA / DNA ratio
# =============================
df_ratio <- df_result %>%
  
  select(sample, rDNA_type, species, RNA, DNA) %>%
  
  pivot_wider(
    names_from = species,
    values_from = c(RNA, DNA),
    names_sep = "_"
  ) %>%
  
  mutate(
    RNA_ratio = RNA_dmel / RNA_dsim,
    DNA_ratio = DNA_dmel / DNA_dsim
  ) %>%
  
  select(sample, rDNA_type, RNA_ratio, DNA_ratio) %>%
  
  arrange(sample, rDNA_type)

# =============================
# 2. 自定义坐标轴
# =============================
x_limits <- c(-2, 2)
y_limits <- c(-7, 7)

x_breaks <- seq(x_limits[1], x_limits[2], by = 1)
y_breaks <- seq(y_limits[1], y_limits[2], by = 2)

x_labels <- x_breaks
y_labels <- y_breaks

# =============================
# 3. 计算每组平均值
#    兼容 HYC1_rep1 / 5A_rep1 等命名
# =============================
df_ratio_mean <- df_ratio %>%
  mutate(
    sample_group = str_remove(sample, "_rep.*"),
    sample_group = case_when(
      sample_group == "5A" ~ "HYC1",
      sample_group == "5B" ~ "HYC2",
      sample_group == "6B" ~ "HYC3",
      sample_group == "7B" ~ "HYC4",
      TRUE ~ sample_group
    )
  ) %>%
  filter(
    sample_group %in% c("HYC1", "HYC2", "HYC4"),
    !is.na(RNA_ratio),
    !is.na(DNA_ratio),
    is.finite(RNA_ratio),
    is.finite(DNA_ratio),
    RNA_ratio > 0,
    DNA_ratio > 0
  ) %>%
  group_by(sample_group, rDNA_type) %>%
  summarise(
    RNA_ratio = mean(RNA_ratio, na.rm = TRUE),
    DNA_ratio = mean(DNA_ratio, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    log2_DNA_ratio  = log2(DNA_ratio),
    log2_rRNA_ratio = log2(RNA_ratio),
    sample_group = factor(
      sample_group,
      levels = c("HYC1", "HYC2", "HYC4")
    )
  )

# 检查：这里不能是 0 行
print(df_ratio_mean)

# =============================
# 4. 如果 df_ratio_mean 为空，提前报错
# =============================
if (nrow(df_ratio_mean) == 0) {
  stop(
    "df_ratio_mean is empty. Check sample names with unique(df_ratio$sample) and rDNA_type/species values."
  )
}

# =============================
# 5. 绘图
# =============================
my_family <- "sans"

p5 <- ggplot(
  df_ratio_mean,
  aes(
    x = log2_DNA_ratio,
    y = log2_rRNA_ratio,
    colour = sample_group
  )
) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    linewidth = 0.25,
    colour = "grey60"
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.25,
    colour = "grey60"
  ) +
  geom_point(
    size = 1.6,
    alpha = 0.9,
    stroke = 0.2
  ) +
  
  facet_wrap(~ rDNA_type, ncol = 2) +
  

  
  scale_colour_manual(
    values = c(
      "HYC1" = "#0072B2",
      "HYC2" = "#fb8b05",
      "HYC4" = "#ed556a"
    ),
    breaks = c("HYC1", "HYC2", "HYC4"),
    name = "Sample"
  ) +
  
  scale_x_continuous(
    limits = x_limits,
    breaks = x_breaks,
    labels = x_labels,
    minor_breaks = seq(x_limits[1], x_limits[2], by = 0.5),
    expand = expansion(mult = c(0.05, 0.05)),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  
  scale_y_continuous(
    limits = y_limits,
    breaks = y_breaks,
    labels = y_labels,
    minor_breaks = seq(y_limits[1], y_limits[2], by = 0.5),
    expand = expansion(mult = c(0.05, 0.05)),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  
  coord_fixed(ratio = 1) +
  
  labs(
    x = expression(log[2]("rDNA " * italic("D. mel") * "/" * italic("D. sim"))),
    y = expression(log[2]("rRNA " * italic("D. mel") * "/" * italic("D. sim")))
  ) +
  
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Facet labels
    strip.text.x = element_text(
      angle = 0,
      hjust = 0.5,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 1, r = 0, b = 3, l = 0)
    ),
    strip.text.y.left = element_text(
      angle = 0,
      hjust = 1,
      size = 6,
      family = my_family,
      color = "black",
      face = "plain",
      margin = margin(t = 0, r = 3, b = 0, l = 1)
    ),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Minor ticks
    axis.minor.ticks.x = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.y = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),
    
    # Compact legend
    legend.position = "right",
    legend.title = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 0, l = 2)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    # Facet spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    strip.placement = "outside",
    strip.switch.pad.grid = unit(0, "pt"),
    strip.clip = "off",
    
    # Plot margin
    plot.margin = margin(t = 0, r = 2, b = 0, l = 2),
    
    # Keep square panels
    aspect.ratio = 1
  ) +
  
  guides(
    colour = guide_legend(
      keywidth = unit(6, "pt"),
      keyheight = unit(6, "pt"),
      ncol = 1,
      bycol = TRUE,
      label.position = "right",
      label.hjust = 0,
      override.aes = list(
        alpha = 0.95,
        size = 1.8
      )
    )
  )
```

Save the plot

```{r}
#| echo: true
#| warning: false
p5_canvas <- p5 + ggview::canvas(75, 50, units="mm",dpi=900,bg="transparent")
p5_canvas

#ggsave( "Fig4D.pdf",p5,dpi=1200,units="mm",width=75, height=50)

```
