---
title: "Fig2. Centromeric Coverage of D.sim contigs"
author: "Jie Hua"
date: 2025-09-29
lang: en
date-modified: 2025-11-20
keywords: ["Centromere","Genomic Coverage","Hybrid Cells"]
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: false
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load the required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(dplyr)
library(data.table)
library(ggplot2)
library(cowplot)
library(viridis) 
library(stringr)
```

# Define the required parameters

```{r}
#| echo: true
#| warning: false

# Define the folder containing the bedgraph files
#folder <- "F:\\Big Paper\\Processed data\\Fig3_Centromeric contigs coverage\\Centromere_Dsim"
# The RepeatMasker annotation
#gff <- "F:\\Big Paper\\Processed data\\Fig3_Centromeric contigs coverage\\Centromere_Dsim\\dsim_centromere_consensus_truncated.fasta.out.gff"
# The RepeatMakser out
#repeatmasker_out <- "F:\\Big Paper\\Processed data\\Fig3_Centromeric contigs coverage\\Centromere_Dsim\\dsim_centromere_consensus_truncated.fasta.out"

# Define the folder containing the bedgraph files
folder <- "H:\\HybridCells_Data\\Fig2_Centromeric Contigs coverage\\Centromere_Dsim"
# The RepeatMasker annotation
gff <- "H:\\HybridCells_Data\\Fig2_Centromeric Contigs coverage\\Centromere_Dsim\\dsim_centromere_consensus_truncated.fasta.out.gff"
# The RepeatMakser out
repeatmasker_out <- "H:\\HybridCells_Data\\Fig2_Centromeric Contigs coverage\\Centromere_Dsim\\dsim_centromere_consensus_truncated.fasta.out"
```

# Load the coverage data

I already used *bamcoverage* to transfer *bam* files to *bed* files.

```{r}
#| echo: true
#| warning: false
cat("The folder containing the bedgraph files is", folder, "\n")
# List the bedGraph files
bedgraph_files <- list.files(folder, pattern = "\\.bedgraph$", full.names = TRUE)
cat("The bedGraph files (total number:", length(bedgraph_files), "):\n")
print(bedgraph_files)
```

Load the *bedgraph* files to R

```{r}
#| echo: true
#| warning: false

data_list <- list()
for (file in bedgraph_files) {
  data <- read.table(file, header = FALSE, sep = "\t", stringsAsFactors = FALSE)
  colnames(data) <- c("Chromosome", "Start", "End", "Counts")
  data$total <- (data$End - data$Start)*data$Counts
  file_name <- basename(file)
  name <- sub(".chopped\\.bedgraph$", "", file_name)
  data_list[[name]] <- data
}
cat("The names of the list containing all the bedgraph tables:\n")
names(data_list)

```

I need to filter the data to select specific samples and chromosomes, and rename them.

```{r}
#| echo: true
#| warning: false
selected_sample <- c("Ras3","ML82","05a","05b_barcode17","06b","07b_barcode21")
renamed_sample  <- c("Ras3","ML82","HYC1","HYC2","HYC3","HYC4")

selected_chromosomes <- c("Contig_14", "Contig_2", "Contig_124", "Contig_137")

renamed_chromosomes <- c("Cen2/3_A_sim", "Cen2/3_B_sim", "Cen4_sim", "CenX_sim")

# Build mapping relationship
sample_map <- setNames(renamed_sample, selected_sample)
chrom_map  <- setNames(renamed_chromosomes, selected_chromosomes)

filtered_list <- lapply(selected_sample, function(smp) {
  df <- data_list[[smp]]
  df <- df[df$Chromosome %in% selected_chromosomes, ]
  df$Chromosome <- chrom_map[df$Chromosome]
  rownames(df) <- NULL
  return(df)
})

names(filtered_list) <- sample_map
```

I want to check if it works.

```{r}
#| echo: true
#| warning: false

print(names(filtered_list))
print(unique(filtered_list$Ras3$Chromosome))
head(filtered_list$Ras3)
```

Merge all tables into one dataframe

```{r}
#| echo: true
#| warning: false

data_all <- bind_rows(
  lapply(names(filtered_list), function(nm) {
    df <- filtered_list[[nm]]
    df$Sample <- nm
    df
  })
)
```

# Normalization

I need to check the distribution of counts first.

```{r}
#| echo: true
#| warning: false
#| fig-align: center
ggplot(data_all, aes(x = Counts, weight = End - Start)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5) +
  facet_wrap(~Sample, scales = "free_y") +
  labs(
    title = "Counts Density Distribution Across Samples",
    x = "Counts",
    y = "Density"
  ) +
  theme_minimal(base_family = "sans") + 
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black")
  )
```

Apparently, I should perform transformation to make it like normal distribution.

```{r}
#| echo: true
#| warning: false

data_all <- data_all %>%
  mutate(Counts_log2 = log2(Counts + 1))
```

Then I could check the distribution with log2 transfered counts

```{r}
#| echo: true
#| warning: false
#| fig-align: center


ggplot(data_all, aes(x = Counts_log2, weight = End - Start)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5) +
  facet_wrap(~Sample, scales = "free_y") +
  labs(
    title = "Log2(Counts+1) Density Distribution Across Samples",
    x = "log2(Counts + 1)",
    y = "Density"
  ) +
  theme_minimal(base_family = "sans") +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 0.6), # 边框稍粗
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black")
  )

```

# Plot the coverage information

```{r}
#| echo: true
#| warning: false


plot_genome_coverage <- function(data_all) {

  # 计算染色体长度并生成标签
  chrom_info <- data_all %>%
    group_by(Chromosome) %>%
    summarise(Length = max(End) - min(Start), .groups = "drop") %>%
    mutate(Chromosome_label = paste0(Chromosome, "\n(", round(Length/1000, 2), " kb)"))

  # 合并标签到数据
  data_all <- data_all %>%
    left_join(chrom_info, by = "Chromosome") %>%
    mutate(
      Chromosome_label = factor(Chromosome_label, levels = chrom_info$Chromosome_label),
      Sample = factor(Sample, levels = unique(Sample))
    )

  # 自定义样本颜色
  sample_colors <- c(
    "Ras3" = "#1f78b4", "ML82" = "#33a02c",
    "HYC1"   = "#e31a1c", "HYC2"   = "#ff7f00",
    "HYC3"   = "#6a3d9a", "HYC4"   = "#b15928"
  )

  p <- ggplot(data_all) +
    geom_rect(
      aes(
        xmin = Start,
        xmax = End,
        ymin = 0,
        ymax = Counts_log2,
        fill = Sample
      ),
      color = NA
    ) +
    facet_grid(
      Sample ~ Chromosome_label,
      scales = "free_x",
      space = "free_x",
      switch ="both"
    ) +
    scale_fill_manual(values = sample_colors) +
    xlab("Chromosome") +
    ylab(expression(log[2](Counts + 1))) +
    theme_bw() +
    theme(
      panel.background = element_rect(fill = "white", color = NA),
      panel.grid = element_blank(),               # 去掉灰色网格线
      panel.spacing = unit(0.2, "lines"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6),
      strip.placement = "outside",
      strip.background = element_blank(),
      strip.text.x = element_text(angle = 0, hjust = 0.5, size = 9), # X轴标签不旋转
      strip.text.y.left = element_text(angle = 0, size = 9), # Sample 标签左边
      axis.text.x = element_blank(),  # X轴刻度数字取消
      axis.ticks.x = element_blank(),
      legend.position = "none",
      strip.clip="off"
    )

  return(p)
}

# 使用示例
plot_genome_coverage(data_all)

```

# Read the annotations

```{r}
#| echo: true
#| warning: false
df_repeat_raw <- read.delim(gff, header = FALSE, stringsAsFactors = FALSE,comment.char = "#")
# Rename the table
colnames(df_repeat_raw) <- c("Chromosome", "Source", "Type", "Start", "End", "score", "strand", "phase", "attributes")
# Extract the target
df_repeat_raw$Target <- str_extract(df_repeat_raw$attributes, "(?<=Motif:)\\S+")
# Rename the chromosomes
chrom_map <- setNames(renamed_chromosomes, selected_chromosomes)
df_repeat_raw$Chromosome <- ifelse(df_repeat_raw$Chromosome %in% selected_chromosomes,
                          chrom_map[df_repeat_raw$Chromosome],
                          df_repeat_raw$Chromosome)
```

Took the selected chromosomes

```{r}
#| echo: true
#| warning: false

df_repeat_raw <- subset(df_repeat_raw, Chromosome %in% renamed_chromosomes)

```

In order to color those repeats like a previous paper. I need to read the .*out* file from RepeatMasker.

```{r}
#| echo: true
#| warning: false

# Read the RepeatMasker out file.
lines <- readLines(repeatmasker_out)
lines <- lines[-c(1:3)]
lines <- lines[lines != ""]
lines <- trimws(lines, which = "both")
raw_list <- strsplit(lines, "\\s+", perl = TRUE)
max_cols <- 15
raw_matrix <- t(sapply(raw_list, function(x){
  length(x) <- max_cols
  x
}))

raw_out <- as.data.frame(raw_matrix, stringsAsFactors = FALSE)


colnames(raw_out) <- c(
  "SW_score", "perc_div", "perc_del", "perc_ins",
  "query_sequence", "query_begin", "query_end", "query_left",
  "strand", "repeat_name", "repeat_class_family",
  "repeat_begin", "repeat_end", "repeat_left", "ID"
)

rep_info <- raw_out %>%
  select(repeat_name, repeat_class_family) %>%
  distinct()

#write.csv(rep_info,"repeats_info.csv",row.names = F)

#rep_info
```

I need to highlight some repeats

```{r}
#| echo: true
#| warning: false

# Dsim repeat library from the original paper
rep_info_new <- rep_info %>%
  mutate(
    repeat_class_family = case_when(
      
      # Specific satallites
      repeat_name == "500U_SAT" ~ "500-bp",
      repeat_name == "365U_SAT" ~ "365-bp",
      repeat_name == "RSP-LIKE_SAT" ~ "Rsp-Like",
      repeat_name == "135U_SAT" ~ "135-bp",
      
      # Jockey-3
      repeat_name == "Jockey-3_Dsim_dmel-like" ~ "Jockey-3",
      repeat_name == "Jockey-3_Dsim_anc-like" ~ "Jockey-3",
      
      # DNA
      repeat_class_family %in% c(
        "DNA/CMC-Transib", "DNA/P", "DNA/hAT-Ac",
        "DNA/hAT-hobo", "DNA/PiggyBac", "DNA/TcMar-Tc1"
      ) ~ "DNA transpon",
      
      # Non-LTR retrotransposon
      repeat_class_family %in% c(
        "Non-LTR_retrotransposon/CR1","Non-LTR_retrotransposon/I","Non-LTR_retrotransposon/Jockey","Non-LTR_retrotransposon/LOA","Non-LTR_retrotransposon/R1"
      ) ~ "Non-LTR retrotransposon",
      
      
      # LTR retrotransposons
      repeat_class_family %in% c(
        "LTR/Copia", "LTR/Gypsy", "LTR/Pao"
      ) ~ "LTR retrotransposon",
      
      # Satallite
      repeat_class_family %in% c(
        "Satellite/Rsp", "Satellite/Satellite"
      ) ~ "Satellite",
      
      

      # Simple & low complexity
      repeat_class_family == "Simple_repeat" ~ "Simple Repeat",
      
      
      # Others
      repeat_class_family %in% c(
        "Other/de_novo", "Other/IGS","RC/Helitron","Low_complexity"
      ) ~ "Others",


      TRUE ~ repeat_class_family  # 其他保持不变
    )
  )

# Dsim repeat library from the RepeatMasker Dfam database
# rep_info_new <- rep_info %>%
#   mutate(
#     repeat_class_family = case_when(
#       repeat_class_family == "DNA/CMC-Transib" ~ "DNA transpon",
#       repeat_class_family == "DNA/P" ~ "DNA transpon",
#       repeat_class_family == "DNA/hAT-Ac" ~ "DNA transpon",
#       repeat_class_family == "DNA/hAT-hobo" ~ "DNA transpon",
#       repeat_class_family == "DNA/PiggyBac" ~ "DNA transpon",
#       repeat_class_family == "DNA/TcMar-Tc1" ~ "DNA transpon",
#       repeat_class_family == "LINE/CR1" ~ "LINEs",
#       repeat_class_family == "LINE/I" ~ "LINEs",
#       repeat_class_family == "LINE/I-Jockey" ~ "LINE/I-Jockey",
#       repeat_class_family == "LINE/R1" ~ "LINEs",
#       repeat_class_family == "LINE/R1-LOA" ~ "LINEs",
#       repeat_class_family == "LINE/R2" ~ "LINE/R2",
#       repeat_class_family == "LTR/Copia" ~ "LTR retrotransposon",
#       repeat_class_family == "LTR/Gypsy" ~ "LTR retrotransposon",
#       repeat_class_family == "LTR/Pao" ~ "LTR retrotransposon",
# 
#       repeat_class_family == "RC/Helentron" ~ "Rolling circles ",
# 
#       repeat_class_family == "Simple_repeat" ~ "Simple Repeat",
#       repeat_class_family == "Low_complexity" ~ "Low complexity",
# 
# 
#       TRUE ~ repeat_class_family  # 其他保持不变
#     )
#   )

# rep_info_new <- rep_info %>%
#   mutate(
#     repeat_class_family = case_when(
#       repeat_name == "(TCTTC)n" ~ "AAGAG",
#       repeat_name == "(GAAGA)n" ~ "AAGAG",
#       repeat_name == "(AAGAT)n" ~ "AAGAT",
#       repeat_name == "(ATAAG)n" ~ "AAGAT",
#       repeat_name == "(GAGAA)n" ~ "AAGAG",
#       repeat_name == "(ATCTT)n" ~ "AATAG",
#       repeat_name == "(AATAG)n" ~ "AATAG",
#       repeat_name == "(ATTAT)n" ~ "AATAT",
#       repeat_name == "(TTATA)n" ~ "AATAT",
#       repeat_name == "(TATTA)n" ~ "AATAT",
#       repeat_name == "(TAGAATAACA)n" ~ "prodsat",
#       repeat_name == "DMSAT6" ~ "dodeca",
#       repeat_name == "Transib5" ~ "DNA transpon",
#       repeat_name == "ProtoP_A" ~ "DNA transpon",
#       TRUE ~ repeat_class_family  # 其他保持不变
#     )
#   )



```

Annotate the **df_repeat** table

```{r}
#| echo: true
#| warning: false

df_repeat <- df_repeat_raw %>%
  left_join(
    rep_info_new %>% select(repeat_name, repeat_class_family),
    by = c("Target" = "repeat_name")
  ) %>%
  rename(Class_family = repeat_class_family)

```

# Add repeats annotation in the plot.

```{r}
#| echo: true
#| warning: false
#| fig-align: center

#
my_family = "sans"


# -------------------------------
# 1. 染色体信息
# -------------------------------
chrom_info <- data_all %>%
  group_by(Chromosome) %>%
  summarise(Length = max(End), .groups="drop") %>%
  mutate(CumStart = cumsum(lag(Length, default=0)),
         Chromosome_label = paste0(Chromosome))

# -------------------------------
# 2. data_all 绘图准备
# -------------------------------
data_all_plot <- data_all %>%
  left_join(chrom_info, by="Chromosome") %>%
  mutate(
    Plot_Start = Start,
    Plot_End   = End,
    Plot_X = (Plot_Start + Plot_End)/2,             # geom_tile 需要中心 x
    Width = Plot_End - Plot_Start,                 # geom_tile 宽度
    Plot_Y = Counts_log2/2,                        # geom_tile 中心 y
    Height = Counts_log2,                          # geom_tile 高度
    Chromosome_label = factor(Chromosome_label, levels = chrom_info$Chromosome_label),
    Sample = factor(Sample, levels = unique(Sample)),
    Sample_group = case_when(
      Sample == "Ras3" ~ "D. mel",
      Sample == "ML82" ~ "D. sim",
      TRUE ~ "Hybrid"
    )
  )

# -------------------------------
# 3. df_repeat 绘图准备
# -------------------------------
#selected_targets <- c("G2_DM", "DOC2_DM")  # 只显示的Target
selected_targets <- rep_info_new$repeat_class_family
df_repeat_plot <- df_repeat %>%
  mutate(
    Counts = 2,
    Counts_log2 = 1,
    total = (End - Start) * Counts,
    Sample = "Repeats",
    Target_plot = ifelse(Class_family %in% selected_targets, Class_family, "Others")
  ) %>%
  left_join(chrom_info, by="Chromosome") %>%
  mutate(
    Plot_Start = Start,
    Plot_End   = End,
    Plot_X = (Plot_Start + Plot_End)/2,
    Width = Plot_End - Plot_Start,
    Plot_Y = Counts_log2/2,
    Height = Counts_log2,
    Chromosome_label = factor(Chromosome_label, levels = chrom_info$Chromosome_label)
  )

# -------------------------------
# 4. 设置颜色
# -------------------------------
sample_colors <- c(
  "D. mel" = "#fb8b05",
  "D. sim" = "#ed556a",
  "Hybrid" = "#61649F"
)

target_levels <- unique(df_repeat_plot$Target_plot)
target_colors <- setNames(rep("gray80", length(target_levels)), target_levels)
#target_colors[selected_targets] <- viridis(length(selected_targets))

target_colors <- c(
  "Satellite"               = "#CCB69E",  # 中灰
  "500-bp"                  = "#A7CDE2",  # 深蓝
  "135-bp"                  = "#1C79B4",  # 橙色
  "365-bp"                  = "#F0E442",  # 黄色
  "Simple Repeat"           = "#E4BBD9",  # 浅蓝
  "Others"                  = "#E0E0E0",  # 浅灰
  "Non-LTR retrotransposon" = "#F5A216",  # 粉紫
  "LTR retrotransposon"     = "#698B2D",  # 青绿色
  "Jockey-3"                = "#D24591",  # 橙黄
  "DNA transpon"            = "#E42220",  # 浅紫/粉
  "Rsp-Like"                = "#B05A27"   # 深蓝，可调亮度避免与 500-bp 混淆
)

# colorblindcheck::palette_check(target_colors, plot = TRUE)


# -------------------------------
# 5. 绘制 data_all 图 (geom_tile)
# -------------------------------
p1 <- ggplot(data_all_plot) +
  geom_tile(aes(
    x = Plot_X, width = Width,
    y = Plot_Y, height = Height,
    fill = Sample_group
  )) +
  facet_grid(Sample ~ Chromosome_label, scales = "free_x", space = "free_x", switch="both") +
  # 范围
  scale_y_continuous(
      breaks = c(0,15),                 # 主刻度
      minor_breaks = seq(0, 15, by = 5),    # 小刻度
      limits = c(0, 18),                    # 可选：固定范围（建议加）
      expand = c(0, 0.2),
      guide = guide_axis(minor.ticks = TRUE),
      labels = function(x) {
        ifelse(x == 0, "0", paste0("", x))
      }
    )+
  scale_fill_manual(values = sample_colors, name="Sample",drop=FALSE) +
  xlab(NULL) + 
  ylab(expression(Log[2](Counts + 1))) +
  theme_bw(base_family = "sans") +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(color = "black",fill = NA,linewidth = 0.1),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Facet labels
    strip.text.x = element_blank(),
    strip.text.y.left = element_text(
      angle = 0,
      hjust = 1,
      size = 6,
      family = my_family,
      color = "black",
      margin = margin(t = 0, r = 3, b = 0, l = 1)
    ),
    
    # Axis text
    axis.text.x = element_blank(),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_blank(),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Minor ticks
    axis.minor.ticks.x = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.y = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),

    
    # Compact legend
    legend.position = "right",
    legend.title = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 1, l = 0)
    ),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      face = "italic",
      margin = margin(t = 0, r = 0, b = 0, l = 0)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    # Facet spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    strip.placement = "outside",
    strip.switch.pad.grid = unit(0, "pt"),
    strip.clip = "off",
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  )

# -------------------------------
# 6. 绘制 df_repeat 图 (geom_tile)
# -------------------------------
p2 <- ggplot(df_repeat_plot) +
  geom_tile(aes(
    x = Plot_X, width = Width,
    y = Plot_Y, height = Height,
    fill = Target_plot
  )) +
  facet_grid(Sample ~ Chromosome_label, scales = "free_x", space = "free_x", switch="both") +
 scale_x_continuous(
  expand = c(0, 0),
  limits = c(0, NA),
  breaks = function(x) {
    seq(
      from = 0,
      to   = ceiling(max(x)/300e3) * 300e3,
      by   = 300e3
    )
  },
  minor_breaks = function(x) {
    seq(
      from = 0,
      to   = ceiling(max(x)/20e3) * 20e3,
      by   = 20e3
    )
  },
  labels = function(x) paste0(x/1e3, ""),
  guide = guide_axis(minor.ticks = TRUE)
)+ 
  scale_fill_manual(values = target_colors, name="Repeats") +
  xlab(expression(italic("D. sim")~"Centrometric Contigs (kb)")) + 
  ylab(NULL) +
  theme_bw(base_family = "sans") +
  theme(
      # Transparent backgrounds
      plot.background = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(fill = "transparent", color = NA),
      
      # Grid
      panel.grid = element_blank(),
      
      # Border
      panel.border = element_rect(
        color = "black",
        fill = NA,
        linewidth = 0.1
      ),
      
      # Avoid double-thick borders
      axis.line = element_blank(),
      
      # Facet labels
      strip.text.x = element_text(
        angle = 0,
        hjust = 0.5,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 1, r = 0, b = 1, l = 0)
      ),
      strip.text.y.left = element_text(
        angle = 0,
        hjust = 1,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 0, r = 1, b = 0, l = 1)
      ),
      
      # Axis text
      axis.text.x = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      axis.text.y = element_blank(),
      
      # Axis titles
      axis.title.x = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      axis.title.y = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      
      # Major ticks
      axis.ticks.x = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.y = element_blank(),
      axis.ticks.length = unit(1.5, "pt"),
      
      # Minor ticks
      axis.minor.ticks.x = element_line(
        linewidth = 0.1,
        color = "black"
      ),
      axis.minor.ticks.y = element_line(
        linewidth = 0.1,
        color = "black"
      ),
      axis.minor.ticks.length.x = unit(0.8, "pt"),
      axis.minor.ticks.length.y = unit(0.8, "pt"),
      
      # Compact legend
      legend.position = "right",
      legend.title = element_text(
        size = 6,
        color = "black",
        family = my_family,
        margin = margin(t = 0, r = 0, b = 1, l = 0)
      ),
      legend.text = element_text(
        size = 6,
        color = "black",
        family = my_family,
        margin = margin(t = 0, r = 0, b = 0, l = 2)
      ),
      legend.key.size = unit(6, "pt"),
      legend.key.width = unit(6, "pt"),
      legend.key.height = unit(6, "pt"),
      legend.key.spacing.x = unit(1, "pt"),
      legend.key.spacing.y = unit(1, "pt"),
      
      legend.spacing.x = unit(1, "pt"),
      legend.spacing.y = unit(1, "pt"),
      legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.spacing = unit(2, "pt"),
      
      # Facet spacing
      panel.spacing.x = unit(0.2, "lines"),
      panel.spacing.y = unit(0.4, "lines"),
      strip.placement = "outside",
      strip.switch.pad.grid = unit(0, "pt"),
      strip.clip = "off",
      
      # Plot margin
      plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
    )


# -------------------------------
# 7. 纵向合并
# -------------------------------
#p <- plot_grid(p1, p2, ncol=1, align="v", axis="lr", rel_heights=c(4,1))
#p


# -------------------------------
# 1. 提取 p1 的 legend（两列）
# -------------------------------
legend1 <- get_legend(
  p1 +
  guides(fill = guide_legend(ncol = 1))          # 两列显示
)

# -------------------------------
# 2. 提取 p2 的 legend（两列）
# -------------------------------
legend2 <- get_legend(
  p2 +
  guides(fill = guide_legend(ncol = 1))
)



# -------------------------------
# 3. 合并 p1 和 p2 图（不显示内部 legend）
# -------------------------------
p_combined <- plot_grid(
  p1 + theme(legend.position="none"),
  p2 + theme(legend.position="none"),
  ncol = 1, align = "v", rel_heights = c(2.5, 1)
)


# -------------------------------
# 4. 纵向排列两个 legend，并控制高度比例
# -------------------------------
legend_combined <- plot_grid(
  legend1,
  legend2,
  ncol = 1,
  rel_heights = c(1.4,6),
  align = "v"
)


# -------------------------------
# 5. 最终合并图和 legend
# -------------------------------
final_plot <- plot_grid(
  p_combined,
  legend_combined,
  ncol = 2,
  rel_widths = c(1, 0.22),
  align = "h",       # 横向对齐
  axis = "l"         # 对齐顶部
)


```

View and save

```{r}
#| echo: true
#| warning: false
#| fig-align: center


final_plot_canvas <- final_plot + ggview::canvas(175,50,units="mm",dpi = 900,bg="transparent")
final_plot_canvas

#ggsave("fig2C.pdf",final_plot,width=175, height=50,units="mm",dpi = 900,bg="transparent")

```
