---
title: "FigS2. Assembly of Ras3 genome"
author: "Jie Hua"
date: 2026-02-01
lang: en
date-modified: 2026-04-08
keywords: ["IF","Probe","Hybrid Cells","DNA-FISH"]
code-copy: true
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: true
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Contig Length

## Load the required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)
```

## Load the input information

```{r}
#| echo: true
#| warning: false

# Read the genome files
# Lab PC
folder <- "H:\\HybridCells_Data\\FigS1_Assembly of Ras3 genome"
# Home PC
#folder <- "F:\\Hybrid Cells Paper Processed Paper\\FigS1_Assembly of Ras3 genome"

# Read the length of all chromosomes in Ras3 and Dmel6
ras3_fai_file <- paste0(folder,"/ras3.fasta.fai")
dmel6_fai_file <- paste0(folder,"/dmel-all-chromosome-r6.56.fasta.fai")
```

Define a function to read the *.fai* file

```{r}
#| echo: true
#| warning: false

# Define a small function to read the fai file
read_fai <- function(fai_file){
  fai <- read.table(fai_file)
  fai <- fai[,1:2]
  colnames(fai) <- c("Chromosome","Length")
  return(fai)
}
```

Read the .fai files

```{r}
#| echo: true
#| warning: false

# Read the length information from fai files
ras3_length <- read_fai(ras3_fai_file)
ras3_length$Genome <- "ras3"
dmel6_length <- read_fai(dmel6_fai_file)
dmel6_length$Genome <- "dm6.56"

ras3_length <- subset(ras3_length, Chromosome %in% c("chrX_mel","chr2L_mel","chr2R_mel","chr3L_mel","chr3R_mel","chr4_mel"))
dmel6_length <- subset(dmel6_length, Chromosome %in% c("X","2L","2R","3L","3R","4"))
lengths <- rbind(ras3_length, dmel6_length)

# Merge the data into one table
ras3_length <- ras3_length %>%
  mutate(Chromosome = gsub("chr|_mel", "", Chromosome))

merged <- bind_rows(dmel6_length, ras3_length)

# Order the chromosomes
order <- c("X", "2L", "2R", "3L", "3R", "4")
merged$Chromosome <- factor(merged$Chromosome, levels = order)
```

## Plot

```{r}
#| echo: true
#| warning: false

# Define custom colors
custome_color <- c("#61649F","#ED556A")

p1 <- ggplot(merged, aes(x = Chromosome, y = Length / 1e6, fill = Genome)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.8),
    width = 0.7,
    alpha = 0.8,
    colour = "grey20",
    linewidth = 0.2
  ) +
  scale_fill_manual(values = custome_color) +
  labs(
    x = "Chromosome",
    y = "Length (Mb)"
  ) +
  scale_y_continuous(
    expand = c(0.05, 0),
    breaks = seq(0, 30, by = 10),
    minor_breaks = seq(0, 33, by = 2),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  theme_bw() +
  theme(
    # 背景透明
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),

    # 网格去掉
    panel.grid = element_blank(),

    # 边框
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.4),
    # 坐标线
    axis.line = element_blank(),

    # 坐标文字
    axis.text.x = element_text(size = 6, color = "black"),
    axis.text.y = element_text(size = 6, color = "black"),
    axis.title.x = element_text(size = 7, color = "black"),
    axis.title.y = element_text(size = 7, color = "black"),

    # 主刻度
    axis.ticks.x = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.y = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.length = unit(1.5, "pt"),

    # 小刻度
    axis.minor.ticks.x = element_line(linewidth = 0.1, color = "black"),
    axis.minor.ticks.y = element_line(linewidth = 0.1, color = "black"),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),

    # 图例
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 6, color = "black"),
    
    # 👇 核心：压缩 legend 和图之间的距离
    legend.margin = margin(t = 0, r = 0, b = -4, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = -1, l = 0),

    # 👇 再微调 item 间距
    legend.spacing.y = unit(0, "pt"),
    
    # 图例色块缩小
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),

    # 边距
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  ) +
  guides(fill = guide_legend(keywidth = unit(6, "pt"),
        keyheight = unit(6, "pt"),nrow = 1, byrow = TRUE))
```

Save the plot

```{r}
#| echo: true
#| warning: false
p1_canvas <- p1 + ggview::canvas(70,50,units="mm",dpi = 900,bg="transparent")
p1_canvas
#ggsave("figS1A.pdf", plot = p1, width = 70, height = 50, units="mm",dpi = 900,bg="transparent")
```

# BUSCO assessment

The raw script was generated from BUSCO. I just modified the styles.

## Load required packages

```{r}
#| echo: true
#| warning: false
library(ggplot2)
library(grid)
```

## Plot

```{r}
#| echo: true
#| warning: false

# Colors
my_colors <- c("#56B4E9", "#3492C7", "#F0E442", "#F04442")
# Bar height ratio
my_bar_height <- 0.75

# Legend
my_title <- "BUSCO Assessment"

# Font
my_family <- "sans"
my_size_ratio <- 1

# !!! SEE YOUR DATA HERE !!! 
# Your data as generated by python, remove or add more
my_species <- c('ras3', 'ras3', 'ras3', 'ras3', 'dm6.56', 'dm6.56', 'dm6.56', 'dm6.56')
my_species <- factor(my_species)
my_species <- factor(my_species,levels(my_species)[c(length(levels(my_species)):1)]) # reorder your species here just by changing the values in the vector :
my_percentage <- c(99.2, 0.4, 0.4, -0.0, 98.8, 0.8, 0.4, 0.0)
my_values <- c(253, 1, 1, 0, 252, 2, 1, 0)

######################################
######################################
######################################
# Code to produce the graph
labsize = 1
if (length(levels(my_species)) > 10){
 labsize = 0.66
}

category <- c(rep(c("S","D","F","M"),c(1)))
category <-factor(category)
category = factor(category,levels(category)[c(4,1,2,3)])
df = data.frame(my_species,my_percentage,my_values,category)

p2 <- ggplot() +
  geom_bar(
    aes(y = my_percentage, x = my_species, fill = category),
    data = df,
    stat = "identity",
    width = my_bar_height,
    position = position_stack(reverse = TRUE),
    colour = "grey20",
    linewidth = 0.2,
    alpha=0.9
  ) +
  coord_flip() +
  scale_y_continuous(
    breaks = c(0, 20, 40, 60, 80, 100),
    labels = c("0", "20", "40", "60", "80", "100"),
    minor_breaks = seq(0, 100, by = 5),
    guide = guide_axis(minor.ticks = TRUE)
  )+
  scale_fill_manual(
    values = my_colors,
    labels = c(
      "Complete (C) and single-copy (S)",
      "Complete (C) and duplicated (D)",
      "Fragmented (F)",
      "Missing (M)"
    )
  ) +
  xlab("") +
  ylab("\n%BUSCOs") +
  guides(
    fill = guide_legend(
      nrow = 2,
      byrow = TRUE,
      override.aes = list(colour = NA)
    )
  ) +
  theme_bw() +
  theme(
    # 保持你原来的 legend 布局逻辑
    legend.position = "top",
    legend.box = "horizontal",
    legend.justification = "right",
    legend.box.just = "left",
    legend.title = element_blank(),
    legend.text = element_text(
      family = my_family,
      size = 5,
      colour = "black"
    ),
    # 图例色块缩小
    legend.key.width = unit(5.5, "pt"),
    legend.key.height = unit(5.5, "pt"),
    
    # 再微调 item 间距
    legend.spacing.y = unit(0, "pt"),
    # 压缩 legend 和图之间的距离
    legend.margin = margin(t = 0, r = 0, b = -4, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = -1, l = 0),

    # 透明背景
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA),

    # 去网格
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),

    # 边框
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.4),

    # 坐标文字
    axis.text.y = element_text(
      family = my_family,
      colour = "black",
      size = 6,
      margin = margin(r = 2)
    ),
    axis.text.x = element_text(
      family = my_family,
      colour = "black",
      size = 6
    ),
    axis.title.x = element_text(
      family = my_family,
      size = 7,
      colour = "black"
    ),
    axis.title.y = element_blank(),

    # 刻度线
    axis.line = element_blank(),
    axis.ticks.x = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.y = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.length = unit(1.5, "pt"),

    axis.minor.ticks.x = element_line(linewidth = 0.1, color = "black"),
    axis.minor.ticks.length.x = unit(0.8, "pt"),

    # 边距
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  )+
  guides(fill = guide_legend(keywidth = unit(5.5, "pt"),
        keyheight = unit(5.5, "pt"),nrow = 2, byrow = TRUE))

for (i in rev(seq_along(levels(my_species)))) {
  detailed_values <- my_values[my_species == levels(my_species)[i]]
  total_buscos <- sum(detailed_values)

  p2 <- p2 +
  annotate(
    "text",
    x = i,
    y = 3,
    hjust = 0,
    size = 7 / 2.845,   # 👈 真正 6 pt
    colour = "black",
    family = my_family,
    label = paste(
      "C:", detailed_values[1] + detailed_values[2],
      " [S:", detailed_values[1],
      ", D:", detailed_values[2],
      "], F:", detailed_values[3],
      ", M:", detailed_values[4],
      ", n:", total_buscos,
      sep = ""
    )
  )
}
```

Save the plot

```{r}
#| echo: true
#| warning: false

p2_canvas <- p2 + ggview::canvas(70,50,units="mm",dpi = 1200,bg="transparent")
p2_canvas
#ggsave("figS1B.pdf", plot = p2, width = 70, height = 50,units="mm", dpi = 900,bg="transparent")
```

# Circulation plot

## Load required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(circlize)
library(dplyr)
library(seqinr)
library(zoo)
library(GenomicRanges)
library(ggbio)
library(grid)
library(ggplot2)
```

## Read the basic information

```{r}
#| echo: true
#| warning: false

# Define the folder of processed data
folder <- "H:\\HybridCells_Data\\FigS1_Assembly of Ras3 genome"

#folder <- "F:\\Hybrid Cells Paper Processed Paper\\FigS1_Assembly of Ras3 genome"

# Read the chromosome length
fai <- read.table(paste0(folder, "/ras3.fasta.fai"), sep = "\t", header = FALSE)
colnames(fai) <- c("chr", "length", "offset", "linebases", "linewidth")

# Target chromosomes
target_chr <- c("chrX_mel","chr2L_mel","chr2R_mel","chr3L_mel","chr3R_mel","chr4_mel")
fai <- fai[fai$chr %in% target_chr, ]

# Read gene annotation
gene_gtf <- read.table(paste0(folder, "/Ras3_liftoff_20250522_polished_transformed_cleaned.gtf"), sep = "\t", header = FALSE, comment.char = "#", quote = "")
colnames(gene_gtf) <- c("chr", "source", "type", "start", "end", "score", "strand", "phase", "attr")
genes <- gene_gtf %>%
  filter(type == "transcript", chr %in% target_chr) %>%
  select(chr, start, end)

# Read repeat annotation
repeat_gff <- read.table(paste0(folder, "/ras3.fasta.repeatmasker.out.gff"), sep = "\t", header = FALSE, comment.char = "#", quote = "")
colnames(repeat_gff) <- c("chr", "source", "type", "start", "end", "score", "strand", "phase", "attr")
repeats <- repeat_gff %>%
  filter(chr %in% target_chr) %>%
  select(chr, start, end)

```

## Calculate some tables

```{r}
#| echo: true
#| warning: false

# Compute GC content
fasta_file <- paste0(folder, "/ras3.fasta")
seqs <- read.fasta(fasta_file, seqtype = "DNA", as.string = FALSE, forceDNAtolower = FALSE)


# Use 100 kb window to calculate the GC
window_size <- 100000
gc_data <- data.frame()

for (chr in target_chr) {
  cat("Calculating GC for", chr, "...\n")
  if (!chr %in% names(seqs)) next
  seq <- toupper(unlist(seqs[[chr]]))
  seq_length <- length(seq)
  starts <- seq(1, seq_length, by = window_size)
  ends <- pmin(starts + window_size - 1, seq_length)
  gc_values <- sapply(1:length(starts), function(i) {
    s <- seq[starts[i]:ends[i]]
    gc_count <- sum(s %in% c("G", "C"))
    gc_count / length(s)
  })
  
  gc_data <- rbind(gc_data, data.frame(chr = chr, start = starts, end = ends, gc = gc_values))
}



# Preprocessing
# Remove 'chr' prefix and '_mel' postfix
fai$chr <- gsub("^chr|_mel$", "", fai$chr)
genes$chr <- gsub("^chr|_mel$", "", genes$chr)
repeats$chr <- gsub("^chr|_mel$", "", repeats$chr)
gc_data$chr <- gsub("^chr|_mel$", "", gc_data$chr)

# Define a function to calculate the counts of genes or repeats
get_counts <- function(df, chr_name, chr_length, window_size = 1e5) {
  starts <- seq(1, chr_length, by = window_size)
  ends <- pmin(starts + window_size - 1, chr_length)
  
  counts <- sapply(1:length(starts), function(i) {
    sub_df <- df[df$chr == chr_name, ]
    sum(sub_df$start <= ends[i] & sub_df$end >= starts[i])
  })
  
  data.frame(chr = chr_name,
             start = starts,
             end = ends,
             count = counts)
}

# Calculate gene density
window_size <- 1e5
gene_density <- do.call(rbind, lapply(fai$chr, function(ch) {
  get_counts(genes, ch, fai$length[fai$chr == ch], window_size)
}))

# Calculate repeat density
repeat_density <- do.call(rbind, lapply(fai$chr, function(ch) {
  get_counts(repeats, ch, fai$length[fai$chr == ch], window_size)
}))

# Calculate/Copy the GC content
gc_density <- gc_data
```

## Prepare GRanges objects

```{r}
#| echo: true
#| warning: false

# ideogram (染色体长度)
gr_chr <- GRanges(
  seqnames = fai$chr,
  ranges = IRanges(start = 1, end = fai$length)
)

# gene density track
gr_gene <- GRanges(
  seqnames = gene_density$chr,
  ranges = IRanges(start = gene_density$start, end = gene_density$end),
  gene_count = gene_density$count
)

# repeat density track
gr_repeat <- GRanges(
  seqnames = repeat_density$chr,
  ranges = IRanges(start = repeat_density$start, end = repeat_density$end),
  repeat_count = repeat_density$count
)

# GC content track
gr_gc <- GRanges(
  seqnames = gc_density$chr,
  ranges = IRanges(start = gc_density$start, end = gc_density$end),
  gc_score = gc_density$gc
)

# ------------------------------
# 2. 缩放 track 以便环形显示
# ------------------------------

amp <- 1.2  # 放大系数，可调节 Y 轴起伏
mcols(gr_gene)$gene_count <- mcols(gr_gene)$gene_count * amp
mcols(gr_repeat)$repeat_count <- mcols(gr_repeat)$repeat_count * amp
mcols(gr_gc)$gc_score <- mcols(gr_gc)$gc_score * amp
```

## Plot A Function: ggbio

```{r}
#| echo: true
#| warning: false

# gene/repeat/gc 的颜色映射
track_colors <- c("Gene Density" = "#F5A216",
                  "Repeat Density" = "#CCB69E",
                  "GC Content" = "#A7CDE2")

# Check the colors
# colorblindcheck::palette_check(track_colors, plot = TRUE)


p3 <- ggbio() + 
  # 染色体 track
  circle(gr_chr, geom = "ideo", fill = "gray90", color = "black", radius = 20, trackWidth = 4) +
  
  # gene density track
  circle(gr_gene, geom = "line", aes(y = gene_count, color = "Gene Density"),
         radius = 15, trackWidth = 4.5, linewidth = 0.5) +
  
  # repeat density track
  circle(gr_repeat, geom = "line", aes(y = repeat_count, color = "Repeat Density"),
         radius = 10, trackWidth = 4.5, linewidth = 0.5) +
  
  # GC content track
  circle(gr_gc, geom = "line", aes(y = gc_score, color = "GC Content"),
         radius = 5, trackWidth = 4.5, linewidth = 0.5) +
  
  # 设置颜色对应图例
  scale_color_manual(name = "Tracks", values = track_colors) +
  
  # 关键：让图例里的线变粗
  guides(
    color = guide_legend(
      override.aes = list(linewidth = 2),   # 这里改数字，越大的线越粗，建议 1.5~3
      keywidth = unit(2, "cm"),
    )
  ) +
  
  # 外圈刻度
  circle(gr_chr, geom = "scale", radius = 24, trackWidth = 1, 
         major.by = 5e6, minor.by = 1e6, label.cex = 0.8, family="sans") +
  
  # 染色体名字
  circle(gr_chr, geom = "text", aes(label = seqnames), 
         radius = 21.5, trackWidth = 1, size = 6, family="sans") +
  
  # legend 设置在左上角，更靠边，线条加粗，所有文字 sans
  theme(
    legend.position = c(-0.2, 1),
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size=20, family="sans"),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.box.background = element_rect(fill = "transparent", colour = NA),

  )

```

Save the plot

```{r}
#| echo: true
#| warning: false

save_circos_all <- function(p, prefix = "Ras3_circos_plot", width = 7, height = 7) {
  # 1. PDF（矢量，透明背景）
  pdf(
    paste0(prefix, ".pdf"),
    width = width,
    height = height + 1,
    family = "sans",
    useDingbats = FALSE,
    bg = "transparent"
  )
  print(p)
  dev.off()
  
  # 2. SVG（最推荐用于AI / Affinity）
  svg(
    paste0(prefix, ".svg"),
    width = width + 1,
    height = height + 1,
    family = "sans",
    bg = "transparent"
  )
  print(p)
  dev.off()
  
  # 3. PNG（1200 dpi，透明背景）
  png(
    paste0(prefix, "_1200dpi.png"),
    width = width,
    height = height,
    units = "in",
    res = 1200,
    family = "sans",
    bg = "transparent"
  )
  print(p)
  dev.off()
  
  cat("全部保存完毕（透明背景）！\n")
  cat("   PDF :", paste0(prefix, ".pdf"), "\n")
  cat("   SVG :", paste0(prefix, ".svg"), "\n")
  cat("   PNG :", paste0(prefix, "_1200dpi.png"), "\n")
}
```

## Plot B Function: ggplot2 version

```{r}
#| echo: true
#| warning: false

# =========================================================
# 0. 颜色
# =========================================================
track_colors <- c(
  "Gene Density"   = "#ed556a",
  "Repeat Density" = "#2177b8",
  "GC Content"     = "#fb8b05"
)

# =========================================================
# 1. GRanges -> data.frame
# =========================================================
df_chr <- as.data.frame(gr_chr) %>%
  dplyr::select(seqnames, start, end) %>%
  dplyr::rename(chr = seqnames) %>%
  dplyr::mutate(chr = as.character(chr))

df_gene <- as.data.frame(gr_gene) %>%
  dplyr::select(seqnames, start, end, gene_count) %>%
  dplyr::rename(chr = seqnames) %>%
  dplyr::mutate(chr = as.character(chr))

df_repeat <- as.data.frame(gr_repeat) %>%
  dplyr::select(seqnames, start, end, repeat_count) %>%
  dplyr::rename(chr = seqnames) %>%
  dplyr::mutate(chr = as.character(chr))

df_gc <- as.data.frame(gr_gc) %>%
  dplyr::select(seqnames, start, end, gc_score) %>%
  dplyr::rename(chr = seqnames) %>%
  dplyr::mutate(chr = as.character(chr))

# =========================================================
# 2. 染色体顺序、长度、gap
#    每个染色体后面都保留 gap，包括最后一个
# =========================================================
chr_info <- df_chr %>%
  dplyr::group_by(chr) %>%
  dplyr::summarise(chr_len = max(end), .groups = "drop")

chr_levels <- unique(df_chr$chr)
chr_info$chr <- factor(chr_info$chr, levels = chr_levels)

gap_size <- 1.5e6

chr_info <- chr_info %>%
  dplyr::arrange(chr) %>%
  dplyr::mutate(
    offset = cumsum(dplyr::lag(chr_len + gap_size, default = 0)),
    genome_start = offset,
    genome_end   = offset + chr_len,
    genome_mid   = (genome_start + genome_end) / 2
  )

total_length <- sum(chr_info$chr_len) + nrow(chr_info) * gap_size

# =========================================================
# 3. 给 track 加 cumulative genome 坐标
# =========================================================
add_genome_pos <- function(df, value_col) {
  df %>%
    dplyr::left_join(chr_info %>% dplyr::select(chr, offset), by = "chr") %>%
    dplyr::mutate(
      gstart = offset + start,
      gend   = offset + end,
      gmid   = (gstart + gend) / 2,
      value  = .data[[value_col]]
    ) %>%
    dplyr::arrange(chr, gmid)
}

df_gene2   <- add_genome_pos(df_gene, "gene_count")
df_repeat2 <- add_genome_pos(df_repeat, "repeat_count")
df_gc2     <- add_genome_pos(df_gc, "gc_score")

# =========================================================
# 4. genome position -> phi
#    12点钟起点，顺时针
# =========================================================
to_phi <- function(pos, total_length) {
  2 * pi * (pos / total_length)
}

phi_to_xy <- function(phi, r) {
  theta <- pi/2 - phi
  data.frame(
    x = r * cos(theta),
    y = r * sin(theta)
  )
}

chr_info <- chr_info %>%
  dplyr::mutate(
    phi_start = to_phi(genome_start, total_length),
    phi_end   = to_phi(genome_end, total_length),
    phi_mid   = to_phi(genome_mid, total_length)
  )

df_gene2   <- df_gene2   %>% dplyr::mutate(phi = to_phi(gmid, total_length))
df_repeat2 <- df_repeat2 %>% dplyr::mutate(phi = to_phi(gmid, total_length))
df_gc2     <- df_gc2     %>% dplyr::mutate(phi = to_phi(gmid, total_length))

# =========================================================
# 5. 半径参数
# =========================================================
r_ideo_in  <- 0.86
r_ideo_out <- 1.00

r_gene_base <- 0.69
r_rep_base  <- 0.52
r_gc_base   <- 0.35

# 起伏更明显
track_amp <- 0.15

# =========================================================
# 6. 数值缩放到各自 track
# =========================================================
rescale01 <- function(x) {
  xr <- range(x, na.rm = TRUE)
  if (diff(xr) == 0) return(rep(0.5, length(x)))
  (x - xr[1]) / diff(xr)
}

df_gene2 <- df_gene2 %>%
  dplyr::mutate(r = r_gene_base + rescale01(value) * track_amp)

df_repeat2 <- df_repeat2 %>%
  dplyr::mutate(r = r_rep_base + rescale01(value) * track_amp)

df_gc2 <- df_gc2 %>%
  dplyr::mutate(r = r_gc_base + rescale01(value) * track_amp)

gene_xy <- dplyr::bind_cols(df_gene2, phi_to_xy(df_gene2$phi, df_gene2$r))
rep_xy  <- dplyr::bind_cols(df_repeat2, phi_to_xy(df_repeat2$phi, df_repeat2$r))
gc_xy   <- dplyr::bind_cols(df_gc2, phi_to_xy(df_gc2$phi, df_gc2$r))

# =========================================================
# 7. 外圈 ideogram polygon
# =========================================================
make_arc_polygon_phi <- function(phi1, phi2, r_in, r_out, n = 300) {
  ph_outer <- seq(phi1, phi2, length.out = n)
  ph_inner <- seq(phi2, phi1, length.out = n)

  out_xy <- phi_to_xy(ph_outer, r_out)
  in_xy  <- phi_to_xy(ph_inner, r_in)

  data.frame(
    x = c(out_xy$x, in_xy$x),
    y = c(out_xy$y, in_xy$y)
  )
}

ideo_df <- dplyr::bind_rows(lapply(seq_len(nrow(chr_info)), function(i) {
  poly <- make_arc_polygon_phi(
    phi1 = chr_info$phi_start[i],
    phi2 = chr_info$phi_end[i],
    r_in = r_ideo_in,
    r_out = r_ideo_out,
    n = 300
  )
  poly$id  <- i
  poly$chr <- as.character(chr_info$chr[i])
  poly
}))

# =========================================================
# 8. 染色体标签
#    右转 90°
# =========================================================
chr_label_df <- chr_info %>%
  dplyr::mutate(r = 0.93) %>%
  dplyr::bind_cols(phi_to_xy(.$phi_mid, .$r)) %>%
  dplyr::mutate(
    angle_raw = 90 - phi_mid * 180 / pi,
    angle = ifelse(
      angle_raw < -90, angle_raw + 180,
      ifelse(angle_raw > 90, angle_raw - 180, angle_raw)
    ),
    angle = angle + 90
  )

# =========================================================
# 9. 外圈刻度
#    刻度文字往里一些，字体更小
# =========================================================
make_ticks_one_chr <- function(chr, chr_len, offset, major_by = 5e6, minor_by = 1e6) {
  major_pos <- seq(0, chr_len, by = major_by)
  minor_pos <- seq(0, chr_len, by = minor_by)

  major_df <- data.frame(
    chr = chr,
    pos_chr = major_pos,
    pos_genome = offset + major_pos,
    type = "major",
    label = paste0(major_pos / 1e6, " M")
  )

  minor_df <- data.frame(
    chr = chr,
    pos_chr = minor_pos,
    pos_genome = offset + minor_pos,
    type = "minor",
    label = NA_character_
  )

  list(major = major_df, minor = minor_df)
}

tick_list <- lapply(seq_len(nrow(chr_info)), function(i) {
  make_ticks_one_chr(
    chr = as.character(chr_info$chr[i]),
    chr_len = chr_info$chr_len[i],
    offset = chr_info$offset[i],
    major_by = 5e6,
    minor_by = 1e6
  )
})

major_tick_df <- dplyr::bind_rows(lapply(tick_list, `[[`, "major")) %>%
  dplyr::mutate(phi = to_phi(pos_genome, total_length))

minor_tick_df <- dplyr::bind_rows(lapply(tick_list, `[[`, "minor")) %>%
  dplyr::mutate(phi = to_phi(pos_genome, total_length))

minor_tick_seg <- minor_tick_df %>%
  dplyr::mutate(
    x    = 1.005 * cos(pi/2 - phi),
    y    = 1.005 * sin(pi/2 - phi),
    xend = 1.015 * cos(pi/2 - phi),
    yend = 1.015 * sin(pi/2 - phi)
  )

major_tick_seg <- major_tick_df %>%
  dplyr::mutate(
    x    = 1.005 * cos(pi/2 - phi),
    y    = 1.005 * sin(pi/2 - phi),
    xend = 1.025 * cos(pi/2 - phi),
    yend = 1.025 * sin(pi/2 - phi)
  )

major_tick_text <- major_tick_df %>%
  dplyr::mutate(
    r = 1.12,
    x = r * cos(pi/2 - phi),
    y = r * sin(pi/2 - phi),
    angle_raw = 90 - phi * 180 / pi,
    angle = ifelse(
      angle_raw < -90, angle_raw + 180,
      ifelse(angle_raw > 90, angle_raw - 180, angle_raw)
    )
  )

# =========================================================
# 10. 作图
# =========================================================
p_circos <- ggplot() +
  # 外圈 ideogram：透明背景
  geom_polygon(
    data = ideo_df,
    aes(x = x, y = y, group = id),
    fill = NA,
    color = "black",
    linewidth = 0.2
  ) +

  # gene track：细一点
  geom_path(
    data = gene_xy,
    aes(x = x, y = y, group = chr, color = "Gene Density"),
    linewidth = 0.2,
    lineend = "round"
  ) +

  # repeat track：细一点
  geom_path(
    data = rep_xy,
    aes(x = x, y = y, group = chr, color = "Repeat Density"),
    linewidth = 0.2,
    lineend = "round"
  ) +

  # gc track：细一点
  geom_path(
    data = gc_xy,
    aes(x = x, y = y, group = chr, color = "GC Content"),
    linewidth = 0.2,
    lineend = "round"
  ) +

  # minor ticks
  geom_segment(
    data = minor_tick_seg,
    aes(x = x, y = y, xend = xend, yend = yend),
    inherit.aes = FALSE,
    linewidth = 0.1,
    color = "black"
  ) +

  # major ticks
  geom_segment(
    data = major_tick_seg,
    aes(x = x, y = y, xend = xend, yend = yend),
    inherit.aes = FALSE,
    linewidth = 0.2,
    color = "black"
  ) +

  # major tick labels：更小、稍微往里
  geom_text(
    data = major_tick_text,
    aes(x = x, y = y, label = label, angle = angle),
    family = "sans",
    size = 6/2.835,
    color = "black"
  ) +

  # chromosome labels
  geom_text(
    data = chr_label_df,
    aes(x = x, y = y, label = chr, angle = angle),
    family = "sans",
    size = 6/2.835,
    color = "black"
  ) +

  scale_color_manual(values = track_colors) +

  coord_fixed(clip = "off") +

  theme_bw() +
  theme(
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),

    panel.grid = element_blank(),
    panel.border = element_blank(),

    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),

    # 图例更靠左上，行间稍微拉开
    legend.position = c(-0.01, 0.99),
    legend.justification = c("left", "top"),
    legend.title = element_blank(),
    legend.text = element_text(size = 6, color = "black", family = "sans"),

    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.spacing.y = unit(1, "pt"),

    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key = element_rect(fill = "transparent", colour = NA),

    plot.margin = margin(t = 1, r = 1, b = 1, l = 1)
  ) +
  guides(
    color = guide_legend(
      keywidth = unit(8, "pt"),
      keyheight = unit(6, "pt"),
      nrow = 3,
      byrow = TRUE,
      override.aes = list(linewidth = 0.6)
    )
  ) +
  xlim(-1.22, 1.22) +
  ylim(-1.22, 1.22)
```

Save the plot

```{r}
#| echo: true
#| warning: false

p_circos_canvas <- p_circos + ggview::canvas(70,70,units="mm",dpi = 1200,bg="transparent")
p_circos_canvas
#ggplot2::ggsave("figS1C.pdf", plot = p_circos, width = 70, height = 70,units="mm", dpi = 900,bg="transparent")

```

# Hi-C contact

Check some useful tutorials:

<https://jserizay.com/HiContacts/articles/HiContacts.html>

<https://jserizay.com/OHCA/docs/devel/>

## Load required packages

```{r}
#| echo: true
#| warning: false
#| message: false

# Load required packages
library(HiCExperiment)
library(HiCool)
library(HiContacts)
library(HiContactsData)
library(fourDNData)
library(DNAZooData)
library(GenomicRanges)
library(ggplot2)
library(grid)
```

## Read the basic files

```{r}
#| echo: true
#| warning: false


# Define the folder containing the files
folder <- "H:\\HybridCells_Data\\FigS1_Assembly of Ras3 genome"
filename <- "ras3_hic_matrix_1000bp.mcool"

# Check available resolutions
availableResolutions(paste0(folder,"/",filename))

# Read the mcool file
data <- HiCExperiment::import(paste0(folder,"/",filename),resolution=128000, format="mcool")

# Check some basic information of data
scores(data)
fileName(data)
resolutions(data)
topologicalFeatures(data)
pairsFile(data)
metadata(data)

# Check the seqnames of data
# seqlevels(data)

# Prepare the metadata
metadata(data) <- list(
  info = "Ras3 Hi-C data from Yuri Lab mapped to Ras3 genome" 
)
metadata(data)


# Zoom to a larger resolution and take specific chromosomes
data_selected <- zoom(data, 128000)
data_selected <- data_selected[c("chrX", "chr2L", "chr2R", "chr3L", "chr3R", "chr4")]
```

## Plot

```{r}
#| echo: true
#| warning: false


# Build the length of chromosomes
chromosome_data <- data.frame(
  Chromosome = c("X", "2L", "2R", "3L", "3R", "4"),
  Length = c(24612742, 23650026, 25463053, 28085216, 32896612, 1180059)
)

# Calculate the cumulative start site and end site
chromosome_data$Start <- c(0, cumsum(head(chromosome_data$Length, -1)))
chromosome_data$End <- chromosome_data$Start + chromosome_data$Length
chromosome_data$Midpoints_x <- (chromosome_data$Start + chromosome_data$End) / 2
chromosome_data$Midpoints_y <- chromosome_data$Midpoints_x * -1

# Use default function to generate a ggplot2 object
hic_plot <- plotMatrix(data_selected, use.scores = 'balanced', scale="log2",chrom_lines = TRUE,cmap = afmhotrColors(), caption=FALSE)

# Add labels and custom styles
hic_plot <- plotMatrix(
  data_selected,
  use.scores = 'balanced',
  scale = "log2",
  chrom_lines = TRUE,
  cmap = afmhotrColors(),
  caption = FALSE
)

hic_plot_final <- hic_plot +

  # -----------------------------
  # X 轴（顶部）
  # -----------------------------
  scale_x_continuous(
    breaks = chromosome_data$Midpoints_x,
    labels = chromosome_data$Chromosome,
    limits = c(0, 136065679),
    expand = c(0, 0),
    position = "top"
  ) +

  # -----------------------------
  # Y 轴（反转）
  # -----------------------------
  scale_y_reverse(
    breaks = chromosome_data$Midpoints_x,
    labels = chromosome_data$Chromosome,
    limits = c(136065679, 0),
    expand = c(0, 0)
  ) +

  # -----------------------------
  # 主题（只改外观，不碰 legend scale）
  # -----------------------------
  theme_bw(base_family = "sans") +
  labs(fill = "Score")+
  theme(
    # ===== 背景透明 =====
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),

    # ===== 去网格 =====
    panel.grid = element_blank(),

    # ===== 边框 =====
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.4),

    # ===== 坐标轴 =====
    axis.title = element_blank(),

    axis.text.x = element_text(size = 6, color = "black", family = "sans"),
    axis.text.y = element_text(size = 6, color = "black", family = "sans"),

    axis.ticks.x = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.y = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.length = unit(1.5, "pt"),

    # Hi-C 禁用 minor ticks（否则会乱）
    axis.minor.ticks.x = element_blank(),
    axis.minor.ticks.y = element_blank(),

    # ===== legend（只改外观，不改类型）=====
    legend.position = "right",

    legend.title = element_text(
      size = 6,
      color = "black",
      family = "sans"
    ),

    legend.text = element_text(
      size = 5,
      color = "black",
      family = "sans"
    ),

    # 压缩距离
    legend.margin = margin(t = 0, r = 0, b = -4, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = -2, l = 0),

    # 不改变 colorbar 结构，只缩紧 spacing
    legend.spacing.y = unit(0, "pt"),

    # 关键：控制 colorbar 厚度（不是 key）
    legend.key.width = unit(20, "pt"),
    legend.key.height = unit(6, "pt"),

    plot.margin = margin(t = 4, r = 2, b = 2, l = 2)
  )+
  guides(
  fill = guide_colorbar(
    barwidth = unit(6,"pt"),
    barheight = unit(20,"pt"),
    frame.colour = "black",
    frame.linewidth = 0.3   # ← 改这里（比如 0.3 / 0.8）
  )
)
```

Save the plot

```{r}
#| echo: true
#| warning: false

hic_plot_final_canvas <- hic_plot_final + ggview::canvas(70,70,units="mm",dpi = 1200,bg="transparent")
hic_plot_final_canvas
#ggplot2::ggsave("figS1D.pdf", plot = hic_plot_final, width = 70, height = 70,units="mm", dpi = 600,bg="transparent")
```

## Other analysis that will not be used in the paper.

```{r}
#| echo: true
#| warning: false

phasing_track <- BSgenome.Dmelanogaster.JieHua.Ras3::BSgenome.Dmelanogaster.JieHua.Ras3
data_compts <- getCompartments(
  data,
  genome = phasing_track,
  BPPARAM = BiocParallel::SerialParam(),
  chromosomes = c("chrX", "chr2L", "chr2R", "chr3L", "chr3R", "chr4")
)

save(data_compts, file="Ras3_original_compartments.RData")
load("Ras3_original_compartments.RData")

data_selected_compts <- data_compts

topologicalFeatures(data_selected_compts, "compartments")


metadata(data_selected_compts)$eigens

library(GenomicRanges)
library(rtracklayer)
coverage(metadata(data_selected_compts)$eigens, weight = 'eigen') |> export('data_selected_compts.bw')
topologicalFeatures(data_selected_compts, "compartments") |> export('data_selected_compts.gff3')

library(ggplot2)
library(patchwork)
microC <- autocorrelate(data_selected)
##  

library(tibble)
p1 <- plotMatrix(microC, use.scores = 'autocorrelated', scale = 'linear', limits = c(-1, 1), caption = FALSE)
eigen <- coverage(metadata(data_selected_compts)$eigens, weight = 'eigen')[[1]]
eigen_df <- tibble(pos = cumsum(runLength(eigen)), eigen = runValue(eigen))
p2 <- ggplot(eigen_df, aes(x = pos, y = eigen)) + 
  geom_area() + 
  theme_void() + 
  coord_cartesian(expand = FALSE) + 
  labs(x = "Genomic position", y = "Eigenvector value")
wrap_plots(p1, p2, ncol = 1, heights = c(10, 1))


# Tests

normalized_hic <- normalize(data_selected)
cowplot::plot_grid(
  plotMatrix(normalized_hic, use.scores = 'count', caption = FALSE),
  plotMatrix(normalized_hic, use.scores = 'balanced', caption = FALSE),
  plotMatrix(normalized_hic, use.scores = 'ICE', caption = FALSE), 
  nrow = 1
)


detrended_hic <- detrend(data_selected)
detrended_hic
cowplot::plot_grid(
  plotMatrix(detrended_hic, use.scores = 'balanced', scale = 'log10', caption = FALSE),
  plotMatrix(detrended_hic, use.scores = 'expected', scale = 'log10', caption = FALSE),
  plotMatrix(detrended_hic, use.scores = 'detrended', scale = 'linear', cmap = bwrColors(), caption = FALSE), 
  nrow = 1
)



# paris? test

pf <- PairsFile(paste0(folder,"/Ras3_hic_mapping.mapped.pairs.gz"))
pf
ps <- distanceLaw(pf, by_chr = TRUE, filtered_chr=c("chrX", "chr2L", "chr2R", "chr3L", "chr3R", "chr4"))

library(ggplot2)
plotPs(ps, aes(x = binned_distance, y = norm_p, color = chr))
plotPsSlope(ps, aes(x = binned_distance, y = slope, color = chr))




# cis/trans ratio
ct <- cisTransRatio(data)
ct <- subset(ct, chr %in% c("chrX", "chr2L", "chr2R", "chr3L", "chr3R", "chr4"))
ggplot(ct, aes(x = chr, y = cis_pct)) + 
  geom_col(position = position_stack()) + 
  theme_bw() + 
  guides(x=guide_axis(angle = 90)) + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = 'Chromosomes', y = '% of cis contacts')



# 其实可以让数据变成matrix进行3D结构模拟的
test <- as.matrix(data_selected, use.scores = "balanced")
test <- as.matrix(data_selected, use.scores = "count")

```
