---
title: "Fig2A Genomic Coverage of hybrid cell lines"
author: "Jie Hua"
date: 2025-11-26
lang: en
date-modified: 2025-11-26
keywords: ["Genomic Coverage","Hybrid Cells"]
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: true
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load the required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(dplyr)
library(data.table)
library(ggplot2)
library(cowplot)
library(viridis) 
library(stringr)

```

# Define the required parameters

```{r}
#| echo: true
#| warning: false

# Define the folder containing the bedgraph files
folder <- "F:\\Big Paper\\Processed data\\Fig3_Genomic_coverage\\100bp_window"
# The genes annotation
gtf <- "F:\\Big Paper\\Processed data\\Fig3_Genomic_coverage\\All.extended_annotation.gtf"
# The RepeatMakser annotation
repeatmasker_gff <- "F:\\Big Paper\\Processed data\\Fig3_Genomic_coverage\\ras3_dsim3_hybrid.repeatmasker.out.gff"

# Lab PC
# Define the folder containing the bedgraph files
#folder <- "H:\\HybridCells_Data\\Fig2_Genomic_coverage\\100bp_window"
# The genes annotation
#gtf <- "H:\\HybridCells_Data\\Fig2_Genomic_coverage\\All.extended_annotation.gtf"
# The RepeatMakser annotation
#repeatmasker_gff <- "H:\\HybridCells_Data\\Fig2_Genomic_coverage\\ras3_dsim3_hybrid.repeatmasker.out.gff"



```

# Load the coverage data

I already used *bamcoverage* to transfer *bam* files to *bed* files.

```{r}
#| echo: true
#| warning: false
cat("The folder containing the bedgraph files is", folder, "\n")
# List the bedGraph files
bedgraph_files <- list.files(folder, pattern = "\\.bedGraph$", full.names = TRUE)
cat("The bedGraph files (total number:", length(bedgraph_files), "):\n")
print(bedgraph_files)
```

Load the *bedgraph* files to R

```{r}
#| echo: true
#| warning: false

data_list <- list()
for (file in bedgraph_files) {
  data <- read.table(file, header = FALSE, sep = "\t", stringsAsFactors = FALSE)
  colnames(data) <- c("Chromosome", "Start", "End", "Counts")
  data$total <- (data$End - data$Start)*data$Counts
  file_name <- basename(file)
  name <- sub("_q30_coverage_100bp\\.bedGraph$", "", file_name)
  data_list[[name]] <- data
}
cat("The names of the list containing all the bedgraph tables:\n")
names(data_list)

```

I need to filter the data to select specific samples and chromosomes, and rename them.

```{r}
#| echo: true
#| warning: false
selected_sample <- c("Ras3","ML82","05a","05b_barcode18","06b","07b_barcode22")
renamed_sample  <- c("Ras3","ML82","HYC1","HYC2","HYC3","HYC4")

selected_chromosomes <- c("chrX_mel", "chr2L_mel", "chr2R_mel", "chr3L_mel","chr3R_mel","chr4_mel","chrX_sim", "chr2L_sim", "chr2R_sim", "chr3L_sim","chr3R_sim","chr4_sim")

renamed_chromosomes <- c("chrX_mel", "chr2L_mel", "chr2R_mel", "chr3L_mel","chr3R_mel","chr4_mel","chrX_sim", "chr2L_sim", "chr2R_sim", "chr3L_sim","chr3R_sim","chr4_sim")

# Build mapping relationship
sample_map <- setNames(renamed_sample, selected_sample)
chrom_map  <- setNames(renamed_chromosomes, selected_chromosomes)

filtered_list <- lapply(selected_sample, function(smp) {
  df <- data_list[[smp]]
  df <- df[df$Chromosome %in% selected_chromosomes, ]
  df$Chromosome <- chrom_map[df$Chromosome]
  rownames(df) <- NULL
  return(df)
})

names(filtered_list) <- sample_map
```

I want to check if it works.

```{r}
#| echo: true
#| warning: false

print(names(filtered_list))
print(unique(filtered_list$Ras3$Chromosome))
head(filtered_list$Ras3)
```

Merge all tables into one dataframe

```{r}
#| echo: true
#| warning: false

data_all_raw <- bind_rows(
  lapply(names(filtered_list), function(nm) {
    df <- filtered_list[[nm]]
    df$Sample <- nm
    df
  })
)
```

# Normalization

I need to check the distribution of counts first.

```{r}
#| echo: true
#| warning: false
#| fig-align: center
ggplot(data_all_raw, aes(x = Counts, weight = End - Start)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5) +
  facet_wrap(~Sample, scales = "free_y") +
  labs(
    title = "Counts Density Distribution Across Samples",
    x = "Counts",
    y = "Density"
  ) +
  theme_minimal(base_family = "sans") + 
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black")
  )
```

Apparently, I should perform transformation to make it like normal distribution.

```{r}
#| echo: true
#| warning: false

data_all <- data_all_raw %>%
  mutate(Counts_log2 = log2(Counts + 1))
```

Then I could check the distribution with log2 transfered counts

```{r}
#| echo: true
#| warning: false
#| fig-align: center

ggplot(data_all, aes(x = Counts_log2, weight = End - Start)) +
  geom_density(fill = "skyblue", color = "black", alpha = 0.5) +
  facet_wrap(~Sample, scales = "free_y") +
  labs(
    title = "Log2(Counts+1) Density Distribution Across Samples",
    x = "log2(Counts + 1)",
    y = "Density"
  ) +
  theme_minimal(base_family = "sans") +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 0.6), # 边框稍粗
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black")
  )

```

# Plot the coverage information

```{r}
#| echo: true
#| warning: false

# ============================================================
# Theme for genome coverage plot
# ============================================================

theme_genome_coverage <- function(my_family = "sans") {
  theme_bw(base_family = my_family) +
    theme(
      # Transparent backgrounds
      plot.background = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(fill = "transparent", color = NA),
      
      # Grid
      panel.grid = element_blank(),
      
      # Border
      panel.border = element_rect(
        color = "black",
        fill = NA,
        linewidth = 0.1
      ),
      
      # Avoid double-thick borders
      axis.line = element_blank(),
      
      # Facet labels
      strip.text.x = element_text(
        angle = 0,
        hjust = 0.5,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 1, r = 0, b = 1, l = 0)
      ),
      strip.text.y.left = element_text(
        angle = 0,
        hjust = 1,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 0, r = 3, b = 0, l = 1)
      ),
      
      # Axis text
      axis.text.x = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      axis.text.y = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      
      # Axis titles
      axis.title.x = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      axis.title.y = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      
      # Major ticks
      axis.ticks.x = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.y = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.length = unit(1.5, "pt"),
      
      # Minor ticks
      axis.minor.ticks.x = element_line(
        linewidth = 0.1,
        color = "black"
      ),
      axis.minor.ticks.y = element_line(
        linewidth = 0.1,
        color = "black"
      ),
      axis.minor.ticks.length.x = unit(0.8, "pt"),
      axis.minor.ticks.length.y = unit(0.8, "pt"),
      
      # Compact legend
      legend.position = "right",
      legend.title = element_blank(),
      legend.text = element_text(
        size = 6,
        color = "black",
        family = my_family,
        face = "italic",
        margin = margin(t = 0, r = 0, b = 0, l = 2)
      ),
      legend.key.size = unit(6, "pt"),
      legend.key.width = unit(6, "pt"),
      legend.key.height = unit(6, "pt"),
      legend.key.spacing.x = unit(1, "pt"),
      legend.key.spacing.y = unit(1, "pt"),
      
      legend.spacing.x = unit(1, "pt"),
      legend.spacing.y = unit(1, "pt"),
      legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.spacing = unit(2, "pt"),
      
      # Facet spacing
      panel.spacing.x = unit(0.2, "lines"),
      panel.spacing.y = unit(0.4, "lines"),
      strip.placement = "outside",
      strip.switch.pad.grid = unit(0, "pt"),
      strip.clip = "off",
      
      # Plot margin
      plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
    )
}


# ============================================================
# Genome coverage plot function
# ============================================================

plot_genome_coverage <- function(
  data_all,
  renamed_chromosomes,
  show_length = FALSE,
  sample_order = NULL,
  my_family = "sans"
) {
  
  # ------------------------------------------------------------
  # Add Sample_group
  # ------------------------------------------------------------
  
  data_all <- data_all %>%
    mutate(
      Sample_group = case_when(
        Sample == "Ras3" ~ "D. mel",
        Sample == "ML82" ~ "D. sim",
        TRUE ~ "Hybrid"
      )
    )
  
  # ------------------------------------------------------------
  # Group colors
  # ------------------------------------------------------------
  
  sample_colors <- c(
    "D. mel" = "#fb8b05",
    "D. sim" = "#ed556a",
    "Hybrid" = "#61649F"
  )
  
  # ------------------------------------------------------------
  # Sample order
  # facet_grid rows follow factor levels from top to bottom
  # ------------------------------------------------------------
  
  if (is.null(sample_order)) {
    sample_order <- unique(data_all$Sample)
  }
  
  missing_samples <- setdiff(sample_order, unique(data_all$Sample))
  
  if (length(missing_samples) > 0) {
    warning(
      "These samples in sample_order are not found in data_all$Sample: ",
      paste(missing_samples, collapse = ", ")
    )
  }
  
  # ------------------------------------------------------------
  # Chromosome information and labels
  # ------------------------------------------------------------
  
  chrom_info <- data_all %>%
    group_by(Chromosome) %>%
    summarise(
      Length = max(End, na.rm = TRUE) - min(Start, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      Chromosome_short = sub("^chr", "", Chromosome),
      Chromosome_label = if (show_length) {
        paste0(
          Chromosome_short,
          "\n(",
          round(Length / 1e6, 2),
          " Mb)"
        )
      } else {
        Chromosome_short
      }
    )
  
  # ------------------------------------------------------------
  # Check chromosome order
  # ------------------------------------------------------------
  
  missing_chromosomes <- setdiff(renamed_chromosomes, chrom_info$Chromosome)
  
  if (length(missing_chromosomes) > 0) {
    warning(
      "These chromosomes in renamed_chromosomes are not found in data_all$Chromosome: ",
      paste(missing_chromosomes, collapse = ", ")
    )
  }
  
  # Only keep chromosomes that exist in data
  renamed_chromosomes_use <- intersect(
    renamed_chromosomes,
    chrom_info$Chromosome
  )
  
  chrom_info <- chrom_info %>%
    slice(match(renamed_chromosomes_use, Chromosome))
  
  # ------------------------------------------------------------
  # Merge labels back to data
  # ------------------------------------------------------------
  
  data_all <- data_all %>%
    left_join(chrom_info, by = "Chromosome") %>%
    filter(!is.na(Chromosome_label)) %>%
    mutate(
      Chromosome_label = factor(
        Chromosome_label,
        levels = chrom_info$Chromosome_label
      ),
      Sample = factor(
        Sample,
        levels = sample_order
      ),
      Sample_group = factor(
        Sample_group,
        levels = c("D. mel", "D. sim", "Hybrid")
      )
    )
  
  # ------------------------------------------------------------
  # Plot
  # ------------------------------------------------------------
  
  p <- ggplot(data_all) +
    geom_rect(
      aes(
        xmin = Start,
        xmax = End,
        ymin = 0,
        ymax = Counts_log2,
        fill = Sample_group
      ),
      color = NA,
      linewidth = 0,
      alpha = 0.95
    ) +
    facet_grid(
      rows = vars(Sample),
      cols = vars(Chromosome_label),
      scales = "free_x",
      space = "free_x",
      switch = "both"
    ) +
    scale_fill_manual(
      values = sample_colors,
      name = "Sample",
      drop = FALSE
    ) +
    scale_x_continuous(
      expand = c(0, 0),
      breaks = function(x) {
        seq(
          from = floor(min(x, na.rm = TRUE) / 10e6) * 10e6,
          to = ceiling(max(x, na.rm = TRUE) / 10e6) * 10e6,
          by = 10e6
        )
      },
      minor_breaks = function(x) {
        seq(
          from = floor(min(x, na.rm = TRUE) / 2e6) * 2e6,
          to = ceiling(max(x, na.rm = TRUE) / 2e6) * 2e6,
          by = 2e6
        )
      },
      guide = guide_axis(minor.ticks = TRUE),
      labels = function(x) {
        paste0(x / 1e6)
      }
    ) +
    scale_y_continuous(
      breaks = c(0, 8),
      minor_breaks = seq(0, 8, by = 2),
      limits = c(0, 10),
      expand = c(0, 0.2),
      guide = guide_axis(minor.ticks = TRUE),
      labels = function(x) {
        ifelse(x == 0, "0", paste0("\u2007\u2007", x))
      }
    ) +
    labs(
      x = "Chromosome (Mb)",
      y = expression(log[2](Counts + 1))
    ) +
    theme_genome_coverage(my_family = my_family) +
    guides(
      fill = guide_legend(
        keywidth = unit(6, "pt"),
        keyheight = unit(6, "pt"),
        ncol = 1,
        bycol = TRUE,
        label.position = "right",
        label.hjust = 0,
        override.aes = list(
        colour = NA,
        alpha = 0.95
      )
      )
    )
  
  return(p)
}

# 使用示例 
p1 <- plot_genome_coverage(data_all, renamed_chromosomes, F)

```

Save the plot

```{r}
#| echo: true
#| warning: false

# Have a look to confirm the size
p1_canvas <- p1+ggview::canvas(166, 50, units="mm",dpi=200,bg="transparent")
p1_canvas
#ggsave("fig2A.pdf",p1,width=166,height=50,units="mm",dpi=900,bg="transparent")
```
