---
title: "Fig.3C Dosage compensation and Transcriptome"
author: "Jie Hua"
date: 2025-10-01
lang: en
date-modified: 2026-05-04
keywords: ["Dosage compensation","Transcriptome"]
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: false
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(tidyr)
library(dplyr)
library(ggplot2)
library(viridis)
library(stringr)
```

# Define some functions

```{r}
#| echo: true
#| warning: false

source("functions.R")

```

# Load the files

## Define a folder

```{r}
#| echo: true
#| warning: false

folder_path <- "H:\\HybridCells_Data\\Fig3_Dosage and Transcriptome"
#folder_path <- "F:\\Big Paper\\Processed data\\Fig3C_Dosage and Transcriptome"
```

## Define main chromosomes

```{r}
#| echo: true
#| warning: false

selected_chromosomes <- c("chr2L_mel","chr2R_mel","chr3L_mel","chr3R_mel","chr4_mel","chrX_mel","chr2L_sim","chr2R_sim","chr3L_sim","chr3R_sim","chr4_sim","chrX_sim")

```

## Read the genomic coverage generates by *mosdepth*

I already got the genomic coverage on all annotated genes.

```{r}
#| echo: true
#| warning: false

folder_genome <- paste0(folder_path,"/genomic")

data_genome <- read_genomic_coverage(folder_genome, pattern="\\.regions\\.bed$",cutoff=1)

info <- visualize_genome(data_genome, selected_chromosomes,facet_by_chrom = TRUE)

genome_totalbases <- read_totalbases(bam_summary=paste0(folder_path,"/","bam_summary_genome.csv"))

data_genome_normalized <- normalize_genome(data_genome,genome_totalbases,selected_chromosomes)

info <- visualize_genome(data_genome_normalized, selected_chromosomes,facet_by_chrom = TRUE)

#write.csv(data_genome, "C:/temp/GenomicCoverageData.csv", row.names = FALSE)
```

## Read transcriptome counts and normalize

```{r}
#| echo: true
#| warning: false


data_transcriptome <- read_transcriptome_counts(transcriptome=paste0(folder_path,"/transcriptome/","All.discovered_gene_grouped_counts.tsv"),gtf=paste0(folder_path,"/transcriptome/","All.extended_annotation.gtf"))

info <- visualize_transcriptome(data_transcriptome,selected_chromosomes,facet_by_chrom = TRUE)

data_transcriptome_normalized  <- normalize_transcriptome(data_transcriptome,selected_chromosomes)

info <- visualize_transcriptome(data_transcriptome_normalized,selected_chromosomes,facet_by_chrom = TRUE)

#write.csv(data_transcriptome, "C:/temp/TranscriptomeCountsData.csv", row.names = FALSE)
```

# Comparison

```{r}
#| echo: true
#| warning: false

# Sample map list
sample_map <- list(
  "Ras3" = c("Ras3_rep1", "Ras3_rep2", "Ras3_rep3"),
  "ML82" = c("ML82_rep1", "ML82_rep2", "ML82_rep3"),
  "HYC1" = c("5A_rep1", "5A_rep2", "5A_rep3"),
  "HYC2" = c("5B_rep1", "5B_rep2", "5B_rep3"),
  "HYC3" = c("6B_rep1", "6B_rep2", "6B_rep3"),
  "HYC4" = c("7B_rep1", "7B_rep2", "7B_rep3"),
  "ML82" = c("ML82_rep1", "ML82_rep2", "ML82_rep3")
)
# Transcriptome to genome
genome_col_map <- list(
  "Ras3" = "Ras3_q30",
  "ML82" = "ML82_q30",
  "HYC1" = "05a_q30",
  "HYC2" = "05b_barcode19_q30",
  "HYC3" = "06b_q30",
  "HYC4" = "07b_barcode23_q30"
)
comparison <- compare_transcriptome_genome(data_transcriptome_normalized, data_genome_normalized, sample_map, genome_col_map)
#write.csv(comparison,"comparison.csv",row.names=F)

```

## Plot the distribution of the expression level matrix

```{r}
#| echo: true
#| warning: false

visualize_comparison_distribution(comparison)
visualize_comparison_distribution_by_chrom(comparison, selected_chromosomes)
visualize_comparison_summary(comparison)
```

# Summarize to final barplot (Mean ± SD)

## Convert to wide-format

```{r}
#| echo: true
#| warning: false

comparison <- as.data.table(comparison)
sample_cols <- setdiff(colnames(comparison), c("chrom","start","end","gene"))

long_comparison  <- melt(
  comparison ,
  id.vars = c("chrom","start","end","gene"),
  measure.vars = sample_cols,
  variable.name = "sample",
  value.name = "expression"
)

long_comparison <- as.data.table(long_comparison)

#write.csv(long_comparison,"long_comparison.csv",row.names=F)
```

## Add group name

```{r}
#| echo: true
#| warning: false

long_comparison[, sample_group := str_extract(sample, "^[0-9A-Za-z]+")]
```

## *D. mel* chromosomes

```{r}
#| echo: true
#| warning: false

my_family="sans"

sample_order <- c("Ras3","HYC1","HYC2","HYC3","HYC4")

mel_dt <- long_comparison[
  grepl("mel$", chrom) &
    (grepl("^[0-9]", sample_group) | sample_group == "Ras3")
]

mel_sum <- mel_dt[, .(
  sum_expr = sum(expression, na.rm = TRUE)
), by = .(sample_group, chrom, sample)]

mel_summary <- mel_sum[, .(
  mean_sum = mean(sum_expr, na.rm = TRUE),
  sd_sum   = sd(sum_expr, na.rm = TRUE)
), by = .(sample_group, chrom)]

mel_summary <- mel_summary %>%
  mutate(
    sample_group = case_when(
      sample_group == "5A" ~ "HYC1",
      sample_group == "5B" ~ "HYC2",
      sample_group == "6B" ~ "HYC3",
      sample_group == "7B" ~ "HYC4",
      TRUE ~ sample_group
    ),
    sample_group = factor(sample_group, levels = sample_order)
  )

p1 <- ggplot(
  mel_summary,
  aes(x = chrom, y = mean_sum, fill = sample_group)
) +
  geom_col(
    width = 0.8,
    position = position_dodge(width = 0.8),
    color = "black",
    linewidth = 0.1
  ) +
  geom_errorbar(
    aes(
      ymin = mean_sum - sd_sum,
      ymax = mean_sum + sd_sum
    ),
    width = 0.2,
    linewidth = 0.2,
    position = position_dodge(width = 0.8),
    color = "black"
  ) +
  scale_fill_viridis_d(
    option = "magma",
    begin = 0,
    end = 0.9,
    limits = sample_order,
    drop = FALSE,
    name = "Sample"
  ) +
  labs(
    title = expression(italic("D. mel") ~ " chromosomes: Hybrids and Ras3"),
    y = "Sum of Relative Expression Level",
    x = "Chromosome"
  ) +
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Title
    plot.title = element_text(
      size = 7,
      color = "black",
      family = my_family,
      face = "plain",
      hjust = 0.5,
      margin = margin(b = 3)
    ),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      angle = 45,
      hjust = 1,
      vjust = 1,
      color = "black",
      family = my_family
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Compact legend
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 0, l = 2)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    # Panel spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  ) +
  guides(
    fill = guide_legend(
      keywidth = unit(6, "pt"),
      keyheight = unit(6, "pt"),
      ncol = 1,
      bycol = TRUE,
      label.position = "right",
      label.hjust = 0,
      override.aes = list(
        colour = NA,
        alpha = 0.95
      )
    )
  )

```

View and save

```{r}
#| echo: true
#| warning: false

p1_canvas <- p1+ggview::canvas(70,50,units="mm",dpi=900,bg="transparent")
p1_canvas
#ggsave("dc_dmel.png",p1,width=7, height=5,units="in",dpi = 1200)

```

## *D. sim* chromosomes

```{r}
#| echo: true
#| warning: false

sample_order_sim <- c("ML82","HYC1","HYC2","HYC3","HYC4")

sim_dt <- long_comparison[
  grepl("sim$", chrom) &
    (grepl("^[0-9]", sample_group) | sample_group == "ML82")
]

sim_sum <- sim_dt[, .(
  sum_expr = sum(expression, na.rm = TRUE)
), by = .(sample_group, chrom, sample)]

sim_summary <- sim_sum[, .(
  mean_sum = mean(sum_expr, na.rm = TRUE),
  sd_sum   = sd(sum_expr, na.rm = TRUE)
), by = .(sample_group, chrom)]

sim_summary <- sim_summary %>%
  mutate(
    sample_group = case_when(
      sample_group == "5A" ~ "HYC1",
      sample_group == "5B" ~ "HYC2",
      sample_group == "6B" ~ "HYC3",
      sample_group == "7B" ~ "HYC4",
      TRUE ~ sample_group
    ),
    sample_group = factor(sample_group, levels = sample_order_sim)
  ) %>%
  filter(sample_group != "Ras3")

p2 <- ggplot(
  sim_summary,
  aes(x = chrom, y = mean_sum, fill = sample_group)
) +
  geom_col(
    width = 0.8,
    position = position_dodge(width = 0.8),
    color = NA,
    linewidth = 0.1
  ) +
  geom_errorbar(
    aes(
      ymin = mean_sum - sd_sum,
      ymax = mean_sum + sd_sum
    ),
    width = 0.2,
    linewidth = 0.2,
    position = position_dodge(width = 0.8),
    color = "black"
  ) +
  scale_fill_viridis_d(
    option = "viridis",
    begin = 0,
    end = 0.9,
    limits = sample_order_sim,
    drop = FALSE,
    name = "Sample"
  ) +
  labs(
    title = expression(italic("D. sim") ~ " chromosomes: Hybrids and ML82"),
    y = "Sum of Relative Expression Level",
    x = "Chromosome"
  ) +
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Title
    plot.title = element_text(
      size = 7,
      color = "black",
      family = my_family,
      face = "plain",
      hjust = 0.5,
      margin = margin(b = 3)
    ),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      angle = 45,
      hjust = 1,
      vjust = 1,
      color = "black",
      family = my_family
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Compact legend
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 0, l = 2)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    # Panel spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  ) +
  guides(
    fill = guide_legend(
      keywidth = unit(6, "pt"),
      keyheight = unit(6, "pt"),
      ncol = 1,
      bycol = TRUE,
      label.position = "right",
      label.hjust = 0,
      override.aes = list(
        colour = NA,
        alpha = 0.95
      )
    )
  )
```

View and save

```{r}
#| echo: true
#| warning: false

p2_canvas <- p2+ggview::canvas(70,50,units="mm",dpi=900,bg="transparent")
p2_canvas

#ggsave("dc_dsim.png",p2,width=70, height=50,units="mm",dpi = 1200,bg="transparent")
```

# Comparison Plot

```{r}
#| echo: true
#| warning: false


sample_sum <- long_comparison %>%
  mutate(
    species = case_when(
      str_detect(chrom, "_mel$") ~ "mel",
      str_detect(chrom, "_sim$") ~ "sim",
      TRUE ~ NA_character_
    ),
    chrom_simple = str_remove(chrom, "_(mel|sim)$")
  ) %>%
  filter(!is.na(species)) %>%
  group_by(sample_group, sample, chrom_simple, species) %>%
  summarise(
    expr_sum = sum(expression, na.rm = TRUE),
    .groups = "drop"
  )

group_stat <- sample_sum %>%
  group_by(sample_group, chrom_simple, species) %>%
  summarise(
    mean_expr = mean(expr_sum),
    sd_expr   = sd(expr_sum),
    .groups = "drop"
  )

plot_df <- group_stat %>%
  mutate(
    direction = if_else(species == "mel", -1, 1),
    mean_plot = mean_expr * direction
  ) %>%
  filter(
    !(species == "mel" & sample_group == "ML82"),
    !(species == "sim" & sample_group == "Ras3")
  )

group_name_map <- c(
  "5A" = "HYC1",
  "5B" = "HYC2",
  "6B" = "HYC3",
  "7B" = "HYC4",
  "ML82" = "ML82",
  "Ras3" = "Ras3"
)

plot_df <- plot_df %>%
  mutate(
    sample_group_label = recode(sample_group, !!!group_name_map)
  )


plot_df <- plot_df %>%
  mutate(
    chrom_simple = factor(
      chrom_simple,
      levels = c("chr2L","chr2R","chr3L","chr3R","chr4","chrX")
    ),
    sample_group_label = factor(
      sample_group_label,
      levels = c("Ras3","ML82","HYC1","HYC2","HYC3","HYC4")
    )
  )

group_colors <- c(
  "HYC1"   = "#0072B2",  # blue
  "HYC2"   = "#fb8b05",  # sky blue
  "HYC3"   = "#ed556a",  # bluish green
  "HYC4"   = "#56B4E9",  # orange
  "ML82" = "#CC79A7",  # reddish purple
  "Ras3" = "#61649F"   # dark purple (no grey)
)



p <- ggplot(
  plot_df,
  aes(
    x = chrom_simple,
    y = mean_plot,
    fill = sample_group_label
  )
) +
  geom_col(
    position = position_dodge(width = 0.8),
    width = 0.7,
    color = NA,
    linewidth = 0.1
  ) +
  geom_errorbar(
    aes(
      ymin = mean_plot - sd_expr,
      ymax = mean_plot + sd_expr
    ),
    position = position_dodge(width = 0.8),
    width = 0.2,
    linewidth = 0.2,
    color = "black"
  ) +
  geom_hline(
    yintercept = 0,
    linewidth = 0.1,
    color = "black"
  ) +
  scale_fill_manual(
    values = group_colors,
    name = "Sample"
  ) +
  scale_y_continuous(
    breaks = c(
      -max(abs(plot_df$mean_plot + plot_df$sd_expr), na.rm = TRUE),
       0,
       max(abs(plot_df$mean_plot + plot_df$sd_expr), na.rm = TRUE)
    ),
    labels = c("D. mel", "0", "D. sim"),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  labs(
    x = "Chromosome",
    y = "Summed RNA/DNA-normalized Expression"
  ) +
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      angle = 45,
      hjust = 1,
      vjust = 1,
      color = "black",
      family = my_family
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Minor ticks
    axis.minor.ticks.x = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.y = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),
    
    # Compact legend
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 0, l = 2)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    # Panel spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  ) +
  guides(
    fill = guide_legend(
      keywidth = unit(6, "pt"),
      keyheight = unit(6, "pt"),
      ncol = 1,
      bycol = TRUE,
      label.position = "right",
      label.hjust = 0,
      override.aes = list(
        colour = NA,
        alpha = 0.95
      )
    )
  )

```

```{r}
#| echo: true
#| warning: false

p_canvas <- p+ggview::canvas(90,60,units="mm",dpi=900,bg="transparent")
p_canvas

#ggsave("dosage_compensation.svg",p,width=90, height=60,units="mm",dpi = 1200)
```

# Comparison Plot version2

```{r}
#| echo: true
#| warning: false

library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(cowplot)
library(grid)

# 如果你前面没有定义 my_family，就先用 sans
my_family <- "sans"

# =============================
# 1. 整理数据
# =============================
sample_sum <- long_comparison %>%
  mutate(
    species = case_when(
      str_detect(chrom, "_mel$") ~ "D.mel",
      str_detect(chrom, "_sim$") ~ "D.sim",
      TRUE ~ NA_character_
    ),
    chrom_simple = str_remove(chrom, "_(mel|sim)$")
  ) %>%
  filter(!is.na(species)) %>%
  group_by(sample_group, sample, chrom_simple, species) %>%
  summarise(
    expr_sum = sum(expression, na.rm = TRUE),
    .groups = "drop"
  )

group_stat <- sample_sum %>%
  group_by(sample_group, chrom_simple, species) %>%
  summarise(
    mean_expr = mean(expr_sum, na.rm = TRUE),
    sd_expr   = sd(expr_sum, na.rm = TRUE),
    .groups = "drop"
  )

# =============================
# 2. 名称映射和顺序
# =============================
group_name_map <- c(
  "5A"   = "HYC1",
  "5B"   = "HYC2",
  "6B"   = "HYC3",
  "7B"   = "HYC4",
  "ML82" = "ML82",
  "Ras3" = "Ras3"
)

chrom_levels   <- c("chr2L", "chr2R", "chr3L", "chr3R", "chr4", "chrX")
sample_levels  <- c("Ras3", "ML82", "HYC1", "HYC2", "HYC3", "HYC4")
species_levels <- c("D.sim", "D.mel")

# =============================
# 3. complete() 补全所有组合
# =============================
plot_df <- group_stat %>%
  mutate(
    sample_group_label = recode(sample_group, !!!group_name_map)
  ) %>%
  select(sample_group_label, chrom_simple, species, mean_expr, sd_expr) %>%
  complete(
    sample_group_label = sample_levels,
    chrom_simple = chrom_levels,
    species = species_levels,
    fill = list(mean_expr = 0, sd_expr = 0)
  ) %>%
  mutate(
    sample_group_label = factor(sample_group_label, levels = sample_levels),
    chrom_simple = factor(chrom_simple, levels = chrom_levels),
    species = factor(species, levels = species_levels)
  ) %>%
  mutate(
    mean_expr = case_when(
      sample_group_label == "ML82" & species == "D.mel" ~ 0,
      sample_group_label == "Ras3" & species == "D.sim" ~ 0,
      TRUE ~ mean_expr
    ),
    sd_expr = case_when(
      sample_group_label == "ML82" & species == "D.mel" ~ 0,
      sample_group_label == "Ras3" & species == "D.sim" ~ 0,
      TRUE ~ sd_expr
    )
  )

# =============================
# 4. 分上下两个 panel 的数据
# =============================
chrom_index <- setNames(seq_along(chrom_levels), chrom_levels)

# ✅ 1. 增大 Chromosome 之间的距离
# 原来是 0.4；这里改成 0.65
gap <- 0.3

plot_df_sim <- plot_df %>%
  filter(species == "D.sim") %>%
  mutate(
    ymin = pmax(mean_expr - sd_expr, 0),
    ymax = mean_expr + sd_expr,
    x_base = chrom_index[as.character(chrom_simple)],
    x = x_base * (1 + gap)
  )

plot_df_mel <- plot_df %>%
  filter(species == "D.mel") %>%
  mutate(
    mean_plot = -mean_expr,
    ymin = -(mean_expr + sd_expr),
    ymax = -pmax(mean_expr - sd_expr, 0),
    x_base = chrom_index[as.character(chrom_simple)],
    x = x_base * (1 + gap)
  )

# 统一上下图的绝对范围
y_abs_max <- max(
  c(
    plot_df_sim$ymax,
    abs(plot_df_mel$ymin)
  ),
  na.rm = TRUE
)

y_breaks <- pretty(c(0, y_abs_max), n = 5)
y_breaks <- y_breaks[y_breaks >= 0]

# ✅ 2. 增加 Y 轴小刻度数量
# 数字越大，小刻度越密；21 比较适中，31 更密
y_minor_breaks <- seq(
  0,
  ceiling(y_abs_max / 2000) * 2000,
  by = 2000
)

x_breaks <- plot_df_sim %>%
  distinct(chrom_simple, x) %>%
  arrange(match(chrom_simple, chrom_levels)) %>%
  pull(x)

# =============================
# 5. 颜色
# =============================
group_colors <- c(
  "Ras3" = "#61649F",
  "ML82" = "#CC79A7",
  "HYC1" = "#0072B2",
  "HYC2" = "#fb8b05",
  "HYC3" = "#ed556a",
  "HYC4" = "#56B4E9"
)

# =============================
# 6. 通用主题
# =============================
base_theme <- theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Grid
    panel.grid = element_blank(),
    
    # Border
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 0.1
    ),
    
    # Avoid double-thick borders
    axis.line = element_blank(),
    
    # Facet labels
    strip.text.x = element_text(
      angle = 0,
      hjust = 0.5,
      size = 6,
      family = my_family,
      color = "black",
      face = "italic",
      margin = margin(t = 1, r = 0, b = 3, l = 0)
    ),
    strip.text.y.left = element_text(
      angle = 0,
      hjust = 1,
      size = 6,
      family = my_family,
      color = "black",
      face = "italic",
      margin = margin(t = 0, r = 3, b = 0, l = 1)
    ),
    
    # Axis text
    axis.text.x = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    
    # Axis titles
    axis.title.x = element_text(
      size = 7,
      color = "black",
      family = my_family
    ),
    axis.title.y = element_text(
      size = 7,
      color = "black",
      family = my_family,
      face="italic"
    ),
    
    # Major ticks
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.length = unit(1.5, "pt"),
    
    # Minor ticks
    axis.minor.ticks.x = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.y = element_line(
      linewidth = 0.1,
      color = "black"
    ),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),
    
    # Compact legend
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 0, l = 2)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    # Facet spacing
    panel.spacing.x = unit(0.2, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    strip.placement = "outside",
    strip.switch.pad.grid = unit(0, "pt"),
    strip.clip = "off",
    
    # General text
    text = element_text(
      family = my_family,
      color = "black"
    ),
    
    # Plot margin
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  )

# =============================
# 7. 上 panel：D.sim
# =============================
p_top_core <- ggplot(
  plot_df_sim,
  aes(x = x, y = mean_expr, fill = sample_group_label)
) +
  geom_col(
    # ✅ 3. 柱子宽度增加
    # position_dodge 也同步加大，保证 errorbar 对齐
    position = position_dodge(width = 1.0),
    width = 0.9,
    color = NA,
    linewidth = 0.1
  ) +
  geom_errorbar(
    aes(ymin = ymin, ymax = ymax),
    position = position_dodge(width = 1.0),
    width = 0.16,
    linewidth = 0.2,
    color = "black"
  ) +
  geom_hline(
    yintercept = 0,
    linewidth = 0.05,
    color = "black"
  ) +
  scale_fill_manual(
    values = group_colors,
    drop = FALSE,
    name = "Sample"
  ) +
  scale_x_continuous(
    breaks = x_breaks,
    labels = chrom_levels,
    expand = c(0.03, 0.03)
  ) +
  scale_y_continuous(
    limits = c(0, y_abs_max * 1.02),
    breaks = y_breaks,
    minor_breaks = y_minor_breaks,
    expand = c(0, 0),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  labs(x = NULL, y = NULL, fill = "Sample") +
  base_theme +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.minor.ticks.x = element_blank(),
    legend.position = "none",
    plot.margin = margin(0, 0, 0, 0)
  )+
  guides(
      fill = guide_legend(
        keywidth = unit(6, "pt"),
        keyheight = unit(6, "pt"),
        ncol = 1,
        bycol = TRUE,
        label.position = "right",
        label.hjust = 0,
        override.aes = list(
        colour = NA,
        alpha = 0.95
      )
      )
    )

# =============================
# 8. 下 panel：D.mel
# =============================
p_bottom_core <- ggplot(
  plot_df_mel,
  aes(x = x, y = mean_plot, fill = sample_group_label)
) +
  geom_col(
    # ✅ 3. 柱子宽度增加
    position = position_dodge(width = 1.0),
    width = 0.9,
    color = NA,
    linewidth = 0.1
  ) +
  geom_errorbar(
    aes(ymin = ymin, ymax = ymax),
    position = position_dodge(width = 1.0),
    width = 0.16,
    linewidth = 0.2,
    color = "black"
  ) +
  geom_hline(
    yintercept = 0,
    linewidth = 0.05,
    color = "black"
  ) +
  scale_fill_manual(
    values = group_colors,
    drop = FALSE,
    name = "Sample"
  ) +
  scale_x_continuous(
    breaks = x_breaks,
    labels = chrom_levels,
    expand = c(0.03, 0.03)
  ) +
  scale_y_continuous(
    limits = c(-y_abs_max * 1.02, 0),
    breaks = -rev(y_breaks),
    minor_breaks = -rev(y_minor_breaks),
    labels = function(x) abs(x),
    expand = c(0, 0),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  labs(x = NULL, y = NULL, fill = "Sample") +
  base_theme +
  theme(
    axis.text.x = element_text(
      size = 6,
      angle = 0,
      hjust = 0.5,
      vjust = 1,
      family = my_family,
      color = "black"
    ),
    axis.text.y = element_text(
      size = 6,
      color = "black",
      family = my_family
    ),
    axis.ticks.x = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    axis.ticks.y = element_line(
      linewidth = 0.2,
      color = "black"
    ),
    legend.position = "none",
    plot.margin = margin(0, 0, 0, 0)
  )

# =============================
# 9. 提取统一图例
# =============================
legend_plot <- ggplot(
  plot_df,
  aes(
    x = chrom_simple,
    y = abs(ifelse(species == "D.mel", -mean_expr, mean_expr)),
    fill = sample_group_label
  )
) +
  geom_col(
    color = "black",
    linewidth = 0.1
  ) +
  scale_fill_manual(
    values = group_colors,
    drop = FALSE,
    name = "Sample"
  ) +
  labs(fill = "Sample") +
  theme_bw(base_family = my_family) +
  theme(
    # Transparent backgrounds
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    # Hide plot body
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    
    # Compact legend
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(
      size = 6,
      color = "black",
      family = my_family,
      margin = margin(t = 0, r = 0, b = 0, l = 2)
    ),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.spacing = unit(2, "pt"),
    
    plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
  ) +
  guides(
    fill = guide_legend(
      keywidth = unit(6, "pt"),
      keyheight = unit(6, "pt"),
      ncol = 1,
      bycol = TRUE,
      label.position = "right",
      label.hjust = 0,
      override.aes = list(
        colour = NA,
        alpha = 0.95
      )
    )
  )

legend_grob <- cowplot::get_legend(legend_plot)

# =============================
# 10. 上下 panel 拼起来
# =============================
panels <- plot_grid(
  p_top_core,
  p_bottom_core,
  ncol = 1,
  align = "v",
  rel_heights = c(1, 1)
)

# =============================
# 11. 左侧物种标签
# =============================
species_labels <- ggdraw() +
  draw_label(
    "D.sim",
    x = 0.95, y = 0.75,
    hjust = 1,
    vjust = 0.5,
    size = 6,
    fontfamily = my_family,
    fontface = "italic"
  ) +
  draw_label(
    "D.mel",
    x = 0.95, y = 0.25,
    hjust = 1,
    vjust = 0.5,
    size = 6,
    fontfamily = my_family,
    fontface = "italic"
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA)
  )

# =============================
# 12. 左侧总 Y 轴标题
# =============================
y_title <- ggdraw() +
  draw_label(
    "Sum(RNA/DNA)",
    angle = 90,
    x = 0.5,
    y = 0.5,
    size = 7,
    fontfamily = my_family,
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA)
  )

# =============================
# 13. 底部总 X 轴标题
# =============================
x_title <- ggdraw() +
  draw_label(
    "Chromosome",
    x = 0.5,
    y = 0.5,
    size = 7,
    fontfamily = my_family
  ) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "transparent", colour = NA),
    panel.background = element_rect(fill = "transparent", colour = NA)
  )

# =============================
# 14. 主体拼接：Y标题 + 物种标签 + panel + legend
# =============================
main_row <- plot_grid(
  y_title,
  species_labels,
  panels,
  legend_grob,
  ncol = 4,
  rel_widths = c(0.06, 0.06, 1, 0.14),
  align = "h"
)

# =============================
# 15. 最终图：再加底部 X 标题
# =============================
final_plot <- plot_grid(
  main_row,
  x_title,
  ncol = 1,
  rel_heights = c(1, 0.06)
)

final_plot <- ggdraw(final_plot) +
  theme(
    plot.background = element_rect(fill = "transparent", colour = NA),
    plot.margin = margin(t = 4, r = 0, b = 0, l = 0)
  )

```

Save the plot

```{r}
#| echo: true
#| warning: false

final_canvas <- final_plot + ggview::canvas(90,60,units="mm",dpi=1200,bg="transparent")

final_canvas

#ggsave("Fig3C.pdf",final_plot,width = 90,height = 60,units="mm",dpi=900,bg = "transparent")
```

# Write some tables out

```{r}
#| echo: true
#| warning: false

new_table <- data_genome_normalized[, c("Gene", setdiff(names(data_genome_normalized),c("Chrom", "Start", "End", "Gene"))),with = FALSE]

# 写出
write.csv(new_table,file = "genome_coverage_matrix.csv",row.names = FALSE)

```

## transcriptome

```{r}
dt <- copy(data_transcriptome_normalized)

# 所有rep列名
rep_cols <- grep("_rep", names(dt), value = TRUE)

# 提取样本名（去掉 _repX）
sample_names <- unique(sub("_rep[0-9]+", "", rep_cols))

for (s in sample_names) {
  
  cols <- grep(paste0("^", s, "_rep"), names(dt), value = TRUE)
  
  dt[, (s) := rowMeans(.SD, na.rm = TRUE), .SDcols = cols]
}

mean_table <- dt[, c("gene_id", sample_names), with = FALSE]

fwrite(mean_table, "transcriptome.csv")

```

## comparison

```{r}
dt <- copy(comparison)

rep_cols <- grep("_rep", names(dt), value = TRUE)

sample_names <- unique(sub("_rep[0-9]+", "", rep_cols))

for (s in sample_names) {
  
  cols <- grep(paste0("^", s, "_rep"), names(dt), value = TRUE)
  
  dt[, (s) := rowMeans(.SD, na.rm = TRUE), .SDcols = cols]
}

mean_table <- dt[, c("chrom", "start", "end", "gene", sample_names),
                 with = FALSE]

mean_table <- dt[, c("chrom", "start", "end", "gene", sample_names),
                 with = FALSE]

fwrite(mean_table, "comparison_sample_means.csv")
```
