---
title: "Fig. S8 Dot Plot"
author: "Jie Hua"
date: 2026-03-09
lang: en
date-modified: 2026-03-09
keywords: ["Genomic Coverage","Transcriptome","Comparison","Dot plot"]
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: false
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(data.table)
library(stringr)
library(pbapply)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpointdensity)
library(scales)
library(grid)
library(patchwork)
```

# Load the files

## Define a folder

```{r}
#| echo: true
#| warning: false

#folder_path <- "H:\\HybridCells_Data\\Fig3_Dosage and Transcriptome"
folder_path <- "F:\\Big Paper\\Processed data\\Fig3_Dosage and Transcriptome"
```

## Load the orthofinder matrix generated with protein-protein.

```{r}
#| echo: true
#| warning: false

orthofinder <- fread(file=paste0(folder_path,"\\dmel_dsim_orthofinder\\Orthogroups.tsv"),header = T,sep = "\t",stringsAsFactors = FALSE,check.names = FALSE)
```

## Load the annotation GTF from IsoQuant

```{r}
#| echo: true
#| warning: false

isoquant_transcript_gtf <- paste0(folder_path,"\\transcriptome\\All.extended_annotation.gtf")

```

## Load the genomic coverage table

```{r}
#| echo: true
#| warning: false

genomic_coverage <- paste0(folder_path,"\\genomic\\genome_coverage_matrix.csv")
```

## Load the transcriptome table

```{r}
#| echo: true
#| warning: false

transcriptome_file <- paste0(folder_path,"\\transcriptome\\transcriptome.csv")
```

## Load the comparison table

```{r}
#| echo: true
#| warning: false

comparison_file <- paste0(folder_path,"\\transcriptome\\comparison_sample_means.csv")

```

# Proceed orthofinder result

Orthofinder found protein-protein level orthogroups. Convert it to transcript-transcript and gene-gene level.

## Convert the orthofinder result to transcript level

```{r}
#| echo: true
#| warning: false

cat("Convert Protein-Protein relationship to Transcript-Transcript level.")

orthofinder_transcript <- orthofinder

cols_to_clean <- colnames(orthofinder_transcript)[-1]

for(col in cols_to_clean){
  orthofinder_transcript[[col]] <- pbsapply(orthofinder_transcript[[col]], function(x){
    # If NA, just return NA.
    if(is.na(x) || trimws(x) == "") return(NA_character_)
    # Split the protein IDs.
    ids <- unlist(strsplit(x, ",\\s*"))
    # Remove the space
    ids <- trimws(ids)
    # Remove the ".p" postfix
    ids <- str_remove(ids, "\\.p\\d+$")
    # Keep the unique IDs
    ids <- unique(ids)
    # Put the IDs back
    paste(ids, collapse = ", ")
  })
}
```

## Convert the orthofinder result to gene level

Read the GTF file and prepare the mapping relationship.

```{r}
#| echo: true
#| warning: false

cat("The GTF file is",isoquant_transcript_gtf)

gtf <- data.table::fread(
  isoquant_transcript_gtf,
  sep = "\t",
  header = FALSE
)

colnames(gtf) <- c("chr","source","feature","start","end","score","strand","frame","attribute")

# Only keep transcript rows
gtf_transcript <- gtf %>%
  dplyr::filter(feature == "transcript")

# Extract transcript_id, gene_id, chr, start
gtf_map <- gtf_transcript[, .(
  transcript_id = sub('.*transcript_id "([^"]+)".*', '\\1', attribute),
  gene_id       = sub('.*gene_id "([^"]+)".*', '\\1', attribute),
  chr           = chr,
  start         = start
)]

# Build mapping
tx2gene <- setNames(gtf_map$gene_id, gtf_map$transcript_id)
tx2chr  <- setNames(gtf_map$chr, gtf_map$transcript_id)
```

Convert the transcript-transcript relationship to gene level.

```{r}
#| echo: true
#| warning: false

# Generic converter
convert_cell <- function(cell_value, map){

  if(is.na(cell_value) || trimws(cell_value) == "")
    return(NA_character_)

  # Split transcript list
  tx_ids <- unlist(strsplit(cell_value, ",\\s*"))
  tx_ids <- trimws(tx_ids)

  # Map values
  values <- map[tx_ids]
  values <- values[!is.na(values)]

  # Keep unique
  values <- unique(values)

  if(length(values) == 0)
    return(NA_character_)

  paste(values, collapse = ", ")
}


orthofinder_gene <- orthofinder_transcript
orthofinder_chr  <- orthofinder_transcript
orthofinder_location  <- orthofinder_transcript

cols_to_convert <- colnames(orthofinder_transcript)[-1]

for(col in cols_to_convert){

  cat("Processing column:", col, "\n")

  # transcript -> gene
  orthofinder_gene[[col]] <- pbsapply(
    orthofinder_transcript[[col]],
    convert_cell,
    map = tx2gene
  )

  # transcript -> chr
  orthofinder_chr[[col]] <- pbsapply(
    orthofinder_transcript[[col]],
    convert_cell,
    map = tx2chr
  )
}
```

Get the common chromosomes

```{r}
#| echo: true
#| warning: false

get_common_chr <- function(chr1, chr2){
  
  if(is.na(chr1) || is.na(chr2))
    return(NA_character_)
  # split multiple chromosomes
  c1 <- unlist(strsplit(chr1, ",\\s*"))
  c2 <- unlist(strsplit(chr2, ",\\s*"))
  # remove species suffix (_mel / _sim etc.)
  c1 <- sub("_.*$", "", c1)
  c2 <- sub("_.*$", "", c2)
  # intersection
  common <- intersect(c1, c2)
  if(length(common) == 0)
    return(NA_character_)
  paste(unique(common), collapse = ", ")
}

orthofinder_chr$common <- pbmapply(
  get_common_chr,
  orthofinder_chr[[2]],
  orthofinder_chr[[3]],
  SIMPLIFY = TRUE
)
```

# Save or load the RData

```{r}
#| echo: true
#| warning: false

save.image(file = "dotplot.RData")
load("dotplot.RData")
```

# Plot for genomic coverage

I already generated the **genome_coverage_matrix.csv** before.

```{r}
#| echo: true
#| warning: false

# Read the genomic coverage information
expr <- read.csv(genomic_coverage,stringsAsFactors=FALSE,check.names=FALSE)
samples <- colnames(expr)[-1]
n_samples <- length(samples)
draft <- data.frame(
  Orthogroup = orthofinder_gene$Orthogroup,
  check.names = FALSE
)

# Prepare a table
for (s in samples) {
  draft[[paste0(s, "_mel")]] <- orthofinder_gene$isoquant_split.mel.pep.nostar
  draft[[paste0(s, "_sim")]] <- orthofinder_gene$isoquant_split.sim.pep.nostar
}

# Define a function to calculate the sum level
get_expr_sum <- function(gene_string, sample_name) {
  if (is.na(gene_string) || trimws(gene_string) == "")
    return(NA_real_)
  genes <- unlist(strsplit(gene_string, ",\\s*"))
  genes <- trimws(genes)
  genes <- unique(genes)
  idx <- match(genes, expr$Gene)
  idx <- idx[!is.na(idx)]
  if (length(idx) == 0)
    return(NA_real_)
  sum(expr[idx, sample_name])
}

# Fill the values
for (s in samples) {
  # mel columns
  draft[[paste0(s, "_mel")]] <- pbsapply(
    draft[[paste0(s, "_mel")]],
    get_expr_sum,
    sample_name = s
  )
  # sim columns
  draft[[paste0(s, "_sim")]] <- pbsapply(
    draft[[paste0(s, "_sim")]],
    get_expr_sum,
    sample_name = s
  )
}

# Remove those orthogroups that only exist in one species
og_with_na <- orthofinder_gene$Orthogroup[
  !complete.cases(orthofinder_gene)
]
draft_clean <- draft[!draft$Orthogroup %in% og_with_na, ]

# Convert to long format
long_df <- draft_clean %>%
  pivot_longer(
    cols = -Orthogroup,
    names_to = c("Sample", "Species"),
    names_sep = "_",
    values_to = "Expression"
  ) %>%
  mutate(Expression = as.numeric(Expression)) %>%
  pivot_wider(
    names_from = Species,
    values_from = Expression,
    values_fn = sum
  )
long_df$mel[is.na(long_df$mel)] <- 0
long_df$sim[is.na(long_df$sim)] <- 0

# Plot
long_df <- subset(long_df, Sample %in% c("5A","6B","5B","7B","Ras3","ML82"))
max_val_log2 <- max(log2(c(long_df$mel, long_df$sim) + 1), na.rm = TRUE)
max_val_log2 <- max_val_log2 * 1.05
```

Add chromosome information

```{r}
#| echo: true
#| warning: false



# =========================
# 1. Prepare data
# =========================
long_df_chr <- merge(
  long_df,
  orthofinder_chr[, c("Orthogroup", "common")],
  by = "Orthogroup",
  all.x = TRUE
)

long_df_chr <- long_df_chr %>%
  filter(!(mel == 0 & sim == 0))

# Rename maps
sample_map <- c(
  "Ras3" = "Ras3",
  "ML82" = "ML82",
  "5A"   = "HYC1",
  "5B"   = "HYC2",
  "6B"   = "HYC3",
  "7B"   = "HYC4"
)

chr_map <- c(
  "chr2L" = "2L",
  "chr2R" = "2R",
  "chr3L" = "3L",
  "chr3R" = "3R",
  "chrX"  = "X",
  "chr4"  = "4"
)

sample_levels <- unname(sample_map)
chr_levels <- unname(chr_map)

long_df_chr <- long_df_chr %>%
  filter(!is.na(common)) %>%
  filter(Sample %in% names(sample_map)) %>%
  filter(common %in% names(chr_map)) %>%
  mutate(
    Sample = recode(Sample, !!!sample_map),
    common = recode(common, !!!chr_map)
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

# =========================
# 2. Classification
# =========================
signal_cutoff <- 0.2

long_df_chr <- long_df_chr %>%
  mutate(
    x_log2 = log2(mel + 1),
    y_log2 = log2(sim + 1),
    total_expr = mel + sim,
    mel_prop = if_else(total_expr > 0, mel / total_expr, NA_real_),
    sim_prop = if_else(total_expr > 0, sim / total_expr, NA_real_),
    change_class = case_when(
      mel_prop > 0.7 ~ "Dmel bias",
      mel_prop < 0.3 ~ "Dsim bias",
      TRUE           ~ "No bias"
    )
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

# =========================
# 3. Summary stats
# =========================
stat_df <- long_df_chr %>%
  group_by(Sample, common) %>%
  summarise(
    n_total = n(),
    n_no_bias = sum(change_class == "No bias", na.rm = TRUE),
    n_dmel    = sum(change_class == "Dmel bias", na.rm = TRUE),
    n_dsim    = sum(change_class == "Dsim bias", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    p_no_bias = scales::percent(n_no_bias / n_total, accuracy = 0.1),
    p_dmel    = scales::percent(n_dmel / n_total, accuracy = 0.1),
    p_dsim    = scales::percent(n_dsim / n_total, accuracy = 0.1)
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

max_val_log2 <- max(
  c(long_df_chr$x_log2, long_df_chr$y_log2),
  na.rm = TRUE
)

bar_df <- stat_df %>%
  select(Sample, common, n_total, n_no_bias, n_dmel, n_dsim) %>%
  pivot_longer(
    cols = c(n_no_bias, n_dmel, n_dsim),
    names_to = "change_class",
    values_to = "n"
  ) %>%
  mutate(
    change_class = recode(
      change_class,
      n_no_bias = "No bias",
      n_dmel    = "Dmel bias",
      n_dsim    = "Dsim bias"
    )
  ) %>%
  group_by(Sample, common) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels),
    change_class = factor(change_class, levels = c("No bias", "Dmel bias", "Dsim bias"))
  )

# =========================
# 4. Shared theme
# =========================
theme_pub_small <- function(my_family = "sans") {
  theme_bw(base_family = my_family) +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),

    panel.grid = element_blank(),
    panel.spacing.x = unit(0.4, "lines"),
    panel.spacing.y = unit(0.4, "lines"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.1),

    strip.placement = "outside",
    strip.text.x = element_text(
      angle = 0, hjust = 0.5, size = 6,
      margin = margin(t = 0, r = 0, b = 2, l = 0),
      color = "black", family = "sans"
    ),
    strip.text.y.left = element_text(
      angle = 0, size = 6,
      color = "black", family = "sans"
    ),

    axis.text.x = element_text(size = 5, color = "black", family = "sans",angle = 45,hjust = 1,vjust = 1),
    axis.text.y = element_text(size = 5, color = "black", family = "sans"),
    axis.title.x = element_text(size = 7, color = "black", family = "sans"),
    axis.title.y = element_text(size = 7, color = "black", family = "sans"),

    axis.ticks = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.length = unit(1.5, "pt"),
    axis.minor.ticks.x = element_line(linewidth = 0.1, color = "black"),
    axis.minor.ticks.y = element_line(linewidth = 0.1, color = "black"),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),

    legend.position = "right",
    legend.margin = margin(t = 0, r = 0, b = 0, l = -2),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2),
    legend.title = element_text(size = 6, color = "black", family = "sans"),
    legend.text  = element_text(size = 6, color = "black", family = "sans"),
    legend.key.size = unit(6, "pt"),
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),
    legend.key.spacing.x = unit(1, "pt"),
    legend.key.spacing.y = unit(1, "pt"),
    legend.spacing.x = unit(1, "pt"),
    legend.spacing.y = unit(1, "pt"),
    legend.box.spacing = unit(2, "pt"),

    strip.clip = "off",
    plot.margin = margin(t = 0, r = 0, b = 0, l = 0)
  )
}

# =========================
# 5. Left plot: scatter
# =========================
p_genomic_chr_scatter <- ggplot(
  long_df_chr,
  aes(x = x_log2, y = y_log2)
) +
  ggpointdensity::geom_pointdensity(size = 0.05) +
  scale_color_viridis_c(
    option = "plasma",
    name = "Point density",
    labels = scales::label_scientific(digits = 2)
  ) +
  geom_abline(
    slope = 1, intercept = 0,
    linetype = "dashed",
    linewidth = 0.2,
    color = "grey40"
  ) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = c(0, 5, 10),
    limits = c(0, 10),
    minor_breaks = seq(0, max_val_log2, by = 2.5),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_y_continuous(
    expand = c(0,0),
    breaks = c(0, 5, 10),
    limits = c(0, 10),
    minor_breaks = seq(0, max_val_log2, by = 2.5),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  facet_grid(
    Sample ~ common,
    switch = "y"
  ) +
  coord_fixed(ratio = 1) +
  labs(
    x = expression(log[2](Dmel + 1)),
    y = expression(log[2](Dsim + 1))
  ) +
  theme_pub_small()

# =========================
# 6. Right plot: stacked bar
# =========================
p_genomic_chr_bar <- ggplot(
  bar_df,
  aes(x = 1, y = prop, fill = change_class)
) +
  geom_col(width = 0.75, color = "black", linewidth = 0.1) +
  facet_grid(
    Sample ~ common,
    switch = "y"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = c(0, 0),
    breaks = c(0, 0.5, 1.0),
    minor_breaks = seq(0, 1, by = 0.25),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = NULL,
    minor_breaks = NULL
  ) +
  scale_fill_manual(
    values = c(
      "No bias"   = "grey70",
      "Dmel bias" = "#1f78b4",
      "Dsim bias" = "#d95f02"
    ),
    name = NULL
  ) +
  labs(
    x = NULL,
    y = "Proportion"
  ) +
  theme_pub_small() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    aspect.ratio = 1
  )

# =========================
# 7. Combine
# =========================
p_genomic_chr_combined <- p_genomic_chr_scatter + p_genomic_chr_bar +
  plot_layout(widths = c(1, 1)) &
  theme(
    legend.position = "right",

    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA)
  )
```

Save plots

```{r}
#| echo: true
#| warning: false

p_genomic_chr_combined_canvas <- p_genomic_chr_combined + ggview::canvas(180,55, units = "mm", dpi = 600, bg = "transparent")
p_genomic_chr_combined_canvas

# ggsave("FigS8A.pdf",p_genomic_chr_combined, width=180,height=55,units="mm",dpi=900, bg = "transparent")
```

## Color with the location of genes

```{r}
#| echo: true
#| warning: false

# =========================================================
# 1) 从 GTF 提取 gene -> chr(start)（只保留 mel genome）
# =========================================================
gtf_gene <- gtf %>%
  filter(feature == "gene") %>%
  mutate(
    gene_id = sub('.*gene_id "([^"]+)".*', '\\1', attribute)
  ) %>%
  select(gene_id, chr, start)

# 只保留 mel 染色体，并去掉后缀 _mel
gtf_mel <- gtf_gene %>%
  filter(grepl("_mel$", chr)) %>%
  mutate(chr = sub("_mel$", "", chr))

# =========================================================
# 2) Orthogroup -> mel gene location（取最小 start 作为代表）
# =========================================================
get_location <- function(gene_string){
  if(is.na(gene_string) || trimws(gene_string) == "")
    return(data.frame(chr = NA_character_, start = NA_real_))

  genes <- unlist(strsplit(gene_string, ",\\s*"))
  genes <- trimws(genes)

  loc <- gtf_mel %>% filter(gene_id %in% genes)
  if(nrow(loc) == 0)
    return(data.frame(chr = NA_character_, start = NA_real_))

  loc <- loc[which.min(loc$start), ]
  data.frame(chr = loc$chr, start = loc$start)
}

loc_list <- pblapply(
  orthofinder_gene$isoquant_split.mel.pep.nostar,
  get_location
)

loc_df <- do.call(rbind, loc_list)

orthofinder_location <- data.frame(
  Orthogroup = orthofinder_gene$Orthogroup,
  chr   = loc_df$chr,
  start = loc_df$start,
  stringsAsFactors = FALSE
)

# =========================================================
# 3) expr 转成长格式
# =========================================================
expr_location <- expr %>%
  rename(gene = Gene) %>%
  pivot_longer(
    cols = -gene,
    names_to = "Sample",
    values_to = "expression"
  )

# =========================================================
# 4) Orthogroup expression（mel vs sim）
# =========================================================
# mel genes
mel_map <- orthofinder_gene %>%
  select(Orthogroup, isoquant_split.mel.pep.nostar) %>%
  rename(gene_list = isoquant_split.mel.pep.nostar)

mel_long <- mel_map %>%
  separate_rows(gene_list, sep = ",\\s*") %>%
  rename(gene = gene_list)

mel_expr <- merge(mel_long, expr_location, by = "gene") %>%
  group_by(Orthogroup, Sample) %>%
  summarise(mel = sum(expression, na.rm = TRUE), .groups = "drop")

# sim genes
sim_map <- orthofinder_gene %>%
  select(Orthogroup, isoquant_split.sim.pep.nostar) %>%
  rename(gene_list = isoquant_split.sim.pep.nostar)

sim_long <- sim_map %>%
  separate_rows(gene_list, sep = ",\\s*") %>%
  rename(gene = gene_list)

sim_expr <- merge(sim_long, expr_location, by = "gene") %>%
  group_by(Orthogroup, Sample) %>%
  summarise(sim = sum(expression, na.rm = TRUE), .groups = "drop")

# mel + sim
long_df <- merge(
  mel_expr,
  sim_expr,
  by = c("Orthogroup", "Sample"),
  all = TRUE
)

# =========================================================
# 5) 合并 genomic location
# =========================================================
long_df_loc <- merge(
  long_df,
  orthofinder_location,
  by = "Orthogroup",
  all.x = TRUE
)

long_df_loc <- subset(long_df_loc, !is.na(chr))

# 只保留主要染色体
chromosomes <- c("chr3R")
long_df_loc <- subset(long_df_loc, chr %in% chromosomes)

long_df_loc <- subset(long_df_loc,
                   Sample %in% c("6B"))

# =========================================================
# 新增：start 分组
# =========================================================
long_df_loc$region <- cut(
  long_df_loc$start,
  breaks = c(-Inf, 6.085e6, 20.08e6, Inf),
  labels = c("start < 6.085M",
             "6.085M – 20.08M",
             "start > 20.08M")
)

# =========================================================
# 6) 计算绘图范围
# =========================================================
max_val_log2 <- max(
  log2(long_df_loc$mel + 1),
  log2(long_df_loc$sim + 1),
  na.rm = TRUE
)

# =========================================================
# 7) 作图（普通点，按 start 上色，Sample × chr 分面）
# =========================================================
p_genomic_location <- ggplot(
  long_df_loc,
  aes(
    x = log2(mel + 1),
    y = log2(sim + 1),
    color = region
  )
) +
  geom_point(size = 3, alpha = 1) +

  scale_color_manual(
    values = c(
      "start < 6.085M" = "#440154",
      "6.085M – 20.08M" = "#21908C",
      "start > 20.08M" = "#FDE725"
    ),
    name = "chr3R region"
  ) +

  geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed",
    linewidth = 0.5,
    color = "black"
  ) +

  coord_fixed(
    ratio = 1,
    xlim = c(0, max_val_log2),
    ylim = c(0, max_val_log2)
  ) +

  labs(
    x = expression(log[2](Dmel + 1)),
    y = expression(log[2](Dsim + 1)),
    title = "Genomic Coverage (chr3R, Sample 6B)"
  ) +

  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 20),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing = unit(0.4, "lines"),
    strip.text = element_text(size = 16,family = "sans",color = "black"),
    legend.position = "right",
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent")
  )

```

```{r}
#| echo: true
#| warning: false

#p_genomic_location_canvas <- p_genomic_location + ggview::canvas(10,10,units="in",dpi=600,bg="white")
#ggsave("p_genomic_location_canvas_6b_3R.png",p_genomic_location, width=10,height=10,units="in",dpi=900)
```

# Plot for transcriptome

I already generated the **transcriptome.csv** before.

```{r}
#| echo: true
#| warning: false

# Read the genomic coverage information
expr2 <- read.csv(transcriptome_file,sep = ",",stringsAsFactors=FALSE,check.names=FALSE)
samples <- colnames(expr2)[-1]
n_samples <- length(samples)
draft2 <- data.frame(
  Orthogroup = orthofinder_gene$Orthogroup,
  check.names = FALSE
)

# Prepare a table
for (s in samples) {
  draft2[[paste0(s, "_mel")]] <- orthofinder_gene$isoquant_split.mel.pep.nostar
  draft2[[paste0(s, "_sim")]] <- orthofinder_gene$isoquant_split.sim.pep.nostar
}

# Define a function to calculate the sum level
get_expr_sum2 <- function(gene_string, sample_name) {
  if (is.na(gene_string) || trimws(gene_string) == "")
    return(NA_real_)
  genes <- unlist(strsplit(gene_string, ",\\s*"))
  genes <- trimws(genes)
  genes <- unique(genes)
  idx <- match(genes, expr2$gene_id)
  idx <- idx[!is.na(idx)]
  if (length(idx) == 0)
    return(NA_real_)
  sum(expr2[idx, sample_name])
}

# Fill the values
for (s in samples) {
  draft2[[paste0(s, "_mel")]] <- pbsapply(
    draft2[[paste0(s, "_mel")]],
    get_expr_sum2,
    sample_name = s
  )
  draft2[[paste0(s, "_sim")]] <- pbsapply(
    draft2[[paste0(s, "_sim")]],
    get_expr_sum2,
    sample_name = s
  )
}

# Remove those orthogroups that only exist in one species
og_with_na <- orthofinder_transcript$Orthogroup[
  !complete.cases(orthofinder_transcript)
]
draft2_clean <- draft2[!draft$Orthogroup %in% og_with_na, ]

# Convert to long format
long_df2 <- draft2_clean %>%
  pivot_longer(
    cols = -Orthogroup,
    names_to = c("Sample", "Species"),
    names_sep = "_",
    values_to = "Expression"
  ) %>%
  mutate(Expression = as.numeric(Expression)) %>%
  pivot_wider(
    names_from = Species,
    values_from = Expression,
    values_fn = sum
  )
long_df2$mel[is.na(long_df2$mel)] <- 0
long_df2$sim[is.na(long_df2$sim)] <- 0
long_df2 <- subset(long_df2,
                   Sample %in% c("5A","6B","5B","7B","Ras3","ML82"))
max_val_log2 <- max(log2(c(long_df2$mel,
                           long_df2$sim) + 1),
                    na.rm = TRUE)

max_val_log2 <- max_val_log2 * 1.05

```

Add chromosome information

```{r}
#| echo: true
#| warning: false
# =========================
# 1. Prepare data
# =========================
long_df_chr2 <- merge(
  long_df2,
  orthofinder_chr[, c("Orthogroup", "common")],
  by = "Orthogroup",
  all.x = TRUE
)


long_df_chr2 <- long_df_chr2 %>%
  filter(!(mel == 0 & sim == 0))

# Rename maps
sample_map <- c(
  "Ras3" = "Ras3",
  "ML82" = "ML82",
  "5A"   = "HYC1",
  "5B"   = "HYC2",
  "6B"   = "HYC3",
  "7B"   = "HYC4"
)

chr_map <- c(
  "chr2L" = "2L",
  "chr2R" = "2R",
  "chr3L" = "3L",
  "chr3R" = "3R",
  "chrX"  = "X",
  "chr4"  = "4"
)

sample_levels <- unname(sample_map)
chr_levels <- unname(chr_map)

long_df_chr2 <- long_df_chr2 %>%
  filter(!is.na(common)) %>%
  filter(Sample %in% names(sample_map)) %>%
  filter(common %in% names(chr_map)) %>%
  mutate(
    Sample = recode(Sample, !!!sample_map),
    common = recode(common, !!!chr_map)
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

# =========================
# 2. Classification
# =========================
signal_cutoff <- 0.2

long_df_chr2 <- long_df_chr2 %>%
  mutate(
    x_log2 = log2(mel + 1),
    y_log2 = log2(sim + 1),
    total_expr = mel + sim,
    mel_prop = if_else(total_expr > 0, mel / total_expr, NA_real_),
    sim_prop = if_else(total_expr > 0, sim / total_expr, NA_real_),
    change_class = case_when(
      mel_prop > 0.7 ~ "Dmel bias",
      mel_prop < 0.3 ~ "Dsim bias",
      TRUE           ~ "No bias"
    )
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

# =========================
# 3. Summary stats
# =========================
stat_df2 <- long_df_chr2 %>%
  group_by(Sample, common) %>%
  summarise(
    n_total = n(),
    n_no_bias = sum(change_class == "No bias", na.rm = TRUE),
    n_dmel    = sum(change_class == "Dmel bias", na.rm = TRUE),
    n_dsim    = sum(change_class == "Dsim bias", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    p_no_bias = scales::percent(n_no_bias / n_total, accuracy = 0.1),
    p_dmel    = scales::percent(n_dmel / n_total, accuracy = 0.1),
    p_dsim    = scales::percent(n_dsim / n_total, accuracy = 0.1)
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

max_val_log2 <- max(
  c(long_df_chr2$x_log2, long_df_chr2$y_log2),
  na.rm = TRUE
)

bar_df2 <- stat_df2 %>%
  select(Sample, common, n_total, n_no_bias, n_dmel, n_dsim) %>%
  pivot_longer(
    cols = c(n_no_bias, n_dmel, n_dsim),
    names_to = "change_class",
    values_to = "n"
  ) %>%
  mutate(
    change_class = recode(
      change_class,
      n_no_bias = "No bias",
      n_dmel    = "Dmel bias",
      n_dsim    = "Dsim bias"
    )
  ) %>%
  group_by(Sample, common) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels),
    change_class = factor(change_class, levels = c("No bias", "Dmel bias", "Dsim bias"))
  )

# =========================
# 4. Shared theme
# =========================

# =========================
# 5. Left plot: scatter
# =========================
p_transcriptome_chr_scatter <- ggplot(
  long_df_chr2,
  aes(x = x_log2, y = y_log2)
) +
  ggpointdensity::geom_pointdensity(size = 0.01) +
  scale_color_viridis_c(
    option = "plasma",
    name = "Point density",
    labels = scales::label_scientific(digits = 2)
  ) +
  geom_abline(
    slope = 1, intercept = 0,
    linetype = "dashed",
    linewidth = 0.2,
    color = "grey40"
  ) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = c(0, 10, 20),
    limits = c(0, 20),
    minor_breaks = seq(0, 20, by = 5),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_y_continuous(
    expand = c(0,0),
    breaks = c(0, 10, 20),
    limits = c(0, 20),
    minor_breaks = seq(0,20, by = 5),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  facet_grid(
    Sample ~ common,
    switch = "y"
  ) +
  coord_fixed(ratio = 1) +
  labs(
    x = expression(log[2](Dmel + 1)),
    y = expression(log[2](Dsim + 1))
  ) +
  theme_pub_small()

# =========================
# 6. Right plot: stacked bar
# =========================
p_transcriptome_chr_bar <- ggplot(
  bar_df2,
  aes(x = 1, y = prop, fill = change_class)
) +
  geom_col(width = 0.75, color = "black", linewidth = 0.1) +
  facet_grid(
    Sample ~ common,
    switch = "y"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = c(0, 0),
    breaks = c(0, 0.5, 1.0),
    minor_breaks = seq(0, 1, by = 0.25),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = NULL,
    minor_breaks = NULL
  ) +
  scale_fill_manual(
    values = c(
      "No bias"   = "grey70",
      "Dmel bias" = "#1f78b4",
      "Dsim bias" = "#d95f02"
    ),
    name = NULL
  ) +
  labs(
    x = NULL,
    y = "Proportion"
  ) +
  theme_pub_small() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    aspect.ratio = 1
  )

# =========================
# 7. Combine
# =========================
p_transcriptome_chr_combined <- p_transcriptome_chr_scatter + p_transcriptome_chr_bar +
  plot_layout(widths = c(1, 1)) &
  theme(
    legend.position = "right",

    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA)
  )
```

Save the plots

```{r}
#| echo: true
#| warning: false

p_transcriptome_chr_combined_canvas <- p_transcriptome_chr_combined + ggview::canvas(180,55, units = "mm", dpi = 600, bg = "transparent")
p_transcriptome_chr_combined_canvas

# ggsave("FigS8B.pdf",p_transcriptome_chr_combined, width=180,height=55,units="mm",dpi=900, bg = "transparent")
```

# Plot for relative expression

I already generated the **comparison_sample_means.csv** before.

```{r}
#| echo: true
#| warning: false

# Read the genomic coverage information
expr3 <- read.csv(comparison_file,sep = ",",stringsAsFactors=FALSE,check.names=FALSE)
samples <- colnames(expr3 )[5:ncol(expr3)]
n_samples <- length(samples)
draft3 <- data.frame(
  Orthogroup = orthofinder_gene$Orthogroup,
  check.names = FALSE
)

# Prepare a table
for (s in samples) {
  draft3[[paste0(s, "_mel")]] <- orthofinder_gene$isoquant_split.mel.pep.nostar
  draft3[[paste0(s, "_sim")]] <- orthofinder_gene$isoquant_split.sim.pep.nostar
}

# Define a function to calculate the sum level
get_expr_sum3 <- function(gene_string, sample_name) {
  if (is.na(gene_string) || trimws(gene_string) == "")
    return(NA_real_)
  genes <- unlist(strsplit(gene_string, ",\\s*"))
  genes <- trimws(genes)
  genes <- unique(genes)
  idx <- match(genes, expr3$gene)
  idx <- idx[!is.na(idx)]
  if (length(idx) == 0)
    return(NA_real_)
  sum(expr3[idx, sample_name])
}

# Fill the values
for (s in samples) {
  draft3[[paste0(s, "_mel")]] <- pbsapply(
    draft3[[paste0(s, "_mel")]],
    get_expr_sum3,
    sample_name = s
  )
  draft3[[paste0(s, "_sim")]] <- pbsapply(
    draft3[[paste0(s, "_sim")]],
    get_expr_sum3,
    sample_name = s
  )
}

# Remove those orthogroups that only exist in one species
og_with_na <- orthofinder_gene$Orthogroup[
  !complete.cases(orthofinder_gene)
]
draft3_clean <- draft3[!draft$Orthogroup %in% og_with_na, ]

# Convert to long format


long_df3 <- draft3_clean %>%
  pivot_longer(
    cols = -Orthogroup,
    names_to = c("Sample", "Species"),
    names_sep = "_",
    values_to = "Expression"
  ) %>%
  mutate(Expression = as.numeric(Expression)) %>%
  pivot_wider(
    names_from = Species,
    values_from = Expression,
    values_fn = sum
  )
long_df3$mel[is.na(long_df3$mel)] <- 0
long_df3$sim[is.na(long_df3$sim)] <- 0
long_df3 <- subset(long_df3,Sample %in% c("5A","6B","5B","7B","Ras3","ML82"))
max_val_log2 <- max(log2(c(long_df3$mel,
                           long_df3$sim) + 1),
                    na.rm = TRUE)

max_val_log2 <- max_val_log2 * 1.05
```

Add chromosome information

```{r}
#| echo: true
#| warning: false

# =========================
# 1. Prepare data
# =========================
long_df_chr3 <- merge(
  long_df3,
  orthofinder_chr[, c("Orthogroup", "common")],
  by = "Orthogroup",
  all.x = TRUE
)


long_df_chr3 <- long_df_chr3 %>%
  filter(!(mel == 0 & sim == 0))

# Rename maps
sample_map <- c(
  "Ras3" = "Ras3",
  "ML82" = "ML82",
  "5A"   = "HYC1",
  "5B"   = "HYC2",
  "6B"   = "HYC3",
  "7B"   = "HYC4"
)

chr_map <- c(
  "chr2L" = "2L",
  "chr2R" = "2R",
  "chr3L" = "3L",
  "chr3R" = "3R",
  "chrX"  = "X",
  "chr4"  = "4"
)

sample_levels <- unname(sample_map)
chr_levels <- unname(chr_map)

long_df_chr3 <- long_df_chr3 %>%
  filter(!is.na(common)) %>%
  filter(Sample %in% names(sample_map)) %>%
  filter(common %in% names(chr_map)) %>%
  mutate(
    Sample = recode(Sample, !!!sample_map),
    common = recode(common, !!!chr_map)
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

# =========================
# 2. Classification
# =========================
signal_cutoff <- 0.2

long_df_chr3 <- long_df_chr3 %>%
  mutate(
    x_log2 = log2(mel + 1),
    y_log2 = log2(sim + 1),
    total_expr = mel + sim,
    mel_prop = if_else(total_expr > 0, mel / total_expr, NA_real_),
    sim_prop = if_else(total_expr > 0, sim / total_expr, NA_real_),
    change_class = case_when(
      mel_prop > 0.7 ~ "Dmel bias",
      mel_prop < 0.3 ~ "Dsim bias",
      TRUE           ~ "No bias"
    )
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

# =========================
# 3. Summary stats
# =========================
stat_df3 <- long_df_chr3 %>%
  group_by(Sample, common) %>%
  summarise(
    n_total = n(),
    n_no_bias = sum(change_class == "No bias", na.rm = TRUE),
    n_dmel    = sum(change_class == "Dmel bias", na.rm = TRUE),
    n_dsim    = sum(change_class == "Dsim bias", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    p_no_bias = scales::percent(n_no_bias / n_total, accuracy = 0.1),
    p_dmel    = scales::percent(n_dmel / n_total, accuracy = 0.1),
    p_dsim    = scales::percent(n_dsim / n_total, accuracy = 0.1)
  ) %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels)
  )

max_val_log2 <- max(
  c(long_df_chr3$x_log2, long_df_chr3$y_log2),
  na.rm = TRUE
)

bar_df3 <- stat_df3 %>%
  select(Sample, common, n_total, n_no_bias, n_dmel, n_dsim) %>%
  pivot_longer(
    cols = c(n_no_bias, n_dmel, n_dsim),
    names_to = "change_class",
    values_to = "n"
  ) %>%
  mutate(
    change_class = recode(
      change_class,
      n_no_bias = "No bias",
      n_dmel    = "Dmel bias",
      n_dsim    = "Dsim bias"
    )
  ) %>%
  group_by(Sample, common) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    Sample = factor(Sample, levels = sample_levels),
    common = factor(common, levels = chr_levels),
    change_class = factor(change_class, levels = c("No bias", "Dmel bias", "Dsim bias"))
  )

# =========================
# 4. Shared theme
# =========================


# =========================
# 5. Left plot: scatter
# =========================
p_comparison_chr_scatter <- ggplot(
  long_df_chr3,
  aes(x = x_log2, y = y_log2)
) +
  ggpointdensity::geom_pointdensity(size = 0.01) +
  scale_color_viridis_c(
    option = "plasma",
    name = "Point density",
    labels = scales::label_scientific(digits = 2)
  ) +
  geom_abline(
    slope = 1, intercept = 0,
    linetype = "dashed",
    linewidth = 0.2,
    color = "grey40"
  ) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = c(0, 10, 20),
    limits = c(0, 20),
    minor_breaks = seq(0, 20, by = 5),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_y_continuous(
    expand = c(0,0),
    breaks = c(0, 10, 20),
    limits = c(0, 20),
    minor_breaks = seq(0,20, by = 5),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  facet_grid(
    Sample ~ common,
    switch = "y"
  ) +
  coord_fixed(ratio = 1) +
  labs(
    x = expression(log[2](Dmel + 1)),
    y = expression(log[2](Dsim + 1))
  ) +
  theme_pub_small()

# =========================
# 6. Right plot: stacked bar
# =========================
p_comparison_chr_bar <- ggplot(
  bar_df3,
  aes(x = 1, y = prop, fill = change_class)
) +
  geom_col(width = 0.75, color = "black", linewidth = 0.2) +
  facet_grid(
    Sample ~ common,
    switch = "y"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, 1),
    expand = c(0, 0),
    breaks = c(0, 0.5, 1.0),
    minor_breaks = seq(0, 1, by = 0.25),
    guide = guide_axis(minor.ticks = TRUE)
  ) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = NULL,
    minor_breaks = NULL
  ) +
  scale_fill_manual(
    values = c(
      "No bias"   = "grey70",
      "Dmel bias" = "#1f78b4",
      "Dsim bias" = "#d95f02"
    ),
    name = NULL
  ) +
  labs(
    x = NULL,
    y = "Proportion"
  ) +
  theme_pub_small() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    aspect.ratio = 1
  )

# =========================
# 7. Combine
# =========================
p_comparison_chr_combined <- p_comparison_chr_scatter + p_comparison_chr_bar +
  plot_layout(widths = c(1, 1)) &
  theme(
    legend.position = "right",

    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background  = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA)
  )
```

Save the plots

```{r}
#| echo: true
#| warning: false

p_comparison_chr_combined_canvas <- p_comparison_chr_combined + ggview::canvas(180,55, units = "mm", dpi = 600, bg = "transparent")
p_comparison_chr_combined_canvas

# ggsave("FigS8C.pdf",p_comparison_chr_combined, width=180,height=55,units="mm",dpi=900, bg = "transparent")
```
