---
title: "FigS3 S5. Genome Ploidy"
author: "Jie Hua"
date: 2025-09-30
lang: en
date-modified: 2026-04-25
keywords: ["Genome Ploidy","Hybrid Cells"]
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: false
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggdist)
library(RColorBrewer)
library(viridis)
library(scales)
```

# Load *.bed* files

Define the folder and find all the files

```{r}
#| echo: true
#| warning: false

# Define the folder containing the bedgraph files
#folder <- "F:\\Big Paper\\Processed data\\Fig3_Genomic_coverage\\100bp_window"

# Lab PC
folder <- "H:\\HybridCells_Data\\Fig2_Genomic_coverage\\100bp_window"

cat("The folder containing the bedgraph files is", folder, "\n")

# List the bedGraph files
bedgraph_files <- list.files(folder, pattern = "\\.bedGraph$", full.names = TRUE)
cat("The bedGraph files (total number:", length(bedgraph_files), "):\n")
print(bedgraph_files)
```

Read all the files into a list, rename them into customized names.

```{r}
#| echo: true
#| warning: false


name_map <- c(
  "05a_q30_coverage_100bp.bedGraph" = "5A",
  "05b_barcode17_q30_coverage_100bp.bedGraph" = "5B_17",
  "05b_barcode18_q30_coverage_100bp.bedGraph" = "5B_18",
  "05b_barcode19_q30_coverage_100bp.bedGraph" = "5B_19",
  "05b_barcode20_q30_coverage_100bp.bedGraph" = "5B_20",
  "06b_q30_coverage_100bp.bedGraph" = "6B",
  "07b_barcode21_q30_coverage_100bp.bedGraph" = "7B_21",
  "07b_barcode22_q30_coverage_100bp.bedGraph" = "7B_22",
  "07b_barcode23_q30_coverage_100bp.bedGraph" = "7B_23",
  "07b_barcode24_q30_coverage_100bp.bedGraph" = "7B_24",
  "ML82_q30_coverage_100bp.bedGraph" = "ML82",
  "Ras3_q30_coverage_100bp.bedGraph" = "Ras3"
)

data_list <- list()
for (file in bedgraph_files) {
  data <- read.table(file, header = FALSE, sep = "\t", stringsAsFactors = FALSE)
  colnames(data) <- c("Chromosome", "Start", "End", "Counts")

  data$total <- (data$End - data$Start) * data$Counts
  
  # Get the sample name from name_map
  file_base <- basename(file)
  sample_name <- name_map[file_base]
  data$Sample <- sample_name
  
  # Save into the list
  data_list[[sample_name]] <- data
}

cat("The names of the list containing all the bed tables:\n")
print(names(data_list))

```

Merge those tables into one table.

```{r}
#| echo: true
#| warning: false

merged_df <- bind_rows(data_list)
```

# Plot the distribution of coverage

```{r}
#| echo: true
#| warning: false

# Selected main chromosomes
selected_chromosomes <- c(
  "chrX_mel","chr2L_mel","chr2R_mel","chr3L_mel","chr3R_mel","chr4_mel",
  "chrX_sim","chr2L_sim","chr2R_sim","chr3L_sim","chr3R_sim","chr4_sim")

# Samples to be kept
selected_samples <- c("Ras3","ML82","5A","5B_19","6B","7B_23")

# Rename the samples）
sample_rename_map <- c(
  "Ras3"="Ras3",
  "ML82"="ML82",
  "5A" = "HYC1",
  "5B_19" = "HYC2",
  "6B" = "HYC3",
  "7B_23" = "HYC4"
)

chromosome_rename_map <- c(
  "chrX_mel" = "X_mel",
  "chr2L_mel" = "2L_mel",
  "chr2R_mel" = "2R_mel",
  "chr3L_mel" = "3L_mel",
  "chr3R_mel" = "3R_mel",
  "chr4_mel" = "4_mel",
  "chrX_sim" = "X_sim",
  "chr2L_sim" = "2L_sim",
  "chr2R_sim" = "2R_sim",
  "chr3L_sim" = "3L_sim",
  "chr3R_sim" = "3R_sim",
  "chr4_sim" = "4_sim"
)


species_colors <- c(
  "Dmel" = "#61649F",
  "Dsim" = "#ED556A"
)

# Filter the chromosomes for Dmel and Dsim control
merged_df <- merged_df %>%
  filter(
    !(
      (str_detect(Chromosome, "_mel$") & Sample == "ML82") |
      (str_detect(Chromosome, "_sim$") & Sample == "Ras3")
    )
  )
```

## Function1: density plot

I decided to give up the density plot, it looks not beautiful.

```{r}
#| echo: true
#| warning: false

df_plot <- merged_df %>%
  filter(
    Chromosome %in% selected_chromosomes,
    Sample %in% selected_samples
  ) %>%
  mutate(
    # Rename chromosome
    Chromosome = recode(Chromosome, !!!chromosome_rename_map),

    # Order
    Chromosome = factor(
      Chromosome,
      levels = chromosome_rename_map[selected_chromosomes]
    ),

    Sample = recode(Sample, !!!sample_rename_map),
    Sample = factor(Sample, levels = sample_rename_map[selected_samples])
  ) %>%
  group_by(Sample, Chromosome)



# Give mel and sim chromosomes different colors
mel_chroms <- grep("_mel$", selected_chromosomes, value = TRUE)
sim_chroms <- grep("_sim$", selected_chromosomes, value = TRUE)
mel_colors <- brewer.pal(n = length(mel_chroms), name = "Blues")
sim_colors <- brewer.pal(n = length(sim_chroms), name = "Oranges")

# Customize colors
chrom_colors <- c(mel_colors, sim_colors)
names(chrom_colors) <- c(mel_chroms, sim_chroms)

# Plot the density plot
p1 <- ggplot(df_plot, aes(x = Counts, color = Chromosome, weight = End - Start)) +
  geom_freqpoly(binwidth = 1, linewidth = 1) +
  facet_wrap(~Sample, scales = "free_y") +
  scale_color_viridis_d(option = "D", end = 0.9) +
  labs(
    title = "",
    x = "Counts",
    y = "Density"
  ) +
  coord_cartesian(xlim = c(0, 75)) +
  theme_minimal(base_family = "sans") +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black")
  )


# ggdist version
df_plot2 <- merged_df %>%
  filter(
    Chromosome %in% selected_chromosomes,
    Sample %in% selected_samples
  ) %>%
  mutate(
    species = ifelse(grepl("_mel$", Chromosome), "Dmel", "Dsim"),

    Chromosome = recode(Chromosome, !!!chromosome_rename_map),
    Chromosome = factor(
      Chromosome,
      levels = chromosome_rename_map[selected_chromosomes]
    ),

    Sample = recode(Sample, !!!sample_rename_map),
    Sample = factor(Sample, levels = sample_rename_map[selected_samples])
  ) %>%
  group_by(Sample, Chromosome)


p1_new <- ggplot(
  df_plot2,
  aes(
    x = Counts,
    y = Chromosome,
    fill = species,
    weight = End - Start
  )
) +
  stat_slab(
    alpha = 1,
    scale = 0.7,
    normalize = "groups",
    side = "top",
    color = NA,
    adjust=1
  ) +
  facet_wrap(~Sample, scales = "free",ncol=6) +
  scale_fill_manual(values = species_colors, name = "Chromosome") +
  scale_x_continuous(
  expand = c(0.05, 0.05),
  breaks = function(x) {
    seq(
      from = floor(min(x) / 20) * 20,
      to   = ceiling(max(x) / 20) * 20,
      by = 20
    )
  },
  minor_breaks = function(x) {
    seq(
      from = floor(min(x) / 5) * 5,
      to   = ceiling(max(x) / 5) * 5,
      by = 5
    )
  },
  guide = guide_axis(minor.ticks = TRUE)
) +
  
  labs(
    x = "Coverage",
    y = "Chromosome"
  ) +
  coord_cartesian(xlim = c(0, 55)) +
  theme_bw(base_family = "sans") +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.box.background = element_rect(fill = "transparent", color = NA),
    
    
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.3),
    
    legend.key = element_rect(fill = "transparent", color = NA),
    strip.background = element_rect(fill = "transparent", color = NA),
    
    strip.text.x = element_text(angle = 0, hjust = 0.5, size = 6), 
    
    # 坐标文字
    axis.text.x = element_text(size = 5, color = "black"),
    axis.text.y = element_text(size = 5, color = "black"),
    axis.title.x = element_text(size = 7, color = "black"),
    axis.title.y = element_text(size = 7, color = "black"),
    
    # 主刻度
    axis.ticks.x = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.y = element_line(linewidth = 0.2, color = "black"),
    axis.ticks.length = unit(1.5, "pt"),

    # 小刻度
    axis.minor.ticks.x = element_line(linewidth = 0.1, color = "black"),
    axis.minor.ticks.y = element_line(linewidth = 0.1, color = "black"),
    axis.minor.ticks.length.x = unit(0.8, "pt"),
    axis.minor.ticks.length.y = unit(0.8, "pt"),
    
    #图例
    legend.position = "null",
    legend.title = element_text(size = 7, color = "black"),
    legend.text = element_text(size = 6, color = "black"),
    
    # 👇 核心：压缩 legend 和图之间的距离
    legend.margin = margin(t = 0, r = 0, b = -4, l = 0),
    legend.box.margin = margin(t = 0, r = 0, b = -1, l = 0),

    # 👇 再微调 item 间距
    legend.spacing.y = unit(0, "pt"),
    
    # 图例色块缩小
    legend.key.width = unit(6, "pt"),
    legend.key.height = unit(6, "pt"),

    
    
  )
```

```{r}
#| echo: true
#| warning: false

# Save the plot

p1_new_canvas <- p1_new + ggview::canvas(160,45,units="mm",dpi=600,bg="transparent")
p1_new_canvas
#ggplot2::ggsave("plodity.pdf",p1_new,width=160, height=45, units="mm",dpi=1200,bg="transparent")

```

## Function2: histogram plot

```{r}
#| echo: true
#| warning: false

# Discrete counts histogram ridge plot

# Optional fallback
if (!exists("my_family")) {
  my_family <- "sans"
}

pt_to_mm <- function(pt) pt / 2.845


df_plot3 <- merged_df %>%
  filter(
    Chromosome %in% selected_chromosomes,
    Sample %in% selected_samples
  ) %>%
  mutate(
    species = ifelse(str_detect(Chromosome, "_mel$"), "Dmel", "Dsim"),
    
    Chromosome = recode(Chromosome, !!!chromosome_rename_map),
    Chromosome = factor(
      Chromosome,
      levels = chromosome_rename_map[selected_chromosomes]
    ),
    
    Sample = recode(Sample, !!!sample_rename_map),
    Sample = factor(
      Sample,
      levels = sample_rename_map[selected_samples]
    ),
    
    Counts = as.integer(Counts),
    weight_bp = End - Start
  )

# ------------------------------------------------------------
# Build discrete weighted histogram
# ------------------------------------------------------------
# Each integer Counts value is one histogram bin.
# Counts == 0 is intentionally retained.
#
# bp:
#   Total genomic span assigned to each Counts bin.
#
# density:
#   Area-normalized histogram height.
#   Since binwidth = 1, sum(density) = 1 for each Sample + Chromosome.
#
# height_raw:
#   Unnormalized visual height.
#   Preserves absolute bp differences across samples/chromosomes.
#
# height_norm:
#   Area-normalized visual height.
#   Makes every Sample + Chromosome histogram comparable in shape.

count_bins <- seq(
  from = 0,
  to = max(df_plot3$Counts, na.rm = TRUE),
  by = 1
)

df_hist3 <- df_plot3 %>%
  group_by(Sample, Chromosome, species, Counts) %>%
  summarise(
    bp = sum(weight_bp, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(Sample, Chromosome, species) %>%
  tidyr::complete(
    Counts = count_bins,
    fill = list(bp = 0)
  ) %>%
  ungroup() %>%
  group_by(Sample, Chromosome, species) %>%
  mutate(
    total_bp = sum(bp, na.rm = TRUE),
    
    # Area-normalized histogram.
    # Since binwidth = 1, sum(density) = 1 for each Sample + Chromosome.
    density = ifelse(total_bp > 0, bp / total_bp, 0)
  ) %>%
  ungroup()


# Global scaling constants.
# These keep visual scaling comparable across all panels.

ridge_height <- 0.72

max_bp_global <- max(df_hist3$bp, na.rm = TRUE)
max_density_global <- max(df_hist3$density, na.rm = TRUE)

df_hist3 <- df_hist3 %>%
  mutate(
    height_raw = ifelse(
      max_bp_global > 0,
      bp / max_bp_global * ridge_height,
      0
    ),
    height_norm = ifelse(
      max_density_global > 0,
      density / max_density_global * ridge_height,
      0
    )
  )

# ------------------------------------------------------------
# Shared theme
# ------------------------------------------------------------

theme_counts_hist <- function() {
  theme_bw(base_family = my_family) +
    theme(
      # Transparent backgrounds
      plot.background = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(fill = "transparent", color = NA),
      
      # Grid
      panel.grid = element_blank(),
      
      # Border
      panel.border = element_rect(
        color = "black",
        fill = NA,
        linewidth = 0.1
      ),
      
      # Important: avoid double-thick borders
      axis.line = element_blank(),
      
      # Facet labels
      strip.text.x = element_text(
        angle = 0,
        hjust = 0.5,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 1, r = 0, b = 1, l = 0)
      ),
      strip.text.y.left = element_text(
        angle = 0,
        hjust = 1,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 0, r = 3, b = 0, l = 1)
      ),
      
      # Axis text
      axis.text.x = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      axis.text.y = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      
      # Axis titles
      axis.title.x = element_text(
        size = 7,
        color = "black",
        family = my_family,
      ),
      axis.title.y = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      
      # Major ticks
      axis.ticks.x = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.y = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.length = unit(1.5, "pt"),
      
      # Minor ticks
      axis.minor.ticks.x = element_line(
        linewidth = 0.1,
        color = "black"
      ),
      axis.minor.ticks.y = element_blank(),
      axis.minor.ticks.length.x = unit(0.8, "pt"),
      axis.minor.ticks.length.y = unit(0.8, "pt"),
      
      # Hide legend because species is already encoded by chromosome suffix
      # Compact legend
      legend.position = "right",
      legend.title = element_text(
        size = 6,
        color = "black",
        family = my_family,
        margin = margin(t = 0, r = 0, b = 1, l = 0)
      ),
      legend.text = element_text(
        size = 6,
        color = "black",
        family = my_family,
        margin = margin(t = 0, r = 0, b = 0, l = 0)
      ),
      legend.key.size = unit(6, "pt"),
      legend.key.width = unit(6, "pt"),
      legend.key.height = unit(6, "pt"),
      legend.key.spacing.x = unit(1, "pt"),
      legend.key.spacing.y = unit(1, "pt"),
      
      legend.spacing.x = unit(1, "pt"),
      legend.spacing.y = unit(1, "pt"),
      legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.spacing = unit(2, "pt"),
      
      # Spacing
      # Facet spacing
      panel.spacing.x = unit(0.2, "lines"),
      panel.spacing.y = unit(0.4, "lines"),
      strip.placement = "outside",
      strip.switch.pad.grid = unit(0, "pt"),
      strip.clip = "off",
      
      # Plot margin
      plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
    )
}

# ------------------------------------------------------------
# 6. Shared scales
# ------------------------------------------------------------

x_max <- 55

scale_x_counts <- scale_x_continuous(
  expand = c(0.05, 0.05),
  breaks = seq(0, 60, by = 20),
  minor_breaks = seq(0, 60, by = 4),
  guide = guide_axis(minor.ticks = TRUE)
)

scale_y_raw <- scale_y_continuous(
  expand = expansion(mult = c(0.05, 0.05)),
  breaks = function(x) {
    pretty(x, n = 2)
  },
  labels = function(x) {
    scales::label_number(
      accuracy = 0.1,
      trim = TRUE
    )(x / 1e6)
  },
  guide = guide_axis(minor.ticks = FALSE)
)

scale_y_norm <- scale_y_continuous(
  expand = expansion(mult = c(0.05, 0.05)),
  breaks = function(x) {
    pretty(x, n = 2)
  },
  labels = function(x) {
    scales::label_number(
      accuracy = 0.01,
      trim = TRUE
    )(x)
  },
  guide = guide_axis(minor.ticks = FALSE)
)


# ------------------------------------------------------------
# Plot A: raw weighted histogram
# Facet by Chromosome and Sample
# ------------------------------------------------------------
# This version is not area-normalized.
# It preserves absolute genomic span differences.
# Use free_y because raw bp values can differ strongly across chromosomes/samples.


bar_half_width <- 0.35

plot2_raw <- ggplot(df_hist3) +
  geom_rect(
    aes(
      xmin = Counts - bar_half_width,
      xmax = Counts + bar_half_width,
      ymin = 0,
      ymax = bp,
      fill = species
    ),
    color = NA,
    alpha = 0.95
  ) +
  facet_grid(
    cols = vars(Sample),
    rows = vars(Chromosome),
    scales = "free_y",
    switch = "both"
  )+
  scale_fill_manual(
    values = species_colors,
    name = "Chromosome"
  ) +
  scale_x_counts +
  scale_y_raw +
  coord_cartesian(
    xlim = c(0, x_max),
    clip = "on"
  ) +
  labs(
    x = "Counts",
    y = "Genomic span (Mb)"
  )+
  theme_counts_hist() +
  theme(
    strip.placement = "outside"
  )

# ------------------------------------------------------------
# Plot B: area-normalized histogram
# Facet by Chromosome and Sample
# ------------------------------------------------------------
# This version normalizes each Sample + Chromosome histogram to area = 1.
# Because binwidth = 1:
#
#   sum(density) = 1
#
# This is better for comparing histogram shapes across samples/chromosomes.

plot2_norm <- ggplot(df_hist3) +
  geom_rect(
    aes(
      xmin = Counts - bar_half_width,
      xmax = Counts + bar_half_width,
      ymin = 0,
      ymax = density,
      fill = species
    ),
    color = NA,
    alpha = 0.95
  ) +
  facet_grid(
    cols = vars(Sample),
    rows = vars(Chromosome),
    scales = "free_y",
    switch = "both"
  )+
  scale_fill_manual(
    values = species_colors,
    name = "Chromosome"
  ) +
  scale_x_counts +
  scale_y_norm +
  coord_cartesian(
    xlim = c(0, x_max),
    clip = "on"
  ) +
  labs(
    x = "Counts",
    y = "Density"
  ) +
  theme_counts_hist() +
  theme(
    strip.placement = "outside"
  )
```

Save the plot

```{r}
#| echo: true
#| warning: false

# Save the plot

plot2_raw_canvas <- plot2_raw + ggview::canvas(170,120,units="mm",dpi=600,bg="transparent")
plot2_raw_canvas
#ggplot2::ggsave("figS3.pdf",plot2_raw,width=170, height=120, units="mm",dpi=900,bg="transparent")


plot2_norm_canvas <- plot2_norm + ggview::canvas(160,120,units="mm",dpi=600,bg="transparent")
plot2_norm_canvas
#ggplot2::ggsave("plodity.pdf",plot2_norm,width=160, height=45, units="mm",dpi=1200,bg="transparent")
```

## Function3: normalized to the length of chromosome

```{r}
#| echo: true
#| warning: false

# ============================================================
# Discrete counts histogram plot
# New branch: df_plot4 / plot3
# Y axis = genomic span normalized by chromosome length (%)
# ============================================================

# ------------------------------------------------------------
# 1. Prepare df_plot4
# ------------------------------------------------------------
# Ras3 is D. melanogaster control, so only _mel chromosomes are shown.
# ML82 is D. simulans control, so only _sim chromosomes are shown.
#
# Chromosome length is inferred from Start/End:
# chr_length_bp = max(End) - min(Start)
#
# This assumes the input intervals cover the chromosome coordinate range.

df_plot4 <- merged_df %>%
  filter(
    Chromosome %in% selected_chromosomes,
    Sample %in% selected_samples,
    !(
      (str_detect(Chromosome, "_mel$") & Sample == "ML82") |
      (str_detect(Chromosome, "_sim$") & Sample == "Ras3")
    )
  ) %>%
  mutate(
    species = ifelse(str_detect(Chromosome, "_mel$"), "Dmel", "Dsim"),
    
    Chromosome = recode(Chromosome, !!!chromosome_rename_map),
    Chromosome = factor(
      Chromosome,
      levels = chromosome_rename_map[selected_chromosomes]
    ),
    
    Sample = recode(Sample, !!!sample_rename_map),
    Sample = factor(
      Sample,
      levels = sample_rename_map[selected_samples]
    ),
    
    Counts = as.integer(Counts),
    weight_bp = End - Start
  ) %>%
  group_by(Sample, Chromosome) %>%
  mutate(
    chr_length_bp = max(End, na.rm = TRUE) - min(Start, na.rm = TRUE)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2. Build discrete weighted histogram
# ------------------------------------------------------------
# Each integer Counts value is one histogram bin.
# Counts == 0 is intentionally retained.
#
# bp:
#   Genomic span assigned to this Counts bin.
#
# chr_length_bp:
#   Chromosome length inferred from the coordinate range.
#
# percent_chr:
#   Percentage of chromosome length represented by this Counts bin.
#
# If the intervals fully cover the chromosome, then:
#
#   sum(percent_chr) ≈ 100
#
# for each Sample + Chromosome.

count_bins <- seq(
  from = 0,
  to = max(df_plot4$Counts, na.rm = TRUE),
  by = 1
)

chr_length_df4 <- df_plot4 %>%
  distinct(Sample, Chromosome, species, chr_length_bp)

df_hist4 <- df_plot4 %>%
  group_by(Sample, Chromosome, species, Counts) %>%
  summarise(
    bp = sum(weight_bp, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(Sample, Chromosome, species) %>%
  tidyr::complete(
    Counts = count_bins,
    fill = list(bp = 0)
  ) %>%
  ungroup() %>%
  left_join(
    chr_length_df4,
    by = c("Sample", "Chromosome", "species")
  ) %>%
  mutate(
    percent_chr = ifelse(
      chr_length_bp > 0,
      bp / chr_length_bp * 100,
      0
    )
  )

# Optional QC table:
# Each Sample + Chromosome should be close to 100 if intervals cover the chromosome.
df_hist4_qc <- df_hist4 %>%
  group_by(Sample, Chromosome, species) %>%
  summarise(
    chr_length_mb = first(chr_length_bp) / 1e6,
    total_bp_in_bins = sum(bp, na.rm = TRUE),
    total_percent_chr = sum(percent_chr, na.rm = TRUE),
    .groups = "drop"
  )



# ------------------------------------------------------------
# 3. Shared theme
# ------------------------------------------------------------

theme_counts_percent <- function() {
  theme_bw(base_family = my_family) +
    theme(
      # Transparent backgrounds
      plot.background = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(fill = "transparent", color = NA),
      
      # Grid
      panel.grid = element_blank(),
      
      # Border
      panel.border = element_rect(
        color = "black",
        fill = NA,
        linewidth = 0.1
      ),
      
      # Avoid double-thick borders
      axis.line = element_blank(),
      
      # Facet labels
      strip.text.x = element_text(
        angle = 0,
        hjust = 0.5,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 1, r = 0, b = 1, l = 0)   # 关键：去掉 strip 自带空白
      ),
      strip.text.y.left = element_text(
        angle = 0,
        hjust = 1,
        vjust = 0.5,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 0, r = 3, b = 0, l = 1)
      ),
      
      # Axis text
      axis.text.x = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      axis.text.y = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      
      # Axis titles
      axis.title.x = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      axis.title.y = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      
      # Major ticks
      axis.ticks.x = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.y = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.length = unit(1.5, "pt"),
      
      # Minor ticks
      axis.minor.ticks.x = element_line(
        linewidth = 0.1,
        color = "black"
      ),
      axis.minor.ticks.length.x = unit(0.8, "pt"),
      axis.minor.ticks.length.y = unit(0, "pt"),
      
      # Hide legend because species is encoded by chromosome suffix
      legend.position = "none",
      
      # Facet spacing
      panel.spacing.x = unit(0.2, "lines"),
      panel.spacing.y = unit(0.4, "lines"),
      strip.placement = "outside",
      strip.switch.pad.grid = unit(0, "pt"),
      strip.clip = "off",
      
      # Plot margin
      plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
    )
}

# ------------------------------------------------------------
# 4. Shared scales
# ------------------------------------------------------------

x_max <- 55
bar_half_width <- 0.45

scale_x_counts <- scale_x_continuous(
  expand = c(0.05, 0.05),
  breaks = seq(0, 60, by = 20),
  minor_breaks = seq(0, 60, by = 4),
  guide = guide_axis(minor.ticks = TRUE)
)

scale_y_percent_chr <- scale_y_continuous(
  expand = expansion(mult = c(0.05, 0.05)),
  breaks = function(x) {
    pretty(x, n = 2)
  },
  labels = function(x) {
    scales::label_number(
      accuracy = 1,
      trim = TRUE
    )(x)
  },
  guide = guide_axis(minor.ticks = FALSE)
)

# ------------------------------------------------------------
# 5. Plot 3: chromosome-length-normalized histogram
# ------------------------------------------------------------
# X axis:
#   Counts
#
# Y axis:
#   Percentage of chromosome length in each Counts bin.
#
# Faceting:
#   Rows = Sample
#   Columns = Chromosome
#
# Empty panels are expected for:
#   ML82 x Dmel chromosomes
#   Ras3 x Dsim chromosomes
#
# because these combinations are filtered out above.

df_hist4_plot <- df_hist4 %>%
  filter(
    Counts >= 0,
    Counts <= x_max
  ) %>%
  mutate(
    xmin_plot = pmax(Counts - bar_half_width, 0),
    xmax_plot = pmin(Counts + bar_half_width, x_max)
  )



plot3 <- ggplot(df_hist4_plot) +
  geom_rect(
    aes(
      xmin = xmin_plot,
      xmax = xmax_plot,
      ymin = 0,
      ymax = percent_chr,
      fill = species
    ),
    color = NA,
    alpha = 0.95
  ) +
  facet_grid(
    rows = vars(Sample),
    cols = vars(Chromosome),
    scales = "free_y",
    switch = "both"
  ) +
  scale_fill_manual(
    values = species_colors,
    name = "Chromosome"
  ) +
  scale_x_counts +
  scale_y_percent_chr +
  coord_cartesian(
    xlim = c(0, x_max),
    clip = "off"
  ) +
  labs(
    x = "Counts",
    y = "Chromosome span (%)"
  ) +
  theme_counts_percent() +
  theme(
    strip.placement = "outside"
  )
```

Save the plot

```{r}
#| echo: true
#| warning: false

plot3_canvas <- plot3 + ggview::canvas(160,60,units="mm",dpi=600,bg="transparent")
plot3_canvas
#ggplot2::ggsave("fig2B.pdf",plot3,width=160, height=60, units="mm",dpi=900,bg="transparent")
```

## Function4: counts=0 barplot

```{r}
#| echo: true
#| warning: false

# ============================================================
# Zero vs non-zero counts summary
# New branch: df_plot5 / plot4
# Y axis = percentage of chromosome length
# ============================================================


# Optional fallback
if (!exists("my_family")) {
  my_family <- "sans"
}


# ------------------------------------------------------------
# 1. Prepare df_plot5
# ------------------------------------------------------------
# Ras3 is D. melanogaster control, so only _mel chromosomes are shown.
# ML82 is D. simulans control, so only _sim chromosomes are shown.

df_plot5 <- merged_df %>%
  filter(
    Chromosome %in% selected_chromosomes,
    Sample %in% selected_samples,
    !(
      (str_detect(Chromosome, "_mel$") & Sample == "ML82") |
      (str_detect(Chromosome, "_sim$") & Sample == "Ras3")
    )
  ) %>%
  mutate(
    species = ifelse(str_detect(Chromosome, "_mel$"), "Dmel", "Dsim"),
    
    Chromosome = recode(Chromosome, !!!chromosome_rename_map),
    Chromosome = factor(
      Chromosome,
      levels = chromosome_rename_map[selected_chromosomes]
    ),
    
    Sample = recode(Sample, !!!sample_rename_map),
    Sample = factor(
      Sample,
      levels = sample_rename_map[selected_samples]
    ),
    
    Counts = as.integer(Counts),
    weight_bp = End - Start,
    
    count_group = ifelse(Counts == 0, "0", ">0"),
    count_group = factor(count_group, levels = c("0", ">0"))
  ) %>%
  group_by(Sample, Chromosome) %>%
  mutate(
    chr_length_bp = max(End, na.rm = TRUE) - min(Start, na.rm = TRUE)
  ) %>%
  ungroup()

# ------------------------------------------------------------
# 2. Summarise zero and non-zero genomic span
# ------------------------------------------------------------

df_zero5 <- df_plot5 %>%
  group_by(Sample, Chromosome, species, count_group) %>%
  summarise(
    bp = sum(weight_bp, na.rm = TRUE),
    chr_length_bp = first(chr_length_bp),
    .groups = "drop"
  ) %>%
  group_by(Sample, Chromosome, species) %>%
  tidyr::complete(
    count_group = factor(c("0", ">0"), levels = c("0", ">0")),
    fill = list(bp = 0)
  ) %>%
  mutate(
    chr_length_bp = first(na.omit(chr_length_bp)),
    percent_chr = ifelse(
      chr_length_bp > 0,
      bp / chr_length_bp * 100,
      0
    ),
    label = paste0(
      scales::number(percent_chr, accuracy = 0.1, trim = TRUE),
      "%"
    )
  ) %>%
  ungroup()

# Optional QC:
# For each Sample + Chromosome, total_percent_chr should be close to 100.

df_zero5_qc <- df_zero5 %>%
  group_by(Sample, Chromosome, species) %>%
  summarise(
    chr_length_mb = first(chr_length_bp) / 1e6,
    total_bp = sum(bp, na.rm = TRUE),
    total_percent_chr = sum(percent_chr, na.rm = TRUE),
    .groups = "drop"
  )


# ------------------------------------------------------------
# 3. Colors
# ------------------------------------------------------------

zero_colors <- c(
  "0"  = "#f2481b",
  ">0" = "#0eb0c9"
)

# ------------------------------------------------------------
# 4. Theme
# ------------------------------------------------------------

theme_zero_percent <- function() {
  theme_bw(base_family = my_family) +
    theme(
      # Transparent backgrounds
      plot.background = element_rect(fill = "transparent", color = NA),
      panel.background = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(fill = "transparent", color = NA),
      
      # Grid
      panel.grid = element_blank(),
      
      # Border
      panel.border = element_rect(
        color = "black",
        fill = NA,
        linewidth = 0.1
      ),
      
      axis.line = element_blank(),
      
      # Facet labels
      strip.text.x = element_text(
        angle = 0,
        hjust = 0.5,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 1, r = 0, b = 1, l = 0)
      ),
      strip.text.y.left = element_text(
        angle = 0,
        hjust = 1,
        size = 6,
        family = my_family,
        color = "black",
        margin = margin(t = 0, r = 3, b = 0, l = 1)
      ),
      
      # Axis text
      axis.text.x = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      axis.text.y = element_text(
        size = 6,
        color = "black",
        family = my_family
      ),
      
      # Axis titles
      axis.title.x = element_text(
        size = 7,
        color = "black",
        family = my_family,
      ),
      axis.title.y = element_text(
        size = 7,
        color = "black",
        family = my_family
      ),
      
      # Major ticks
      axis.ticks.x = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.y = element_line(
        linewidth = 0.2,
        color = "black"
      ),
      axis.ticks.length = unit(1.5, "pt"),
      
      # Minor ticks
      axis.minor.ticks.x = element_blank(),
      axis.minor.ticks.y = element_blank(),
      axis.minor.ticks.length.x = unit(0, "pt"),
      axis.minor.ticks.length.y = unit(0, "pt"),
      
      # Compact legend
      legend.position = "right",
      legend.title = element_blank(),
      legend.text = element_text(
        size = 6,
        color = "black",
        family = my_family,
        margin = margin(t = 0, r = 0, b = 0, l = 2)
      ),
      legend.key.size = unit(6, "pt"),
      legend.key.width = unit(6, "pt"),
      legend.key.height = unit(6, "pt"),
      legend.key.spacing.x = unit(1, "pt"),
      legend.key.spacing.y = unit(1, "pt"),
      
      legend.spacing.x = unit(1, "pt"),
      legend.spacing.y = unit(1, "pt"),
      legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.margin = margin(t = 0, r = 0, b = 0, l = 0),
      legend.box.spacing = unit(2, "pt"),
      
      # Facet spacing
      panel.spacing.x = unit(0.2, "lines"),
      panel.spacing.y = unit(0.4, "lines"),
      strip.placement = "outside",
      strip.switch.pad.grid = unit(0, "pt"),
      strip.clip = "off",
      
      # Plot margin
      plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
    )
}

# ------------------------------------------------------------
# 5. Scales
# ------------------------------------------------------------

scale_y_zero_percent <- scale_y_continuous(
  limits = c(0, 150),
  expand = expansion(mult = c(0, 0.6)),
  breaks = c(0, 100),
  labels = function(x) {
    scales::label_number(
      accuracy = 1,
      trim = TRUE
    )(x)
  },
  guide = guide_axis(minor.ticks = FALSE)
)

# ------------------------------------------------------------
# 6. Plot 4: zero vs non-zero chromosome span
# ------------------------------------------------------------

plot4 <- ggplot(df_zero5) +
  geom_col(
    aes(
      x = count_group,
      y = percent_chr,
      fill = count_group
    ),
    width = 0.65,
    color = NA,
    linewidth = 0.15,
    alpha = 0.95
  ) +
  geom_text(
    aes(
      x = count_group,
      y = percent_chr,
      label = label
    ),
    vjust = -0.25,
    size = 5 / 2.845,
    family = my_family,
    color = "black"
  ) +
  facet_grid(
    rows = vars(Sample),
    cols = vars(Chromosome),
    scales = "fixed",
    switch = "both"
  ) +
  scale_fill_manual(
    values = zero_colors,
    labels = c(
      "0" = "Counts = 0",
      ">0" = "Counts > 0"
    )
  ) +
  scale_y_zero_percent +
  coord_cartesian(
    ylim = c(0, 105),
    clip = "off"
  ) +
  labs(
    x = "Counts",
    y = "Chromosome span (%)"
  ) +
  theme_zero_percent() +
  guides(
    fill = guide_legend(
      keywidth = unit(6, "pt"),
      keyheight = unit(6, "pt"),
      ncol = 1,
      bycol = TRUE,
      label.position = "right",
      label.hjust = 0,
      override.aes = list(
      colour = NA,
      alpha = 0.95
    )
    )
  )

```

Save the plot

```{r}
#| echo: true
#| warning: false


plot4_canvas <- plot4 + ggview::canvas(170,50,units="mm",dpi=300,bg="transparent")
plot4_canvas
#ggplot2::ggsave("fig2B.pdf",plot4,width=170, height=50, units="mm",dpi=900,bg="transparent")
```

# Check an individual chromosome

```{r}
#| echo: true
#| warning: false

df_plot6 <- df_plot2

df_plot6 %>%
  filter(Sample == "HYC4", Chromosome == "4_mel") %>%
  summarise(
    n_rows = n(),
    nonzero_n = sum(Counts > 0, na.rm = TRUE),
    zero_n = sum(Counts == 0, na.rm = TRUE),
    mean_counts = mean(Counts, na.rm = TRUE),
    median_counts = median(Counts, na.rm = TRUE),
    max_counts = max(Counts, na.rm = TRUE),
    total_weight = sum(End - Start, na.rm = TRUE),
    nonzero_weight = sum((End - Start)[Counts > 0], na.rm = TRUE)
  )


df_plot6 %>%
  filter(Sample == "HYC4", Chromosome == "4_mel") %>%
  ggplot(aes(x = Counts, weight = End - Start)) +
  geom_histogram(binwidth = 1, color = "black", fill = "grey80") +
  theme_bw()
```

As shown in Figure @fig-counts-distribution, the 5B samples have two peaks.

# Barplot

```{r}
#| echo: true
#| warning: false
#| label: fig-mean_coverage
#| fig-align: center
#| fig-width: 10
#| fig-height: 8
#| fig-dpi: 600
#| fig-format: svg
#| fig-cap: "Mean coverage"
#| fig-cap-location: bottom

selected_chromosomes <- c(
  "chrX_mel","chr2L_mel","chr2R_mel","chr3L_mel","chr3R_mel","chr4_mel",
  "chrX_sim","chr2L_sim","chr2R_sim","chr3L_sim","chr3R_sim","chr4_sim")

# 你想保留的 sample
selected_samples <- c("5A","5B_19","6B","7B_23","ML82","Ras3" )

# 重命名规则（旧名 = 新名）
sample_rename_map <- c(
  "5A" = "HYC1",
  "5B_19" = "HYC2",
  "6B" = "HYC3",
  "7B_23" = "HYC4",
  "ML82"="ML82",
  "Ras3"="Ras3"
)


merged_df_without_zero <- subset(merged_df, Counts > 0)

df_cov <- merged_df_without_zero %>%
  filter(
    Chromosome %in% selected_chromosomes,
    Sample %in% selected_samples   # ✅ 筛选 sample
  ) %>%
  mutate(
    Chromosome = factor(Chromosome, levels = selected_chromosomes),

    # ✅ 重命名 sample
    Sample = recode(Sample, !!!sample_rename_map),

    # ✅ 顺序也一起控制（很关键）
    Sample = factor(Sample, levels = sample_rename_map[selected_samples])
  ) %>%
  group_by(Sample, Chromosome) %>%
  summarise(
    mean_coverage = sum(Counts * (End - Start)) / sum(End - Start),
    .groups = "drop"
  )

mel_chroms <- grep("_mel$", levels(df_cov$Chromosome), value = TRUE)
sim_chroms <- grep("_sim$", levels(df_cov$Chromosome), value = TRUE)

mel_colors <- brewer.pal(n = length(mel_chroms), name = "Blues")
sim_colors <- brewer.pal(n = length(sim_chroms), name = "Oranges")

chrom_colors <- c(mel_colors, sim_colors)
names(chrom_colors) <- c(mel_chroms, sim_chroms)


p2 <- ggplot(df_cov, aes(x = Chromosome, y = mean_coverage, fill = Chromosome)) +
  geom_col() +
  facet_wrap(~Sample, scales = "free_y") +
  scale_fill_manual(values = chrom_colors) +
  labs(
    title = "Mean Coverage (B)",
    x = "Chromosome",
    y = "Mean Coverage"
  ) +
  theme_minimal(base_family = "sans") +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

```

## Median coverage

```{r}
#| echo: true
#| warning: false
#| label: fig-median_coverage
#| fig-align: center
#| fig-width: 10
#| fig-height: 8
#| fig-dpi: 600
#| fig-format: svg
#| fig-cap: "Mean coverage"
#| fig-cap-location: bottom


selected_chromosomes <- c(
  "chrX_mel","chr2L_mel","chr2R_mel","chr3L_mel","chr3R_mel","chr4_mel",
  "chrX_sim","chr2L_sim","chr2R_sim","chr3L_sim","chr3R_sim","chr4_sim")

# 你想保留的 sample
selected_samples <- c("5A","5B_17","6B","7B_21","ML82","Ras3" )

# 重命名规则（旧名 = 新名）
sample_rename_map <- c(
  "5A" = "HYC1",
  "5B_17" = "HYC2",
  "6B" = "HYC3",
  "7B_21" = "HYC4",
  "ML82"="ML82",
  "Ras3"="Ras3"
)


weighted_median <- function(x, w, na.rm = TRUE) {
  if (na.rm) {
    keep <- !(is.na(x) | is.na(w))
    x <- x[keep]
    w <- w[keep]
  }
  
  if (length(x) == 0 || sum(w) == 0) {
    return(NA_real_)
  }
  
  ord <- order(x)
  x <- x[ord]
  w <- w[ord]
  
  cum_w <- cumsum(w) / sum(w)
  x[which(cum_w >= 0.5)[1]]
}

# Change between 0 to -1 to have different plots.
merged_df_without_zero <- subset(merged_df, Counts > -1)


df_cov <- merged_df_without_zero %>%
  filter(
    Chromosome %in% selected_chromosomes,
    Sample %in% selected_samples
  ) %>%
  mutate(
    Chromosome = factor(Chromosome, levels = selected_chromosomes),
    Sample = recode(Sample, !!!sample_rename_map),
    Sample = factor(Sample, levels = sample_rename_map[selected_samples]),
    width = End - Start
  ) %>%
  group_by(Sample, Chromosome) %>%
  summarise(
    mean_coverage = sum(Counts * width, na.rm = TRUE) / sum(width[!is.na(Counts)]),
    median_coverage = weighted_median(Counts, width),
    .groups = "drop"
  )

mel_chroms <- grep("_mel$", levels(df_cov$Chromosome), value = TRUE)
sim_chroms <- grep("_sim$", levels(df_cov$Chromosome), value = TRUE)

mel_colors <- brewer.pal(n = length(mel_chroms), name = "Blues")
sim_colors <- brewer.pal(n = length(sim_chroms), name = "Oranges")

chrom_colors <- c(mel_colors, sim_colors)
names(chrom_colors) <- c(mel_chroms, sim_chroms)


p3 <- ggplot(df_cov, aes(x = Chromosome, y = median_coverage, fill = Chromosome)) +
  geom_col() +
  facet_wrap(~Sample, scales = "free_y") +
  scale_fill_manual(values = chrom_colors) +
  labs(
    title = "Median Coverage (A)",
    x = "Chromosome",
    y = "Mean Coverage"
  ) +
  theme_minimal(base_family = "sans") +
  theme(
    panel.grid = element_blank(),
    panel.border = element_rect(color = "black", fill = NA),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

```
