---
title: "Fig1. Quantification of probe foci"
author: "Jie Hua"
date: 2025-10-01
lang: en
date-modified: 2026-04-08
keywords: ["IF","Probe","Hybrid Cells","DNA-FISH"]
code-copy: true
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: true
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)
library(viridis)
library(ggsignif)
library(dunn.test)
library(tinytable)
```

# Load quantification data

I have already counted the number of probe foci. The result is in multiple csv file.

## Define a function to import the *.csv* files

```{r}
#| echo: true
#| warning: false

# Write a loop to read all the csv files in one folder, save data to a list and then merge them into one data frame
data_import <- function(folder, pattern){
  # Get all the files in the folder
  cat ("The folder is: ", folder, "\n")
  files <- list.files(folder, pattern = pattern, full.names = TRUE)
  # Create an empty list to store the data
  data_list <- list()
  # Loop through all the files
  for (file in files){
    cat ("Reading file: ", file, "\n")
    # Read the data
    data <- read.csv(file)
    # Store the data in the list
    data_list[[file]] <- data
  }
  # Merge all the data into one data frame
  data <- do.call(rbind, data_list)
  rownames(data) <- NULL
  return(data)
}
```

## Read the tables

```{r}
#| echo: true
#| warning: false

# Read the quantification result
data_raw <- data_import(folder="F:\\Big Paper\\Processed data\\Fig1 DNA-FISH probe dot quantification",pattern="\\.csv$")

#data_raw <- data_import(folder="H:\\HybridCells_Data\\Fig1 Probe Foci Quantification\\Processed_data_Fig1sup_Probe Dots Quantification",pattern="\\.csv$")
```

Have a quick check

```{r}
#| echo: true
#| warning: false
#| label: Rawdata preview

tinytable::tt(head(data_raw, 10))
```

Rename those samples

Four hybrid cell lines were renamed as "HYC1" to "HYC4".

```{r}
#| echo: true
#| warning: false

data_raw <- data_raw %>%
  mutate(Sample = recode(Sample,
                         "5a" = "HYC1",
                         "5b" = "HYC2",
                         "6b" = "HYC3",
                         "7b" = "HYC4"))

data_raw <- data_raw %>%
  mutate(Probe = recode(Probe,
                        "probe1"="Dmel/Dsim",
                        "probe23"="Dsim",
                        "probe24"="Dmel"
    
  ))
```

# Summarize the data

```{r}
#| echo: true
#| warning: false

data_summarise <- data_raw %>%
  dplyr::summarise(
    Mean = mean(Number),
    SD = sd(Number),
    Count = n(),
    .by = c(Sample, Probe)
  )
```

A quick boxplot for the data

```{r}
#| echo: true
#| warning: false
#| fig-width: 8
#| fig-height: 5
#| fig-align: center

# Dot plot does not work. Too many dots.
boxplot <- ggplot(data_raw, aes(x = Sample, y = Number, fill = Probe)) +
  geom_boxplot(
    alpha = 0.45,
    color = "black",
    linewidth = 0.25,
    outlier.size = 0.4,
    outlier.alpha = 0.25,
    position = position_dodge(width = 0.75)
  ) +
  stat_summary(
    fun = median,
    geom = "text",
    aes(label = round(after_stat(y), 1), group = Probe),
    position = position_dodge(width = 0.75),
    vjust = -0.45,
    color = "black",
    size = 2.6,
    fontface = "bold"
  ) +
  scale_fill_viridis_d(
    option = "D",
    end = 0.9
  ) +
  labs(
    x = "Sample",
    y = "Number",
    fill = "Probe"
  ) +
  theme_classic(base_size = 8) +
  theme(
    axis.line = element_line(linewidth = 0.3, color = "black"),
    axis.ticks = element_line(linewidth = 0.25, color = "black"),
    axis.text = element_text(color = "black", size = 7),
    axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1),
    axis.title = element_text(color = "black", size = 8),
    
    legend.position = "right",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7, face = "italic"),
    legend.key.size = unit(4, "mm"),
    
    plot.margin = margin(4, 8, 4, 4)
  )

boxplot
```

# Define a function to perform static test and make plot

## Function A: auto Y-axis

```{r}
#| echo: true
#| warning: false

report_probe <- function(data_raw, probe, exclude_samples,custom_order, custom_colors){
  
  # Select specific probe and exclude some sample
  subset_data <- data_raw[data_raw$Probe == probe, ]
  subset_data <- subset_data[!subset_data$Sample %in% exclude_samples, ]
  
  # Shapiro.test
  shapiro_results <- lapply(split(subset_data, subset_data$Sample), function(group) {
    if (length(unique(group$Number)) > 1) {
      return(shapiro.test(group$Number)$p.value)
    } else {
      return(NA)
    }
  })
  
  # Print the Shapiro.test result
  cat(paste("Probe:", probe, "\n"))
  for (sample in names(shapiro_results)) {
    p_value <- shapiro_results[[sample]]
    if (!is.na(p_value)) {
      cat(sprintf("Sample: %s Shapiro-Wilk p-value: %.2e\n", sample, p_value))
    } else {
      cat(sprintf("Sample: %s Shapiro-Wilk test skipped (all values are the same)\n", sample))
    }
  }
  

  # Kruskal-Wallis test
  kruskal_result <- kruskal.test(Number ~ Sample, data = subset_data)
  cat(sprintf("Kruskal-Wallis p-value for Probe %s: %.2e\n\n", probe, kruskal_result$p.value))
  kruskal_p_value <- kruskal_result$p.value
  kruskal_p_text <- paste("Kruskal-Wallis p-value:", format(kruskal_p_value, digits = 2))
  
  # Dunn test（Bonferroni correction）
  dunn_result <- dunn.test(subset_data$Number, subset_data$Sample, method = "bonferroni")
  
  # Extract the result
  comparisons <- data.frame(
    group = dunn_result$comparisons,
    p_value = dunn_result$P.adjusted
  )
  
  # Add Probe Information
  comparisons$Probe <- probe
  # 筛选感兴趣的组对（vs SL2 或 ML82）
  filtered_comparisons <- comparisons[
    (grepl("Ras3|ML82", comparisons$group) & grepl("HYC1|HYC2|HYC3|HYC4", comparisons$group)) |
      grepl("ML82 - Ras3", comparisons$group),
  ]
  comparisons <- filtered_comparisons
  
  # 计算 Y 轴的下限，避免覆盖数据
  y_min <- min(subset_data$Number)
  y_min <- ifelse(y_min >= 0, 0, y_min * 1.2)
  
  # 创建箱线图
  boxplot <- ggplot(subset_data, aes(x = Sample, y = Number, fill = Sample)) +
    geom_boxplot(
      size = 0.5,               # 较细的边框
      colour = "grey20",        # 浅灰边框，不挡住中位数
      outlier.color = "grey40",  # 离群点也浅灰色
      fatten = 3,
    )+
    labs(
      #title = paste("Boxplot of Number for Probe:", probe),
      x = "Sample",
      y = bquote("Number of "*italic(.(probe))*" Probe Foci")
    ) +
    scale_x_discrete(limits = custom_order) +
    scale_fill_manual(values = custom_colors) +
    scale_y_continuous(expand = c(0, 0), limits = c(y_min, max(subset_data$Number + 8))) +
    theme_minimal() +
    theme(
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6),
      #axis.line = element_line(color = "black", linewidth  = 0.5),
      #plot.title = element_text(hjust = 0.5, size = 14, family = "sans", face = "bold", color = "black"),
      axis.text.x = element_text(hjust = 0.5,size = 15,family = "sans",color="black"),
      axis.text.y = element_text(size = 15,family = "sans",color="black"),
      axis.title.x = element_text(size = 15,family = "sans",color="black"),
      axis.title.y = element_text(size = 15,family = "sans",color="black"),
      legend.position = "none",
      panel.grid = element_blank(),
      )+
    # 添加 Kruskal-Wallis p 值到左上角
    annotate(
      "text",
      x = 1, y = max(subset_data$Number) + 7.5,
      hjust = 0,# 调整位置
      label = kruskal_p_text,
      size = 5, color = "black"
    )
  
  # 遍历每个比较结果，添加显著性标注到图中
  for (i in 1:nrow(comparisons)) {
    group <- unlist(strsplit(comparisons$group[i], " - "))
    p_value <- comparisons$p_value[i]
    
    # 根据 p 值设置显著性标注
    significance <- ifelse(p_value < 0.001, "***",
                           ifelse(p_value < 0.01, "**",
                                  ifelse(p_value < 0.05, "*", "ns")))
    
    # 在图中添加显著性
    boxplot <- boxplot +
      geom_signif(
        comparisons = list(group),
        annotations = significance,
        y_position = max(subset_data$Number) + 0.6*i,
        textsize = 5,
        map_signif_level = FALSE,
        size = 0.5
      )
  }
  
  return(list(boxplot=boxplot, comparisons=comparisons))
  
}

```

## Function B: fixed Y-axis

```{r}
#| echo: true
#| warning: false


report_probe2 <- function(data_raw, probe, exclude_samples, custom_order, custom_colors) {
  
  subset_data <- data_raw[data_raw$Probe == probe, ]
  subset_data <- subset_data[!subset_data$Sample %in% exclude_samples, ]
  
  # Shapiro.test
  shapiro_results <- lapply(split(subset_data, subset_data$Sample), function(group) {
    if (length(unique(group$Number)) > 1) {
      return(shapiro.test(group$Number)$p.value)
    } else {
      return(NA)
    }
  })
  
  cat(paste("Probe:", probe, "\n"))
  for (sample in names(shapiro_results)) {
    p_value <- shapiro_results[[sample]]
    if (!is.na(p_value)) {
      cat(sprintf("Sample: %s Shapiro-Wilk p-value: %.2e\n", sample, p_value))
    } else {
      cat(sprintf("Sample: %s Shapiro-Wilk test skipped (all values are the same)\n", sample))
    }
  }
  
  # Kruskal-Wallis test
  kruskal_result <- kruskal.test(Number ~ Sample, data = subset_data)
  cat(sprintf("Kruskal-Wallis p-value for Probe %s: %.2e\n\n", probe, kruskal_result$p.value))
  
  kruskal_p_value <- kruskal_result$p.value
  kruskal_p_text <- paste("Kruskal-Wallis p-value:", format(kruskal_p_value, digits = 2))
  
  # Dunn test
  dunn_result <- dunn.test(subset_data$Number, subset_data$Sample, method = "bonferroni")
  
  comparisons <- data.frame(
    group = dunn_result$comparisons,
    p_value = dunn_result$P.adjusted
  )
  
  comparisons$Probe <- probe
  
  filtered_comparisons <- comparisons[
    (grepl("Ras3|ML82", comparisons$group) & grepl("HYC1|HYC2|HYC3|HYC4", comparisons$group)) |
      grepl("ML82 - Ras3", comparisons$group),
  ]
  comparisons <- filtered_comparisons
  
  # ===== 固定坐标轴与注释位置 =====
  y_lim <- c(0, 17)
  kw_y <- 16.5
  sig_start <- 6.5
  sig_step <- 0.93
  
  boxplot <- ggplot(subset_data, aes(x = Sample, y = Number, fill = Sample)) +
    geom_boxplot(
      size = 0.2,
      colour = "grey20",
      outlier.color = NA,
      fatten = 1.5
    ) +
    labs(
      x = "",
      y = bquote(""*italic(.(probe))*" Probe Foci")
    ) +
    scale_x_discrete(limits = custom_order) +
    scale_fill_manual(values = custom_colors) +
    scale_y_continuous(
      expand = c(0.05, 0),
      limits = y_lim,
      breaks = seq(0, 15, by = 5),         # 主刻度
      minor_breaks = seq(0, 15, by = 1),   # 小刻度
      guide = guide_axis(minor.ticks = TRUE)
    )+
    theme_bw() +
    theme(
      # 背景透明
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background  = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(fill = "transparent", color = NA),
      panel.grid = element_blank(),
      panel.spacing.x = unit(0.6, "lines"),
      panel.spacing.y = unit(0.5, "lines"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.7),
      axis.text.x = element_text(size = 6, color = "black"),
      axis.text.y = element_text(size = 6, color = "black"),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 7),
      
      # 主刻度
      axis.ticks.x = element_line(linewidth = 0.2, color = "black"),
      axis.ticks.y = element_line(linewidth = 0.2, color = "black"),
      axis.ticks.length = unit(1.5, "pt"),
      # 小刻度：需要单独控制
      axis.minor.ticks.x = element_line(linewidth = 0.1, color = "black"),
      axis.minor.ticks.length.x = unit(0.8, "pt"),
      # 图例
      legend.position = "none",
      strip.clip = "off",
      plot.margin = margin(t = 4, r = 2, b = 0, l = 2)
      
      
      
    ) +
    annotate(
      "text",
      x = 1,
      y = kw_y,
      hjust = 0,
      label = kruskal_p_text,
      size = 1.8,
      color = "black"
    )
  
  for (i in 1:nrow(comparisons)) {
    group <- unlist(strsplit(comparisons$group[i], " - "))
    p_value <- comparisons$p_value[i]
    
    significance <- ifelse(
      p_value < 0.001, "***",
      ifelse(p_value < 0.01, "**",
             ifelse(p_value < 0.05, "*", "ns"))
    )
    
    boxplot <- boxplot +
      geom_signif(
        comparisons = list(group),
        annotations = significance,
        y_position = sig_start + sig_step * (i - 1),
        textsize = 1.3,
        vjust = 0.1,
        margin_top = 0.01,
        map_signif_level = FALSE,
        size = 0.25
      )
  }
  
  return(list(boxplot = boxplot, comparisons = comparisons))
}
```

## Function C: remove significance stars

```{r}
#| echo: true
#| warning: false


report_probe3 <- function(data_raw, probe, exclude_samples, custom_order, custom_colors) {
  
  subset_data <- data_raw[data_raw$Probe == probe, ]
  subset_data <- subset_data[!subset_data$Sample %in% exclude_samples, ]
  
  # ===== Y-axis =====
  y_lim <- c(0, 6)
  
  boxplot <- ggplot(subset_data, aes(x = Sample, y = Number, fill = Sample)) +
    geom_boxplot(
      linewidth = 0.2,
      colour = "grey65",
      box.colour = "grey65",
      whisker.colour = "grey65",
      median.colour = "black",
      median.linewidth = 0.3,
      outlier.color = NA,
      fatten = 1.5
    )+
    labs(
      x = "",
      y = bquote("" * italic(.(probe)) * " Probe Foci")
    ) +
    scale_x_discrete(limits = custom_order) +
    scale_fill_manual(values = custom_colors) +
    scale_y_continuous(
      expand = c(0.02, 0),
      limits = y_lim,
      breaks = seq(0, 5, by = 1),
      minor_breaks = NULL
    ) +
    theme_bw(base_family = "Arial") +
    theme(
      # Transparent background
      panel.background = element_rect(fill = "transparent", color = NA),
      plot.background  = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(fill = "transparent", color = NA),
      
      panel.grid = element_blank(),
      panel.spacing.x = unit(0.6, "lines"),
      panel.spacing.y = unit(0.5, "lines"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.3),
      
      # X 轴文字旋转 45 度
      axis.text.x = element_text(
        size = 6,
        color = "black",
        angle = 45,
        hjust = 1,
        vjust = 1
      ),
      axis.text.y = element_text(size = 6, color = "black"),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 7),
      
      # 坐标轴刻度
      axis.ticks.x = element_line(linewidth = 0.2, color = "black"),
      axis.ticks.y = element_line(linewidth = 0.2, color = "black"),
      axis.ticks.length = unit(1.5, "pt"),
      
      # 不显示图例
      legend.position = "none",
      strip.clip = "off",
      plot.margin = margin(t = 2, r = 1, b = 1, l = 1),
      text = element_text(family = "Arial")
    )
  
  return(list(
    boxplot = boxplot,
    data = subset_data
  ))
}

```

# Plot for every probe

## Prepare colors

```{r}
#| echo: true
#| warning: false
#| fig-align: center
#| fig-width: 6
#| fig-height: 3

# Take three type of color from viridis
#cols <- viridis(3)
cols <- c("#DCCC77","#88CDEE","#D68C99")
# Have a preview of those colors
barplot(1:3, col = cols, names.arg = c("0%", "50%", "100%"))

# Define the order and color of each sample
custom_order <- c("Ras3", "ML82", "HYC1", "HYC2", "HYC3", "HYC4")
custom_colors <- c("Ras3" = cols[1], "ML82" = cols[3], "HYC1" = cols[2], "HYC2" = cols[2], "HYC3" = cols[2], "HYC4" = cols[2])
```

## *Dmel/Dsim* probe

```{r}
#| echo: true
#| warning: false
#| fig-align: center
#| fig-width: 5
#| fig-height: 6


probe1 <- report_probe3(data_raw, "Dmel/Dsim", exclude_samples=c("SL2"), custom_order, custom_colors)

probe1_boxplot_canvas <- probe1$boxplot + ggview::canvas(46.6, 32.6, units="mm",dpi=900,bg="transparent")
probe1_boxplot_canvas

#ggsave( "fig1C.pdf",probe1$boxplot,dpi=1200,units="mm",width=46.6, height=32.6, device = cairo_pdf)
```

## *Dsim* probe

```{r}
#| echo: true
#| warning: false
#| fig-align: center
#| fig-width: 5
#| fig-height: 6


probe23 <- report_probe3(data_raw, "Dsim", exclude_samples=c("SL2"), custom_order, custom_colors)

probe23_boxplot_canvas <- probe23$boxplot + ggview::canvas(46.6, 32.6, units="mm",dpi=900,bg="transparent")
probe23_boxplot_canvas

#ggsave( "fig1E.pdf",probe23$boxplot,dpi=1200,units="mm",width=46.6, height=32.6,device = cairo_pdf)
```

## *Dmel* probe

```{r}
#| echo: true
#| warning: false
#| fig-align: center
#| fig-width: 5
#| fig-height: 6


probe24 <- report_probe3(data_raw, "Dmel", exclude_samples=c("SL2"), custom_order, custom_colors)

probe24_boxplot_canvas <- probe24$boxplot + ggview::canvas(46.6, 32.6, units="mm",dpi=900,bg="transparent")
probe24_boxplot_canvas

#ggsave( "fig1G.pdf",probe24$boxplot,dpi=1200,units="mm",width=46.6, height=32.6,device = cairo_pdf)
```
