---
title: "Fig.3D, S9 Comparative analysis of relative expression level"
author: "Jie Hua"
date: 2025-12-03
lang: en
date-modified: 2026-05-26
keywords: ["Dosage compensation","Transcriptome","Comparison"]
format:
  html:
    theme: cosmo
    page-layout: full
    toc: true
    toc-location: left
    section-divs: true
    code-fold: false
    code-overflow: wrap
    code-line-numbers: true
    code-copy: hover
    number-sections: true
    highlight-style: tango
editor: visual
---

# Load required packages

```{r}
#| echo: true
#| warning: false
#| message: false

library(data.table)
library(stringr)
library(pbapply)
library(tidyr)
library(dplyr)
library(limma)
library(clusterProfiler)
library(ggplot2)
library(ggrepel)
library(viridis)
library(org.Dhybrid.eg.db)
library(tibble)
library(ape)
library(ggtree)
library(patchwork)
library(RColorBrewer)
library(scales)
library(tidytext)
```

# Load the files

## Define a folder

```{r}
#| echo: true
#| warning: false

folder_path <- "H:\\HybridCells_Data\\Fig3_Dosage and Transcriptome"
#folder_path <- "F:\\Big Paper\\Processed data\\Fig3C_Dosage and Transcriptome"
```

## Load the matrix from RNA/DNA

```{r}
#| echo: true
#| warning: false

comparison_dt <- fread(paste0(folder_path,"\\comparison.csv"),header = T)
```

## Load the orthofinder matrix generated with protein-protein.

```{r}
#| echo: true
#| warning: false

mel_sim_orthofinder <- fread(file=paste0(folder_path,"\\dmel_dsim_orthofinder\\Orthogroups.tsv"),header = T,sep = "\t")
```

## Load the annotation GTF from IsoQuant

```{r}
#| echo: true
#| warning: false

isoquant_transcript_gtf <- paste0(folder_path,"\\transcriptome\\All.extended_annotation.gtf")

```

## Load the annotations from eggnogmapper

```{r}
#| echo: true
#| warning: false

eggnog_annotation_file <- paste0(folder_path,"\\proteome_annotation\\Isoquant.transcripts.fa.eggnog.emapper.annotations")

```

# Read the eggnog-mapper annotation

Read the table

```{r}
#| echo: true 
#| warning: false  

col_names <- c(   "query", "seed_ortholog", "evalue", "score", "eggNOG_OGs", "max_annot_lvl",   "COG_category", "Description", "Preferred_name", "GOs", "EC", "KEGG_ko",   "KEGG_Pathway", "KEGG_Module", "KEGG_Reaction", "KEGG_rclass", "BRITE",   "KEGG_TC", "CAZy", "BiGG_Reaction", "PFAMs" )

eggnog_annotation <- read.table(eggnog_annotation_file,sep = "\t",header = FALSE,comment.char = "#",stringsAsFactors = FALSE,quote = "")  

colnames(eggnog_annotation) <- col_names 
```

According to orthogroups information, build orthogroup-GO relationship.

```{r}
#| echo: true
#| warning: false

# 1. 建立蛋白 -> GO 映射，并处理空值
eggnog_map <- eggnog_annotation[, c("query", "GOs")]
eggnog_map$GOs[eggnog_map$GOs == "-"] <- ""

# 2. 定义函数：输入蛋白ID向量或逗号分隔字符串，返回 GO 并集
get_go_union <- function(protein_str) {
  # 拆分蛋白ID
  proteins <- trimws(unlist(strsplit(protein_str, ",")))
  
  # 找到对应的 GOs
  gos <- eggnog_map$GOs[eggnog_map$query %in% proteins]
  
  # 拆分每个 GO 字符串成单个 GO
  gos_list <- unlist(strsplit(gos, ","))
  
  # 去掉空值，取唯一并排序
  unique_gos <- sort(unique(gos_list[gos_list != ""]))
  
  # 合并成字符串
  paste(unique_gos, collapse = ",")
}

# 3. 对 mel_sim_orthofinder 两列应用 pbapply
# mel列
mel_sim_orthofinder$GO_union_mel <- pbapply::pbsapply(
  mel_sim_orthofinder$isoquant_split.mel.pep.nostar, 
  get_go_union
)
# Sim
mel_sim_orthofinder$GO_union_sim <- pbapply::pbsapply(
  mel_sim_orthofinder$isoquant_split.sim.pep.nostar, 
  get_go_union
)


# 4. 查看结果
head(mel_sim_orthofinder[, c("GO_union_mel", "GO_union_sim")])



# 函数：取两个逗号分隔 GO 字符串的并集
get_go_union_two <- function(go1, go2) {
  go1_split <- unlist(strsplit(go1, ","))
  go2_split <- unlist(strsplit(go2, ","))
  
  go1_split <- go1_split[go1_split != ""]
  go2_split <- go2_split[go2_split != ""]
  
  union_go <- sort(unique(c(go1_split, go2_split)))
  paste(union_go, collapse = ",")
}

# 使用 pbapply 的 pbsapply 显示进度条
mel_sim_orthofinder$GO_union_all <- pbapply::pbsapply(
  seq_len(nrow(mel_sim_orthofinder)), 
  function(i) get_go_union_two(
    mel_sim_orthofinder$GO_union_mel[i], 
    mel_sim_orthofinder$GO_union_sim[i]
  )
)

```

# Covert protein-protein relationship to transcript-transcript relationship

```{r}
#| echo: true
#| warning: false

orthofinder <- as.data.table(mel_sim_orthofinder)

# 需要处理的列

cols_to_clean <- c("isoquant_split.mel.pep.nostar", "isoquant_split.sim.pep.nostar")

for(col in cols_to_clean){
  orthofinder[[col]] <- pbsapply(orthofinder[[col]], function(x){
    if(is.na(x) || x == "") return(NA)
    # 按逗号分割
    ids <- unlist(strsplit(x, ",\\s*"))
    # 去掉 .p1/.p2/.p3 后缀
    ids <- str_replace(ids, "\\.p\\d+$", "")
    # 去重
    ids <- unique(ids)
    # 合并回单元格
    paste(ids, collapse = ", ")
  })
}

orthofinder_subset <- orthofinder[, c("Orthogroup", "GO_union_all")]

# ----------------------------
# 2. 保存为 CSV
# ----------------------------
#write.csv(orthofinder_subset, file = "orthofinder_term2gene.csv", row.names = FALSE)
```

# Read GTF file and build transcript-gene relationship

```{r}
#| echo: true
#| warning: false


# Define a function to get the mapping relationship
read_isoquant_gtf <- function(isoquant_transcript_gtf){
  # Read GTF file
  gtf <- fread(isoquant_transcript_gtf, sep="\t", header=FALSE, data.table=TRUE, quote = "")
  
  # Name the columns
   colnames(gtf) <- c("chr","source","type","start","end","score","strand","frame","attributes")
  
  # Extract only transcripts
  trans <- gtf[type == "transcript"]
  
  # Extract gene_id and transcript_id
  trans[, gene_id := str_match(attributes, 'gene_id "([^"]+)"')[,2]]
  trans[, transcript_id := str_match(attributes, 'transcript_id "([^"]+)"')[,2]]
  
  # Return the transcript-gene mapping relationship
  dt_matrix <- trans[, .(transcripts = paste(transcript_id, collapse=",")), by=gene_id]
  
  return(dt_matrix)
}

# Extract the relationship

transcript_gene <- read_isoquant_gtf(isoquant_transcript_gtf)

```

Confirm if there is duplications

```{r}
#| echo: true
#| warning: false

cat("Check the duplications in the GTF: ",anyDuplicated(transcript_gene),"\n")

# Split the multiple transcript IDs

trans_gene_expanded <- transcript_gene[, .(
  transcript = unlist(str_split(transcripts, ",\\s*"))
), by = gene_id]

cat("Check the duplications in the expanded table: ",anyDuplicated(trans_gene_expanded),"\n")

```

# Convert transcript-transcript relationship to gene-gene relationship

```{r}
#| echo: true
#| warning: false

# Set a lookup table
trans2gene <- setNames(trans_gene_expanded$gene_id, trans_gene_expanded$transcript)

# The columns that need to be converted
cols_to_replace <- c("isoquant_split.mel.pep.nostar", "isoquant_split.sim.pep.nostar")

# Perform conversion
for(col in cols_to_replace){
  orthofinder[[col]] <- pbsapply(orthofinder[[col]], function(x){
    if(is.na(x) || x == "") return(NA_character_)
    
    # 拆分逗号分隔的 transcript
    transcripts <- str_split(x, ",\\s*")[[1]]
    
    # 替换为 gene_id，如果找不到就报错
    genes <- sapply(transcripts, function(tr){
      gene <- trans2gene[tr]
      if(is.na(gene)) stop(paste("Transcript not found in lookup:", tr))
      gene
    })
    
    # 去重并合并
    paste(unique(genes), collapse=",")
  }, cl = 1)  # cl = 1 表示单线程，如果想用多线程可以改
}


#write.csv(orthofinder,file="orthofinder_gene.csv",row.names = F)
```

# Calculate the Allele-specific expression

```{r}
#| echo: true
#| warning: false

# Step 0: 准备表达矩阵
gene_expr_mat <- as.matrix(comparison_dt[, -c("chrom","start","end","gene"), with=FALSE])
rownames(gene_expr_mat) <- comparison_dt$gene
sample_cols <- colnames(gene_expr_mat)

# Step 1: 构建基因名字占位表格
orthogroup_dt <- data.table(Orthogroup = orthofinder$Orthogroup)

# 将每个 Orthogroup 的 mel/sim 基因拆开，按列填充
for(s in sample_cols){
  orthogroup_dt[[paste0(s, "_mel")]] <- str_split(orthofinder$isoquant_split.mel.pep.nostar, ",\\s*")
  orthogroup_dt[[paste0(s, "_sim")]] <- str_split(orthofinder$isoquant_split.sim.pep.nostar, ",\\s*")
}

# Step 2: 用 pbapply 遍历每个单元格，把基因名替换成表达量加和
orthogroup_expr_dt <- copy(orthogroup_dt)

cols_to_sum <- setdiff(colnames(orthogroup_expr_dt), "Orthogroup")

for(i in seq_along(cols_to_sum)){
  col <- cols_to_sum[i]
  cat(sprintf("Processing column %d/%d: %s\n", i, length(cols_to_sum), col))
  
  orthogroup_expr_dt[[col]] <- pbsapply(orthogroup_expr_dt[[col]], function(genes){
    genes <- genes[genes != ""]       # 移除空字符串
    genes_exist <- genes[genes %in% rownames(gene_expr_mat)]
    if(length(genes_exist) == 0) return(0)
    sum(gene_expr_mat[genes_exist, str_remove(col, "_mel|_sim"), drop=TRUE], na.rm = TRUE)
  })
}

```

Have a quick check of the table

```{r}
#| echo: true
#| warning: false

head(orthogroup_expr_dt)

```

# Save all objects to .RData

```{r}
#| echo: true
#| warning: false

#save.image(file = "data.RData")

load("data.RData")

```

# Enrichment analysis of orthogroups

```{r}
#| echo: true
#| warning: false

# 假设 orthogroup_expr_dt 已经是 data.table
dt <- copy(orthogroup_expr_dt)

# 找出所有样本名（去掉 _mel/_sim），假设所有列都符合 "sample_repX_mel/sim" 格式
all_cols <- colnames(dt)
expr_cols <- all_cols[!all_cols %in% "Orthogroup"]

# 提取样本前缀（去掉 _mel/_sim）
sample_prefix <- unique(sub("_(mel|sim)$", "", expr_cols))

# 初始化新的矩阵
new_expr_dt <- data.table(Orthogroup = dt$Orthogroup)

# 对每个样本计算 mel+sim
for(s in sample_prefix){
  mel_cols <- grep(paste0("^", s, "_mel$"), expr_cols, value = TRUE)
  sim_cols <- grep(paste0("^", s, "_sim$"), expr_cols, value = TRUE)
  
  # 如果有多个重复列（rep1, rep2, rep3）就相加
  new_expr_dt[[s]] <- rowSums(dt[, c(mel_cols, sim_cols), with = FALSE])
}

# 查看结果
head(new_expr_dt)

```

# PCA and heatmap

```{r}
#| echo: true
#| warning: false

# 1️⃣ 提取表达矩阵
expr_mat <- as.matrix(new_expr_dt[, -1, with = FALSE])
rownames(expr_mat) <- new_expr_dt$Orthogroup

# 2️⃣ 去掉全零行（Orthogroup）
expr_mat <- expr_mat[rowSums(expr_mat) != 0, ]

# 3️⃣ 可选：去掉方差为 0 的行（或列）
expr_mat <- expr_mat[apply(expr_mat, 1, var) != 0, ]

# 4️⃣ PCA（样本为行，gene为列）
pca_res <- prcomp(t(expr_mat), scale. = TRUE)

pca_var <- pca_res$sdev^2
pca_var_perc <- round(pca_var / sum(pca_var) * 100, 1)

# 构建 PCA 数据框
pca_df <- data.frame(
  Sample = colnames(expr_mat),
  PC1 = pca_res$x[, 1],
  PC2 = pca_res$x[, 2]
)

# 可视化 PCA
pca_transcriptome <- ggplot(pca_df, aes(x = PC1, y = PC2, label = Sample)) +
  geom_point(size = 3, color = "steelblue") +
  geom_text_repel(                
    size = 5,
    family = "sans",
    color = "black",
    max.overlaps = Inf            
  ) +
  theme_gray(base_size = 12) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      size = 20,
      family = "sans",
      color = "black"
    ),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1,
      size = 15,
      family = "sans",
      color = "black"
    ),
    axis.text.y = element_text(
      size = 15,
      family = "sans",
      color = "black"
    ),
    axis.title.x = element_text(
      size = 20,
      family = "sans",
      color = "black"
    ),
    axis.title.y = element_text(
      size = 20,
      family = "sans",
      color = "black"
    ),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(fill = "transparent"),
    panel.grid = element_blank(),
    panel.spacing = unit(0.4, "lines"),
    strip.background = element_rect(
      fill = "transparent",
      color = "NA",
      linewidth = 0.8
    ),
    strip.text = element_text(
      size = 16,
      family = "sans",
      color = "black"
    ),
    legend.position = "none"
  ) +
  ggtitle("PCA of Samples") +
  # 横纵轴显示方差百分比
  xlab(paste0("PC1 (", pca_var_perc[1], "%)")) +
  ylab(paste0("PC2 (", pca_var_perc[2], "%)"))

```

ggplot2 version heatmap

```{r}
#| echo: true
#| warning: false

source("functions.R")

res_heatmap2 <- ggtree_heatmap(
  mat = expr_mat,
  scale = "row",
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  show_row_tree = FALSE,
  show_col_tree = TRUE,
  show_rownames = FALSE,
  show_colnames = TRUE,
  border_color = NA,
  legend_name = "Z-score",
  col_angle = 45
)

#res_heatmap2$plot

```

Save plot

```{r}
#| echo: true
#| warning: false

pca_transcriptome_canvas <- pca_transcriptome + ggview::canvas(7.5, 7.5, units="in")

pca_transcriptome_canvas
#ggsave("pca_transcrioptome.pdf",pca_transcriptome,width=7.5, height=7.5,units="in",dpi = 1200)


```

# Comparative analysis

## H4 to Ras3

```{r}
#| echo: true
#| warning: false

#1. 提取需要比较的列
# 取表达矩阵（去掉 Orthogroup 列）
expr_mat <- as.matrix(new_expr_dt[, -1, with = FALSE])
rownames(expr_mat) <- new_expr_dt$Orthogroup

expr_mat <- log2(expr_mat + 1)

# 选择 Ras3 和 7B 的列
ras3_cols <- grep("^Ras3", colnames(expr_mat), value = TRUE)
b7_cols   <- grep("^7B", colnames(expr_mat), value = TRUE)

expr_sub <- expr_mat[, c(ras3_cols, b7_cols)]

#2. 构建分组信息
group <- factor(c(rep("B7", length(b7_cols)), rep("Ras3", length(ras3_cols))))
design <- model.matrix(~group)
colnames(design) <- c("Intercept", "B7_vs_Ras3")

#3. 使用 limma 进行差异分析
fit <- lmFit(expr_sub, design)
fit <- eBayes(fit)

# 查看差异表达结果
res_all <- topTable(fit, coef = "B7_vs_Ras3", number = Inf, sort.by = "none")
head(res_all)

```

Perform Enrich analysis

```{r}
#| echo: true
#| warning: false

# ----------------------------
# 提取上调和下调蛋白质
# ----------------------------
# 设置阈值
padj_cutoff <- 0.05
logFC_cutoff <- 1

up_genes   <- rownames(res_all)[res_all$adj.P.Val < padj_cutoff & res_all$logFC > logFC_cutoff]
down_genes <- rownames(res_all)[res_all$adj.P.Val < padj_cutoff & res_all$logFC < -logFC_cutoff]
```

up

```{r}
#| echo: true
#| warning: false


ego_bp <- enrichGO(
  gene = up_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


goplot(ego_bp)

ego_cc <- enrichGO(
  gene = up_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = up_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)
```

```{r}
#| echo: true
#| warning: false

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)
```

Save the Plot

```{r}
#| echo: true
#| warning: false

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("H4_Ras3_mel_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("H4_Ras3_mel_GO_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")
```

Other plots

```{r}
#| echo: true
#| warning: false

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
source("functions.R")
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)


plot_ego_up_emapplot+ggview::canvas(90,90,"mm",600,bg="transparent")
```

Save the plot

```{r}
#| echo: true
#| warning: false

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("emap_pub.png",p_emap_pub,width=160,height=90,units="mm",bg="transparent")

```

down

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = down_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


goplot(ego_bp)

ego_cc <- enrichGO(
  gene = down_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = down_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("sim_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("sim_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("emap_pub.png",p_emap_pub,width=160,height=90,units="mm",bg="transparent")

```

## H4_mel to H4_sim

Took 7B samples out.

```{r}
#| echo: true
#| warning: false


data_7B <- orthogroup_expr_dt[
  , c("Orthogroup", grep("^7B", names(orthogroup_expr_dt), value = TRUE)),
  with = FALSE
]

```

```{r}
#| echo: true
#| warning: false

# 假设 orthogroup_expr_dt 是 data.table
# 1. 提取表达矩阵，去掉第一列 Orthogroup
expr_mat <- as.matrix(orthogroup_expr_dt[, -1, with = FALSE])
rownames(expr_mat) <- orthogroup_expr_dt$Orthogroup

# 可选：log2 转换
expr_mat <- log2(expr_mat + 1)

# 2. 选择 7B 的列
b7_mel_cols <- grep("^7B.*_mel$", colnames(expr_mat), value = TRUE)
b7_sim_cols <- grep("^7B.*_sim$", colnames(expr_mat), value = TRUE)

expr_sub <- expr_mat[, c(b7_mel_cols, b7_sim_cols)]

# 3. 构建分组信息
group <- factor(c(rep("mel", length(b7_mel_cols)), rep("sim", length(b7_sim_cols))))
design <- model.matrix(~group)
colnames(design) <- c("Intercept", "mel_vs_sim")

# 4. 使用 limma 进行差异分析
fit <- lmFit(expr_sub, design)
fit <- eBayes(fit)

# 5. 查看差异表达结果
res_7B <- topTable(fit, coef = "mel_vs_sim", number = Inf, sort.by = "none")

head(res_7B)
```

```{r}
#| echo: true
#| warning: false


# 统计数量
n_up   <- sum(res_7B$adj.P.Val < 0.05 & res_7B$logFC > 1)
n_down <- sum(res_7B$adj.P.Val < 0.05 & res_7B$logFC < -1)

# 分类显著性
res_7B$Significance <- "NotSig"
res_7B$Significance[res_7B$adj.P.Val < 0.05 & res_7B$logFC > 1] <- "Up"
res_7B$Significance[res_7B$adj.P.Val < 0.05 & res_7B$logFC < -1] <- "Down"

# 计算 -log10(adj.P.Val)
res_7B$negLogAdjP <- -log10(res_7B$adj.P.Val + 1e-10) # 防止log(0)
```

Plot

```{r}
#| echo: true
#| warning: false

# ----------------------------
# 提取上调和下调蛋白质
# ----------------------------
# 设置阈值
padj_cutoff <- 0.05
logFC_cutoff <- 1

mel_genes   <- rownames(res_7B)[res_7B$adj.P.Val < padj_cutoff & res_7B$logFC > logFC_cutoff]
sim_genes <- rownames(res_7B)[res_7B$adj.P.Val < padj_cutoff & res_7B$logFC < -logFC_cutoff]
```

Mel specific

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("H4_mel_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("mel_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```

Sim bias

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("h4_dsim_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("dsim_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("dsim_treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("dsim_emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```

## H1_mel to H1_sim

Took 5A samples out.

```{r}
#| echo: true
#| warning: false


data_5A <- orthogroup_expr_dt[
  , c("Orthogroup", grep("^5A", names(orthogroup_expr_dt), value = TRUE)),
  with = FALSE
]

```

```{r}
#| echo: true
#| warning: false

# 假设 orthogroup_expr_dt 是 data.table
# 1. 提取表达矩阵，去掉第一列 Orthogroup
expr_mat <- as.matrix(orthogroup_expr_dt[, -1, with = FALSE])
rownames(expr_mat) <- orthogroup_expr_dt$Orthogroup

# 可选：log2 转换
expr_mat <- log2(expr_mat + 1)

# 2. 选择 7B 的列
A5_mel_cols <- grep("^5A.*_mel$", colnames(expr_mat), value = TRUE)
A5_sim_cols <- grep("^5A.*_sim$", colnames(expr_mat), value = TRUE)

expr_sub <- expr_mat[, c(A5_mel_cols, A5_sim_cols)]

# 3. 构建分组信息
group <- factor(c(rep("mel", length(A5_mel_cols)), rep("sim", length(A5_sim_cols))))
design <- model.matrix(~group)
colnames(design) <- c("Intercept", "mel_vs_sim")

# 4. 使用 limma 进行差异分析
fit <- lmFit(expr_sub, design)
fit <- eBayes(fit)

# 5. 查看差异表达结果
res_5A <- topTable(fit, coef = "mel_vs_sim", number = Inf, sort.by = "none")

head(res_5A)
```

```{r}
#| echo: true
#| warning: false


# 统计数量
n_up   <- sum(res_5A$adj.P.Val < 0.05 & res_5A$logFC > 1)
n_down <- sum(res_5A$adj.P.Val < 0.05 & res_5A$logFC < -1)

# 分类显著性
res_5A$Significance <- "NotSig"
res_5A$Significance[res_5A$adj.P.Val < 0.05 & res_5A$logFC > 1] <- "Up"
res_5A$Significance[res_5A$adj.P.Val < 0.05 & res_5A$logFC < -1] <- "Down"

# 计算 -log10(adj.P.Val)
res_5A$negLogAdjP <- -log10(res_5A$adj.P.Val + 1e-10) # 防止log(0)
```

Plot

```{r}
#| echo: true
#| warning: false

# ----------------------------
# 提取上调和下调蛋白质
# ----------------------------
# 设置阈值
padj_cutoff <- 0.05
logFC_cutoff <- 1

mel_genes   <- rownames(res_5A)[res_5A$adj.P.Val < padj_cutoff & res_5A$logFC > logFC_cutoff]
sim_genes <- rownames(res_5A)[res_5A$adj.P.Val < padj_cutoff & res_5A$logFC < -logFC_cutoff]
```

Mel specific

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("h1_mel_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("mel_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```

Sim bias

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("dsim_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("dsim_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("dsim_treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("dsim_emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```

## H2_mel to H2_sim

Took 5A samples out.

```{r}
#| echo: true
#| warning: false


data_5B <- orthogroup_expr_dt[
  , c("Orthogroup", grep("^5B", names(orthogroup_expr_dt), value = TRUE)),
  with = FALSE
]

```

```{r}
#| echo: true
#| warning: false

# 假设 orthogroup_expr_dt 是 data.table
# 1. 提取表达矩阵，去掉第一列 Orthogroup
expr_mat <- as.matrix(orthogroup_expr_dt[, -1, with = FALSE])
rownames(expr_mat) <- orthogroup_expr_dt$Orthogroup

# 可选：log2 转换
expr_mat <- log2(expr_mat + 1)

# 2. 选择 7B 的列
B5_mel_cols <- grep("^5B.*_mel$", colnames(expr_mat), value = TRUE)
B5_sim_cols <- grep("^5B.*_sim$", colnames(expr_mat), value = TRUE)

expr_sub <- expr_mat[, c(B5_mel_cols, B5_sim_cols)]

# 3. 构建分组信息
group <- factor(c(rep("mel", length(B5_mel_cols)), rep("sim", length(B5_sim_cols))))
design <- model.matrix(~group)
colnames(design) <- c("Intercept", "mel_vs_sim")

# 4. 使用 limma 进行差异分析
fit <- lmFit(expr_sub, design)
fit <- eBayes(fit)

# 5. 查看差异表达结果
res_5B <- topTable(fit, coef = "mel_vs_sim", number = Inf, sort.by = "none")

head(res_5B)
```

```{r}
#| echo: true
#| warning: false


# 统计数量
n_up   <- sum(res_5B$adj.P.Val < 0.05 & res_5B$logFC > 1)
n_down <- sum(res_5B$adj.P.Val < 0.05 & res_5B$logFC < -1)

# 分类显著性
res_5B$Significance <- "NotSig"
res_5B$Significance[res_5B$adj.P.Val < 0.05 & res_5B$logFC > 1] <- "Up"
res_5B$Significance[res_5B$adj.P.Val < 0.05 & res_5B$logFC < -1] <- "Down"

# 计算 -log10(adj.P.Val)
res_5B$negLogAdjP <- -log10(res_5B$adj.P.Val + 1e-10) # 防止log(0)
```

Plot

```{r}
#| echo: true
#| warning: false

# ----------------------------
# 提取上调和下调蛋白质
# ----------------------------
# 设置阈值
padj_cutoff <- 0.05
logFC_cutoff <- 1

mel_genes   <- rownames(res_5B)[res_5B$adj.P.Val < padj_cutoff & res_5B$logFC > logFC_cutoff]
sim_genes <- rownames(res_5B)[res_5B$adj.P.Val < padj_cutoff & res_5B$logFC < -logFC_cutoff]
```

Mel specific

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("h2mel_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("mel_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```

Sim bias

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("dsim_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("dsim_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("dsim_treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("dsim_emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```

## H3_mel to H3_sim

Took 5A samples out.

```{r}
#| echo: true
#| warning: false


data_6B <- orthogroup_expr_dt[
  , c("Orthogroup", grep("^6B", names(orthogroup_expr_dt), value = TRUE)),
  with = FALSE
]

```

```{r}
#| echo: true
#| warning: false

# 假设 orthogroup_expr_dt 是 data.table
# 1. 提取表达矩阵，去掉第一列 Orthogroup
expr_mat <- as.matrix(orthogroup_expr_dt[, -1, with = FALSE])
rownames(expr_mat) <- orthogroup_expr_dt$Orthogroup

# 可选：log2 转换
expr_mat <- log2(expr_mat + 1)

# 2. 选择 7B 的列
B6_mel_cols <- grep("^6B.*_mel$", colnames(expr_mat), value = TRUE)
B6_sim_cols <- grep("^6B.*_sim$", colnames(expr_mat), value = TRUE)

expr_sub <- expr_mat[, c(B6_mel_cols, B6_sim_cols)]

# 3. 构建分组信息
group <- factor(c(rep("mel", length(B6_mel_cols)), rep("sim", length(B6_sim_cols))))
design <- model.matrix(~group)
colnames(design) <- c("Intercept", "mel_vs_sim")

# 4. 使用 limma 进行差异分析
fit <- lmFit(expr_sub, design)
fit <- eBayes(fit)

# 5. 查看差异表达结果
res_6B <- topTable(fit, coef = "mel_vs_sim", number = Inf, sort.by = "none")

head(res_6B)
```

```{r}
#| echo: true
#| warning: false


# 统计数量
n_up   <- sum(res_6B$adj.P.Val < 0.05 & res_6B$logFC > 1)
n_down <- sum(res_6B$adj.P.Val < 0.05 & res_6B$logFC < -1)

# 分类显著性
res_6B$Significance <- "NotSig"
res_6B$Significance[res_6B$adj.P.Val < 0.05 & res_6B$logFC > 1] <- "Up"
res_6B$Significance[res_6B$adj.P.Val < 0.05 & res_6B$logFC < -1] <- "Down"

# 计算 -log10(adj.P.Val)
res_6B$negLogAdjP <- -log10(res_6B$adj.P.Val + 1e-10) # 防止log(0)
```

Plot

```{r}
#| echo: true
#| warning: false

# ----------------------------
# 提取上调和下调蛋白质
# ----------------------------
# 设置阈值
padj_cutoff <- 0.05
logFC_cutoff <- 1

mel_genes   <- rownames(res_6B)[res_6B$adj.P.Val < padj_cutoff & res_6B$logFC > logFC_cutoff]
sim_genes <- rownames(res_6B)[res_6B$adj.P.Val < padj_cutoff & res_6B$logFC < -logFC_cutoff]
```

Mel specific

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = mel_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("h3_mel_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("mel_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```

Sim bias

```{r}
#| echo: true
#| warning: false

ego_bp <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)


ego_cc <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "CC",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

ego_mf <- enrichGO(
  gene = sim_genes,
  OrgDb = org.Dhybrid.eg.db,
  keyType = "GID",
  ont = "MF",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.05,
  readable = TRUE
)

#---------------------------
# 2. drop level 1:2
#---------------------------
ego_bp2 <- dropGO(ego_bp, level = 1:2)
ego_cc2 <- dropGO(ego_cc, level = 1:2)
ego_mf2 <- dropGO(ego_mf, level = 1:2)

#---------------------------
# 3. 转成 data.frame 并手动加类别
#---------------------------
df_bp <- as.data.frame(ego_bp2)
df_cc <- as.data.frame(ego_cc2)
df_mf <- as.data.frame(ego_mf2)

df_bp$ONTOLOGY <- "BP"
df_cc$ONTOLOGY <- "CC"
df_mf$ONTOLOGY <- "MF"

# 可选：统一列顺序，防止老版本 bind_rows 时列位置乱掉
common_cols <- union(
  union(colnames(df_bp), colnames(df_cc)),
  colnames(df_mf)
)

df_bp <- df_bp[, common_cols[common_cols %in% colnames(df_bp)], drop = FALSE]
df_cc <- df_cc[, common_cols[common_cols %in% colnames(df_cc)], drop = FALSE]
df_mf <- df_mf[, common_cols[common_cols %in% colnames(df_mf)], drop = FALSE]

# 合并
df_all <- bind_rows(df_bp, df_cc, df_mf)

p_bp <- bp_dotplot(df_bp) 

p_all <- all_dotplot(df_all)

p_bp_canvas <- p_bp + ggview::canvas(80,60,units="mm",dpi=600,bg="transparent")
p_bp_canvas


p_all_canvas <- p_all + ggview::canvas(100,170,units="mm",dpi=600,bg="transparent")
p_all_canvas

#ggsave("dsim_GO.pdf",p_bp,width=80,height=60,units="mm",bg="transparent")

#ggsave("dsim_all.png",p_all,width=100,height=170,units="mm",dpi=900,bg="transparent")

ego_up_drop <- dropGO(ego_bp, level = 1:2)
edox2 <- enrichplot::pairwise_termsim(ego_up_drop)

plot_ego_up_treeplot <- treeplot(edox2, cladelab_offset=20) + hexpand(.2)
p_pub <- modify_treeplot_pub(
  plot_ego_up_treeplot,
  show_canvas = FALSE,
  make_static = TRUE
)

plot_ego_up_emapplot <- emapplot(edox2, node_label = "group")

p_emap_pub <- modify_emapplot_pub(
  plot_ego_up_emapplot,
  label_size = 18,        # 把圆上文字变大
  font_family = "sans",
  point_size_range = c(2, 7),
  ellipse_show_border = FALSE
)

p_pub + ggview::canvas(160, 70, units = "mm", dpi = 600, bg = "transparent")

#ggview::save_ggplot(p_pub,"treeplot.png",width=160,height=70,units="mm",dpi=900,bg="transparent")
#ggsave("dsim_treeplot.pdf",p_pub,width=160,height=90,units="mm",bg="transparent")
#ggsave("treeplot.png",p_pub,width=160,height=90,units="mm",dpi=900,bg="transparent")


p_emap_pub + ggview::canvas(120, 90, units = "mm", dpi = 600, bg = "transparent")
#ggsave("dsim_emap_pub.png",p_emap_pub,width=120,height=90,units="mm",bg="transparent")


```
